---
title: "Puzzle Engagement Experiment — Data Pipeline & Analysis"
author: 
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preprocessing

Data preprocessing of .json file from Firestore.

```{r load-and-flatten}
# install.packages(c("jsonlite", "dplyr", "tidyr", "purrr", "lme4", "lmtest", "pwr"))

library(jsonlite)
library(dplyr)
library(tidyr)
library(purrr)
library(lme4)
library(clubSandwich)
library(lmtest)
library(pwr)
library(knitr)

# Load raw JSON — simplifyVector = FALSE preserves irregular arrays as lists
raw_data <- fromJSON("firestore_export.json", simplifyVector = FALSE)

cat("Users:", length(raw_data$users), "\n")
cat("GameLogs:", length(raw_data$gameLogs), "\n")

# ── Flatten users ──────────────────────────────────────────────────────────────
users_df <- tibble(
  document_id = names(raw_data$users),
  data = raw_data$users
) %>%
  unnest_wider(data)

# List columns on users:
#   puzzleOrder   — array of puzzle id strings  → comma-separated string
#   resumeRounds  — array of round numbers (empty for single-session users) → comma-separated or NA
#   breakLog      — array of {timestamp, duration} maps → derive breaksTaken + totalBreakTime scalars
users_df <- users_df %>%
  mutate(
    puzzleOrder    = map_chr(puzzleOrder,
                             ~ paste(.x, collapse = ",")),
    resumeRounds   = map_chr(resumeRounds,
                             ~ if (length(.x) == 0) NA_character_ else paste(.x, collapse = ",")),
    breaksTaken    = map_int(breakLog,
                             ~ if (is.null(.x)) 0L else length(.x)),
    totalBreakTime = map_dbl(breakLog,
                             ~ if (is.null(.x)) 0 else sum(map_dbl(.x, "duration"), na.rm = TRUE))
  ) %>%
  select(-breakLog)

# ── Flatten gameLogs ───────────────────────────────────────────────────────────
gamelogs_df <- tibble(
  document_id = names(raw_data$gameLogs),
  data = raw_data$gameLogs
) %>%
  unnest_wider(data)

# rawSubmissionTimestamps is a variable-length list; drop it for the master join.
# Filter replays — isReplay == TRUE marks repeated attempts after completion.
# Exception: sharedDevice == "new_player" means a different person on the same household
# device; the app marks their gamelogs isReplay=TRUE but they are genuinely new participants.
new_player_ids <- users_df %>%
  filter(!is.na(sharedDevice) & sharedDevice == "new_player") %>%
  pull(userId)

gamelogs_df <- gamelogs_df %>%
  select(-rawSubmissionTimestamps) %>%
  filter(is.na(isReplay) | isReplay == FALSE | userId %in% new_player_ids)

# ── Master join ────────────────────────────────────────────────────────────────
# Drop treatment/isReplay from gameLogs (canonical copies live on users).
# sessionComplete is on users (true once all 5 puzzles done), not on gamelogs.
experiment_master <- gamelogs_df %>%
  select(-treatment, -isReplay) %>%
  left_join(users_df %>% arrange(userId, desc(sessionComplete)) %>% distinct(userId, .keep_all = TRUE), by = "userId") %>%
  select(-starts_with("document_id"))

# Remove personally identifying information
#experiment_master <- experiment_master %>%
#  select(-any_of(c(
#    "ipAddress",
#    "studyId",
#    "sessionId"
#  )))

glimpse(experiment_master)
```

```{r}
# One row per randomized participant
participant_df <- users_df %>%
  arrange(userId, desc(sessionComplete)) %>%
  distinct(userId, .keep_all = TRUE) %>%
  select(-any_of(c(
    "ipAddress",
    "prolificPid",
    "studyId",
    "sessionId"
  )))

# Overall participation and completion
participant_df %>%
  summarise(
    user_documents = nrow(users_df),
    unique_participants = n(),
    completed = sum(sessionComplete == TRUE, na.rm = TRUE),
    incomplete = sum(sessionComplete == FALSE, na.rm = TRUE),
    missing_completion = sum(is.na(sessionComplete))
  )

# Randomization and completion by treatment
participant_df %>%
  count(treatment, sessionComplete)

# Number of rounds recorded per participant
round_counts <- experiment_master %>%
  group_by(userId) %>%
  summarise(
    logged_rounds = n_distinct(roundNumber),
    game_log_rows = n(),
    .groups = "drop"
  )

participant_check <- participant_df %>%
  left_join(round_counts, by = "userId") %>%
  mutate(
    logged_rounds = replace_na(logged_rounds, 0L),
    game_log_rows = replace_na(game_log_rows, 0L)
  )

participant_check %>%
  count(treatment, sessionComplete, logged_rounds)
```


## Covariate Balance Check

Check randomization frequencies across treatment arms.

```{r balance-check}
user_level <- experiment_master %>%
  distinct(userId, .keep_all = TRUE)

user_level %>%
  count(treatment) %>%
  kable(col.names = c("Treatment", "N"), caption = "Treatment Assignment")

user_level %>%
  group_by(treatment) %>%
  summarize(
    N                  = n(),
    Age                = round(mean(age, na.rm = TRUE), 1),
    `Quant Exposure`   = round(mean(quantitativeExposure, na.rm = TRUE), 2),
    `Puzzle Skill`     = round(mean(puzzleSkill, na.rm = TRUE), 2),
    `Puzzle Enjoyment` = round(mean(puzzleEnjoyment, na.rm = TRUE), 2),
    `Energy Level`     = round(mean(energyLevel, na.rm = TRUE), 2),
    `% Desktop`        = round(mean(deviceType == "Desktop", na.rm = TRUE) * 100, 1)
  ) %>%
  kable(caption = "Covariate Balance by Treatment")

user_level %>%
  count(treatment, gender) %>%
  group_by(treatment) %>%
  mutate(pct = paste0(round(n / sum(n) * 100, 0), "%")) %>%
  kable(col.names = c("Treatment", "Gender", "N", "%"), caption = "Gender by Treatment")




```

## Attrition
```{r}
# ── Attrition by treatment ────────────────────────────────────────────────────

# Count returning/replay sessions excluded from the attrition denominator
returning_exclusions <- users_df %>%
  filter(sharedDevice == "returning") %>%
  distinct(document_id, .keep_all = TRUE)

cat(
  "Returning/replay sessions excluded:",
  nrow(returning_exclusions),
  "\n"
)

# One record per eligible randomized participant
# Keep ordinary sessions and legitimate new players on shared devices;
# exclude returning participants who replayed the experiment.
participant_attrition <- users_df %>%
  filter(
    is.na(sharedDevice) |
      sharedDevice == "new_player"
  ) %>%
  arrange(userId, desc(sessionComplete)) %>%
  distinct(userId, .keep_all = TRUE)

# Count distinct retained game-log rounds for each eligible participant
participant_round_counts <- experiment_master %>%
  group_by(userId) %>%
  summarize(
    logged_rounds = n_distinct(roundNumber),
    .groups = "drop"
  )

# Retain eligible participants with zero game logs
participant_attrition <- participant_attrition %>%
  left_join(participant_round_counts, by = "userId") %>%
  mutate(
    logged_rounds = replace_na(logged_rounds, 0L),

    round_status = case_when(
      logged_rounds == 0 ~ "0 rounds",
      logged_rounds >= 1 & logged_rounds <= 4 ~ "1–4 rounds",
      logged_rounds == 5 ~ "5 rounds",
      TRUE ~ "Other"
    ),

    round_status = factor(
      round_status,
      levels = c("0 rounds", "1–4 rounds", "5 rounds", "Other")
    )
  )

# Attrition summary by treatment
attrition_summary <- participant_attrition %>%
  group_by(treatment) %>%
  summarize(
    N = n(),

    `Zero Logs N` = sum(logged_rounds == 0),
    `Zero Logs %` =
      round(mean(logged_rounds == 0) * 100, 1),

    `Partial (1–4) N` =
      sum(logged_rounds >= 1 & logged_rounds <= 4),
    `Partial (1–4) %` =
      round(mean(logged_rounds >= 1 & logged_rounds <= 4) * 100, 1),

    `Five Rounds N` = sum(logged_rounds == 5),
    `Five Rounds %` =
      round(mean(logged_rounds == 5) * 100, 1),

    .groups = "drop"
  )

attrition_summary %>%
  knitr::kable(
    caption = paste(
      "Observed Puzzle-Round Completion by Treatment",
      "Among Eligible Participants"
    )
  )


```

```{r}
#Add Attrition by Number of Rounds (Puzzles) Completed 
treatment_totals <- participant_attrition %>%
  count(treatment, name = "Treatment N")

exact_round_summary <- participant_attrition %>%
  filter(logged_rounds >= 1 & logged_rounds <= 4) %>%
  count(treatment, logged_rounds, name = "N") %>%
  tidyr::complete(
    treatment,
    logged_rounds = 1:4,
    fill = list(N = 0)
  ) %>%
  left_join(treatment_totals, by = "treatment") %>%
  mutate(
    `Percent of Treatment Group` =
      round(N / `Treatment N` * 100, 1)
  )

exact_round_summary %>%
  select(treatment, logged_rounds, N, `Percent of Treatment Group`) %>%
  knitr::kable(
    col.names = c(
      "Treatment",
      "Logged Rounds",
      "N",
      "% of Treatment Group"
    ),
    caption = "Partial Completion by Exact Number of Logged Rounds"
  )
```


## Outcome Summary

```{r outcome-summary}
experiment_master %>%
  group_by(treatment) %>%
  summarize(
    `Rounds`          = n(),
    `Solve Rate`      = paste0(round(mean(isSolved, na.rm = TRUE) * 100, 1), "%"),
    `Avg Time (s)`    = round(mean(timeEngaged, na.rm = TRUE), 1),
    `Avg Attempts`    = round(mean(attemptCount, na.rm = TRUE), 2),
    `Skip Rate`       = paste0(round(mean(didSkip, na.rm = TRUE) * 100, 1), "%"),
    `Avg Rage Clicks` = round(mean(rageClicks, na.rm = TRUE), 2)
  ) %>%
  kable(caption = "Outcome Summary by Treatment")
```

## Data Cleaning

Session-level and round-level exclusions applied in order below.

```{r clean-data}
# ── Exclusion criteria ─────────────────────────────────────────────────────────
# Session-level (user excluded entirely):
#   S1. sessionComplete == FALSE: only keep fully finished 5-round sessions
#   S2. age < 18: adults only
#   S3. sharedDevice == "returning": duplicate session from same device
#   S4. prolificRejected == TRUE: flagged by Prolific platform
#   S5. total tabSwitchTimePaused > 600s across session (verified away from browser)
#   S6. any round: didSkip + tabSwitchTimePaused == 0 + timeEngaged > per-puzzle p97
#       of skip distribution — walked away without switching tabs, came back and skipped
#   S7. 3+ zero-attempt skips with timeEngaged in 30–35s — skip button unlocks at 30s;
#       clicking immediately on 3+ puzzles signals waiting out the timer, not genuine engagement

# Per-puzzle p97 skip-time thresholds (computed from all skip rounds pre-cleaning)
skip_p97 <- experiment_master %>%
  filter(didSkip == TRUE) %>%
  group_by(puzzleId) %>%
  summarize(p97_skip = quantile(timeEngaged, 0.97), .groups = "drop")

# S6: users with at least one idle-walk-away skip round
idle_skip_users <- experiment_master %>%
  left_join(skip_p97, by = "puzzleId") %>%
  filter(didSkip == TRUE &
           (is.na(tabSwitchTimePaused) | tabSwitchTimePaused == 0) &
           timeEngaged > p97_skip) %>%
  pull(userId) %>%
  unique()

# S7: users with 5 zero-attempt skips at 30–35s (skip button unlocks at 30s; immediate click = timer-waiting)
cluster_skip_users <- experiment_master %>%
  filter(didSkip == TRUE & timeEngaged >= 30 & timeEngaged <= 35) %>%
  group_by(userId) %>%
  summarize(n_cluster_skips = n(), .groups = "drop") %>%
  filter(n_cluster_skips == 5) %>%
  pull(userId)

# Compute per-user total tab-switch pause time (S5)
user_tab_paused <- experiment_master %>%
  filter(isSolved | didSkip) %>%
  group_by(userId) %>%
  summarize(total_tab_paused = sum(tabSwitchTimePaused, na.rm = TRUE), .groups = "drop")

clean <- experiment_master %>%
  filter(sessionComplete == TRUE) %>%
  filter(age >= 18) %>%
  filter(is.na(sharedDevice) | sharedDevice == "new_player") %>%
  filter(is.na(prolificRejected) | prolificRejected != TRUE) %>%
  filter(!userId %in% idle_skip_users) %>%
  filter(!userId %in% cluster_skip_users) %>%
  filter(isSolved | didSkip) %>%
  left_join(user_tab_paused, by = "userId") %>%
  filter(total_tab_paused <= 600) %>%
  select(-total_tab_paused) %>%
  mutate(
    logTE      = log(timeEngaged + 1),
    treatment  = relevel(factor(treatment), ref = "no_label"),
    puzzleId   = relevel(factor(puzzleId),  ref = "puzzle_01"),
    deviceType = factor(deviceType)
  )

cat("Idle-walk-away sessions excluded (S6):", length(idle_skip_users), "\n")
cat("Cluster-skip sessions excluded (S7):", length(cluster_skip_users), "\n")
cat("Clean rounds:", nrow(clean), "\n")
cat("Complete users:", n_distinct(clean$userId), "\n")

clean %>% count(treatment) %>% kable(col.names = c("Treatment", "Rounds"), caption = "Rounds per Treatment")

clean %>%
  group_by(puzzleId, treatment) %>%
  summarize(n = n(), skip_pct = paste0(round(mean(didSkip) * 100, 0), "%"), .groups = "drop") %>%
  pivot_wider(names_from = treatment, values_from = c(n, skip_pct)) %>%
  arrange(puzzleId) %>%
  kable(caption = "Skip Rate per Puzzle by Treatment")
```

## Exclusion Flags: AI Assistance Detection

Two-tier flag applied after deduplication on (userId, puzzleId,
roundNumber). Hard puzzles are puzzle_01 (number sequence), puzzle_02
(alphabet wheel), puzzle_05 (tile sequence). Thresholds are empirical
percentiles of solve-time distribution on first-attempt solves.

Rules: R1. Self-reported: honestyCheck == "assisted" R2a. Any single
hard puzzle solved in ≤ p5 time on first attempt (puzzle_01 ≤ 21s,
puzzle_02 ≤ 41s, puzzle_05 ≤ 23s) R2b. 2+ hard puzzles solved in ≤ p10
time on first attempt (puzzle_01 ≤ 30s, puzzle_02 ≤ 58s, puzzle_05 ≤
37s)

```{r exclusion-flags}
# Deduplicate: keep one row per user × puzzle × round
# Smarter dedup: prefer completed rows (isSolved | didSkip) over incomplete ones
dedup <- clean %>%
  arrange(userId, puzzleId, roundNumber, desc(isSolved | didSkip), didSkip) %>%
  distinct(userId, puzzleId, roundNumber, .keep_all = TRUE)

hard_puzzles <- c("puzzle_01", "puzzle_02", "puzzle_05")

# Solve-time percentile thresholds on hard puzzles (computed on clean data)
# p5  — flag any single hard puzzle solved below this time on first attempt (R2a)
# p10 — flag if 2+ hard puzzles below this time on first attempt (R2b)
p5_thresh  <- c(puzzle_01 = 21, puzzle_02 = 41, puzzle_05 = 23)
p10_thresh <- c(puzzle_01 = 30, puzzle_02 = 58, puzzle_05 = 37)

ai_flags <- dedup %>%
  mutate(
    below_p5  = puzzleId %in% hard_puzzles & isSolved & attemptCount == 1 &
                ((puzzleId == "puzzle_01" & timeEngaged <= p5_thresh["puzzle_01"]) |
                 (puzzleId == "puzzle_02" & timeEngaged <= p5_thresh["puzzle_02"]) |
                 (puzzleId == "puzzle_05" & timeEngaged <= p5_thresh["puzzle_05"])),
    below_p10 = puzzleId %in% hard_puzzles & isSolved & attemptCount == 1 &
                ((puzzleId == "puzzle_01" & timeEngaged <= p10_thresh["puzzle_01"]) |
                 (puzzleId == "puzzle_02" & timeEngaged <= p10_thresh["puzzle_02"]) |
                 (puzzleId == "puzzle_05" & timeEngaged <= p10_thresh["puzzle_05"]))
  ) %>%
  group_by(userId, prolificPid, treatment) %>%
  summarize(
    flag_R1  = any(honestyCheck == "assisted", na.rm = TRUE),  # self-reported
    flag_R2a = sum(below_p5,  na.rm = TRUE) >= 1,             # any p5 blitz
    flag_R2b = sum(below_p10, na.rm = TRUE) >= 2,             # 2+ p10 blitz
    ai_flagged = flag_R1 | flag_R2a | flag_R2b,
    .groups = "drop"
  )

# Manual overrides: participants reviewed and cleared despite automated flag
# 6977b9ea4ad70bb8a33c9b2d (local_lhde027og): p01 20s/a1 (1s below p5), p05 35s/a1 (2s below p10)
#   — margins too tight to be confident; struggled on p02 (118s/a3, tab switch); solo self-report
manual_clear <- c("local_lhde027og")

all_flags <- ai_flags %>%
  mutate(any_flag = ai_flagged & !userId %in% manual_clear)

cat("=== Exclusion summary ===\n")
cat("R1  Admitted assistance:         ", sum(ai_flags$flag_R1),  "\n")
cat("R2a Any hard puzzle below p5:    ", sum(ai_flags$flag_R2a), "\n")
cat("R2b 2+ hard puzzles below p10:   ", sum(ai_flags$flag_R2b), "\n")
cat("TOTAL excluded (R1|R2a|R2b):     ", sum(all_flags$any_flag), "\n")

# Final analysis dataset
# Dedup within (userId, puzzleId, roundNumber): prefer completed row (isSolved | didSkip)
df <- clean %>%
  left_join(all_flags %>% select(userId, any_flag), by = "userId") %>%
  filter(!any_flag) %>%
  mutate(
    zero_att_skip = as.integer(didSkip & attemptCount == 0),
    logAtt        = log(attemptCount + 1)
  ) %>%
  arrange(userId, puzzleId, roundNumber, desc(isSolved | didSkip), didSkip) %>%
  distinct(userId, puzzleId, roundNumber, .keep_all = TRUE)

cat("\nFinal dataset: ", n_distinct(df$userId), "users |", nrow(df), "rounds\n")
cat("Rounds per treatment:\n")
print(count(df, treatment))



# Create recruitment-source subsamples
df_prolific <- df %>%
  filter(!is.na(prolificPid))

df_organic <- df %>%
  filter(is.na(prolificPid))

cat("\nSample sizes:\n")
cat("Full:", n_distinct(df$userId), "users |", nrow(df), "rounds\n")
cat(
  "Prolific:",
  n_distinct(df_prolific$userId),
  "users |",
  nrow(df_prolific),
  "rounds\n"
)
cat(
  "Organic:",
  n_distinct(df_organic$userId),
  "users |",
  nrow(df_organic),
  "rounds\n"
)


```

## Primary Model 1: log(timeEngaged)

Mixed effects model with random intercept per user (rounds nested within
users). Reference: treatment = no_label, puzzleId = puzzle_01.
Cluster-robust SEs (CR1, Satterthwaite df) account for within-user
correlation. Lean spec (treatment + puzzleId + puzzleSkill) is primary;
full covariate model is robustness check.

```{r primary-model}
run_primary <- function(data) {
  # Lean model
  m_lean    <- lmer(logTE ~ treatment + puzzleId + puzzleSkill + (1 | userId),
                    data = data, REML = FALSE)
  vcov_lean <- vcovCR(m_lean, cluster = data$userId, type = "CR1")
  rt_lean   <- coef_test(m_lean, vcov = vcov_lean, test = "Satterthwaite")
  
  # Full model
  m_full    <- lmer(logTE ~ treatment + puzzleId +
                      puzzleSkill + age + quantitativeExposure + energyLevel +
                      deviceType + tabSwitchTimePaused + (1 | userId),
                    data = data, REML = FALSE)
  vcov_full <- vcovCR(m_full, cluster = data$userId, type = "CR1")
  rt_full   <- coef_test(m_full, vcov = vcov_full, test = "Satterthwaite")
  
  trt_lean <- grepl("^treatment", rt_lean$Coef)
  trt_full <- grepl("^treatment", rt_full$Coef)
  
  bind_rows(
    data.frame(model = "Lean", term = rt_lean$Coef[trt_lean],
               est = round(rt_lean$beta[trt_lean], 3),
               se  = round(rt_lean$SE[trt_lean], 3),
               t   = round(rt_lean$tstat[trt_lean], 3),
               p   = round(rt_lean$p_Satt[trt_lean], 4)),
    data.frame(model = "Full", term = rt_full$Coef[trt_full],
               est = round(rt_full$beta[trt_full], 3),
               se  = round(rt_full$SE[trt_full], 3),
               t   = round(rt_full$tstat[trt_full], 3),
               p   = round(rt_full$p_Satt[trt_full], 4))
  )
}

bind_rows(
  run_primary(df)          %>% mutate(sample = "Full"),
  run_primary(df_prolific) %>% mutate(sample = "Prolific"),
  run_primary(df_organic)  %>% mutate(sample = "Organic")
) %>%
  select(sample, model, term, est, se, t, p) %>%
  arrange(term, model, sample) %>%
  mutate(p_adj = round(p.adjust(p, method = "BH"), 4)) %>%
  kable(caption = "Primary Model (logTE): Lean and Full by Sample (cluster-robust SEs)")
```

## Primary Model 2: Zero-Attempt Skip Rate

Zero-attempt skips (didSkip + attemptCount == 0) measure cold
disengagement — giving up without trying. This is the primary binary
outcome. Normal-approximation z-tests (adequate with n \> 200).

```{r zero-att-skip-model}
run_zas <- function(data) {
  m <- glmer(
    zero_att_skip ~ treatment + puzzleId + scale(puzzleSkill) + (1 | userId),
    data = data, family = binomial,
    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
  )
  cs   <- coef(summary(m))
  rows <- grep("^treatment", rownames(cs))
  data.frame(
    term = rownames(cs)[rows],
    OR   = round(exp(cs[rows, "Estimate"]), 3),
    z    = round(cs[rows, "z value"], 3),
    p    = round(cs[rows, "Pr(>|z|)"], 4)
  )
}

df_prolific <- df %>% filter(!is.na(prolificPid))
df_organic  <- df %>% filter(is.na(prolificPid))

zas_full     <- run_zas(df)          %>% mutate(sample = "Full")
zas_prolific <- run_zas(df_prolific) %>% mutate(sample = "Prolific")
zas_organic  <- run_zas(df_organic)  %>% mutate(sample = "Organic")

bind_rows(zas_prolific, zas_organic, zas_full) %>%
  select(sample, term, OR, z, p) %>%
  arrange(term, sample) %>%
  kable(caption = "Zero-Attempt Skip Model: OR by Sample (exploratory; raw p-values)")
```

## Consistency Check: Attempt Count Among Skippers

Among rounds that ended in a skip, did the label change how hard people
tried before giving up? Complements the zero-att skip finding:
easy_label → more attempts before quitting (not just fewer cold quits).
Outcome: log(attemptCount + 1). Models use cluster-robust SEs.

```{r att-skippers}
# All puzzles
df_skip <- df %>% filter(didSkip == TRUE, puzzleId != "puzzle_04")
cat("Skipped rounds (excl. puzzle_04):", nrow(df_skip), "from", n_distinct(df_skip$userId), "users\n")

m_att_all  <- lmer(logAtt ~ treatment + puzzleId + puzzleSkill + (1 | userId),
                   data = df_skip, REML = FALSE)
rv_att_all <- vcovCR(m_att_all, cluster = df_skip$userId, type = "CR1")
rt_att_all <- coef_test(m_att_all, vcov = rv_att_all, test = "Satterthwaite")
trt_a      <- grepl("treatment", rt_att_all$Coef)

data.frame(
  Estimate = round(rt_att_all$beta[trt_a], 4),
  SE       = round(rt_att_all$SE[trt_a], 4),
  t        = round(rt_att_all$tstat[trt_a], 3),
  p_Satt   = round(rt_att_all$p_Satt[trt_a], 4),
  row.names = rt_att_all$Coef[trt_a]
) %>% kable(caption = "log(attempts+1) among skippers — all hard/medium puzzles (exploratory)")

# Hard number sequences only (puzzle_01 + puzzle_05)
df_hard_skip <- df %>% filter(puzzleId %in% c("puzzle_01", "puzzle_05"), didSkip == TRUE)
cat("\nHard puzzle (p01+p05) skipped rounds:", nrow(df_hard_skip),
    "from", n_distinct(df_hard_skip$userId), "users\n")

m_att_hard  <- lmer(logAtt ~ treatment + puzzleId + puzzleSkill + (1 | userId),
                    data = df_hard_skip, REML = FALSE)
rv_att_hard <- vcovCR(m_att_hard, cluster = df_hard_skip$userId, type = "CR1")
rt_att_hard <- coef_test(m_att_hard, vcov = rv_att_hard, test = "Satterthwaite")
trt_h       <- grepl("treatment", rt_att_hard$Coef)

data.frame(
  Estimate = round(rt_att_hard$beta[trt_h], 4),
  SE       = round(rt_att_hard$SE[trt_h], 4),
  t        = round(rt_att_hard$tstat[trt_h], 3),
  p_Satt   = round(rt_att_hard$p_Satt[trt_h], 4),
  row.names = rt_att_hard$Coef[trt_h]
) %>% kable(caption = "log(attempts+1) among skippers — puzzle_01 + puzzle_05 only (exploratory)")

df_hard_skip %>%
  group_by(treatment, puzzleId) %>%
  summarize(n = n(), mean_att = round(mean(attemptCount), 2),
            pct_zero_att = round(mean(attemptCount == 0) * 100, 1), .groups = "drop") %>%
  arrange(puzzleId, treatment) %>%
  kable(caption = "Raw means — avg attempts before skip on hard puzzles")
```

## Robustness: Overall Skip Rate & Solve Rate

```{r robustness-models}
# Overall skip rate
m_skip <- suppressWarnings(glmer(
  didSkip ~ treatment + puzzleId + puzzleSkill + (1 | userId),
  data = df, family = binomial, control = glmerControl(optimizer = "bobyqa")
))
cs_skip <- coef(summary(m_skip))
cat("--- Overall skip (didSkip) ---\n")
for (r in c("treatmenteasy_label", "treatmenthard_label"))
  cat(sprintf("  %s: OR=%.3f  p=%.4f\n", r, exp(cs_skip[r, "Estimate"]),
              2 * pnorm(abs(cs_skip[r, "z value"]), lower.tail = FALSE)))

# Solve rate
m_solve <- suppressWarnings(glmer(
  isSolved ~ treatment + puzzleId + puzzleSkill + (1 | userId),
  data = df, family = binomial, control = glmerControl(optimizer = "bobyqa")
))
cs_solve <- coef(summary(m_solve))
cat("\n--- Solve rate (isSolved) ---\n")
for (r in c("treatmenteasy_label", "treatmenthard_label"))
  cat(sprintf("  %s: OR=%.3f  p=%.4f\n", r, exp(cs_solve[r, "Estimate"]),
              2 * pnorm(abs(cs_solve[r, "z value"]), lower.tail = FALSE)))

# logTE among solvers only
clean_solved <- df %>% filter(isSolved == TRUE)
m_solved <- lmer(logTE ~ treatment + puzzleId + puzzleSkill + (1 | userId),
                 data = clean_solved, REML = FALSE)
rv_sol <- vcovCR(m_solved, cluster = clean_solved$userId, type = "CR1")
rt_sol <- coef_test(m_solved, vcov = rv_sol, test = "Satterthwaite")
trt_s <- grepl("treatment", rt_sol$Coef)
cat("\n--- logTE among solved rounds only ---\n")
print(data.frame(
  Estimate = round(rt_sol$beta[trt_s], 4),
  SE       = round(rt_sol$SE[trt_s], 4),
  t        = round(rt_sol$tstat[trt_s], 3),
  p_Satt   = round(rt_sol$p_Satt[trt_s], 4),
  row.names = rt_sol$Coef[trt_s]
))
```


## Secondary Model (Persistence under failure)
```{r}

#Look at only the skipped puzzles, what was the time before skipping under the treatment groups compared to control. 

# Keep only puzzle rounds that ended in a skip
df_skip_time <- df %>%
  filter(didSkip == TRUE) %>%
  drop_na(timeEngaged, treatment, puzzleId, puzzleSkill, userId) %>%
  mutate(
    treatment = relevel(factor(treatment), ref = "no_label"),
    puzzleId  = relevel(factor(puzzleId), ref = "puzzle_01"),
    logTE      = log(timeEngaged + 1)
  )

cat(
  "Skipped rounds:", nrow(df_skip_time),
  "\nParticipants:", n_distinct(df_skip_time$userId),
  "\n"
)

# Raw descriptive results
df_skip_time %>%
  group_by(treatment) %>%
  summarize(
    Participants = n_distinct(userId),
    `Skipped Rounds` = n(),
    `Mean Time to Skip (s)` = round(mean(timeEngaged), 1),
    `Median Time to Skip (s)` = round(median(timeEngaged), 1),
    .groups = "drop"
  ) %>%
  knitr::kable(
    caption = "Time Engaged Among Skipped Puzzle Rounds"
  )

# Model time to skip
m_skip_time <- lmer(
  logTE ~ treatment + puzzleId + puzzleSkill + (1 | userId),
  data = df_skip_time,
  REML = FALSE
)

# Participant-clustered standard errors
vcov_skip_time <- vcovCR(
  m_skip_time,
  cluster = df_skip_time$userId,
  type = "CR1"
)

test_skip_time <- coef_test(
  m_skip_time,
  vcov = vcov_skip_time,
  test = "Satterthwaite"
)

# Display treatment effects only
treatment_rows <- grepl("^treatment", test_skip_time$Coef)

skip_time_results <- data.frame(
  Comparison = test_skip_time$Coef[treatment_rows],
  Estimate = test_skip_time$beta[treatment_rows],
  SE = test_skip_time$SE[treatment_rows],
  t = test_skip_time$tstat[treatment_rows],
  p = test_skip_time$p_Satt[treatment_rows]
) %>%
  mutate(
    `Approximate % Difference` =
      100 * (exp(Estimate) - 1),
    `BH-adjusted p` =
      p.adjust(p, method = "BH")
  ) %>%
  mutate(
    across(
      c(Estimate, SE, t, p, `Approximate % Difference`,
        `BH-adjusted p`),
      ~ round(.x, 4)
    )
  )

skip_time_results %>%
  knitr::kable(
    caption = "Label Effects on Time Engaged Among Skipped Puzzles"
  )
```

