---
title: "Puzzle Engagement Experiment — Data Pipeline & Analysis"
author: 
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preprocessing

Data preprocessing of .json file from Firestore.

```{r load-and-flatten}
# install.packages(c("jsonlite", "dplyr", "tidyr", "purrr", "lme4", "lmtest", "pwr"))

library(jsonlite)
library(dplyr)
library(tidyr)
library(purrr)
library(lme4)
library(clubSandwich)
library(lmtest)
library(pwr)

# Load raw JSON — simplifyVector = FALSE preserves irregular arrays as lists
raw_data <- fromJSON("firestore_export.json", simplifyVector = FALSE)

cat("Users:", length(raw_data$users), "\n")
cat("GameLogs:", length(raw_data$gameLogs), "\n")

# ── Flatten users ──────────────────────────────────────────────────────────────
users_df <- tibble(
  document_id = names(raw_data$users),
  data = raw_data$users
) %>%
  unnest_wider(data)

# List columns on users:
#   puzzleOrder   — array of puzzle id strings  → comma-separated string
#   resumeRounds  — array of round numbers (empty for single-session users) → comma-separated or NA
#   breakLog      — array of {timestamp, duration} maps → derive breaksTaken + totalBreakTime scalars
users_df <- users_df %>%
  mutate(
    puzzleOrder    = map_chr(puzzleOrder,
                             ~ paste(.x, collapse = ",")),
    resumeRounds   = map_chr(resumeRounds,
                             ~ if (length(.x) == 0) NA_character_ else paste(.x, collapse = ",")),
    breaksTaken    = map_int(breakLog,
                             ~ if (is.null(.x)) 0L else length(.x)),
    totalBreakTime = map_dbl(breakLog,
                             ~ if (is.null(.x)) 0 else sum(map_dbl(.x, "duration"), na.rm = TRUE))
  ) %>%
  select(-breakLog)

# ── Flatten gameLogs ───────────────────────────────────────────────────────────
gamelogs_df <- tibble(
  document_id = names(raw_data$gameLogs),
  data = raw_data$gameLogs
) %>%
  unnest_wider(data)

# rawSubmissionTimestamps is a variable-length list; drop it for the master join.
# Filter replays — isReplay == TRUE marks repeated attempts after completion.
# Exception: sharedDevice == "new_player" means a different person on the same household
# device; the app marks their gamelogs isReplay=TRUE but they are genuinely new participants.
new_player_ids <- users_df %>%
  filter(!is.na(sharedDevice) & sharedDevice == "new_player") %>%
  pull(userId)

gamelogs_df <- gamelogs_df %>%
  select(-rawSubmissionTimestamps) %>%
  filter(is.na(isReplay) | isReplay == FALSE | userId %in% new_player_ids)

# ── Master join ────────────────────────────────────────────────────────────────
# Drop treatment/isReplay from gameLogs (canonical copies live on users).
# sessionComplete is on users (true once all 5 puzzles done), not on gamelogs.
experiment_master <- gamelogs_df %>%
  select(-treatment, -isReplay) %>%
  left_join(users_df %>% arrange(userId, desc(sessionComplete)) %>% distinct(userId, .keep_all = TRUE), by = "userId") %>%
  select(-starts_with("document_id"))

glimpse(experiment_master)
```

## Covariate Balance Check

Check randomization frequencies across treatment arms.

```{r balance-check}
user_level <- experiment_master %>%
  distinct(userId, .keep_all = TRUE) %>%
  mutate(source = ifelse(!is.na(prolificPid), "Prolific", "Organic"))

# Treatment counts
user_level %>%
  count(treatment) %>%
  kable(col.names = c("Treatment", "N"), caption = "Treatment Assignment")

# Continuous covariates: mean by arm + ANOVA p
continuous_vars <- c("age", "quantitativeExposure")
cont_means <- user_level %>%
  group_by(treatment) %>%
  summarize(across(all_of(continuous_vars), ~ round(mean(.x, na.rm = TRUE), 2))) %>%
  pivot_longer(-treatment, names_to = "Covariate", values_to = "val") %>%
  pivot_wider(names_from = treatment, values_from = val)

cont_p <- map_dfr(continuous_vars, function(v) {
  p <- summary(aov(as.formula(paste(v, "~ treatment")), data = user_level))[[1]]["treatment", "Pr(>F)"]
  data.frame(Covariate = v, p_value = round(p, 4))
})

left_join(cont_means, cont_p, by = "Covariate") %>%
  kable(caption = "Continuous Covariate Balance (mean, ANOVA p)")

# Ordinal covariates: median by arm + Fisher's p
ordinal_vars <- c("puzzleSkill", "puzzleEnjoyment", "energyLevel")
ord_medians <- user_level %>%
  group_by(treatment) %>%
  summarize(across(all_of(ordinal_vars), ~ median(.x, na.rm = TRUE))) %>%
  pivot_longer(-treatment, names_to = "Covariate", values_to = "val") %>%
  pivot_wider(names_from = treatment, values_from = val)

ord_p <- map_dfr(ordinal_vars, function(v) {
  p <- fisher.test(table(user_level$treatment, user_level[[v]]),
                   simulate.p.value = TRUE)$p.value
  data.frame(Covariate = v, p_value = round(p, 4))
})

left_join(ord_medians, ord_p, by = "Covariate") %>%
  kable(caption = "Ordinal Covariate Balance (median, Fisher's exact p)")

# Categorical covariates: % per arm + chi-square / Fisher's p
user_level %>%
  group_by(treatment) %>%
  summarize(
    `% Prolific` = paste0(round(mean(source == "Prolific") * 100, 1), "%"),
    `% Desktop`  = paste0(round(mean(deviceType == "Desktop", na.rm = TRUE) * 100, 1), "%"),
    `% Female`   = paste0(round(mean(gender == "Female", na.rm = TRUE) * 100, 1), "%")
  ) %>%
  kable(caption = "Categorical Covariate Balance (% by arm)")

data.frame(
  Variable = c("source", "deviceType", "gender"),
  p_value  = c(
    round(fisher.test(table(user_level$treatment, user_level$source), 
                      simulate.p.value = TRUE)$p.value, 4),
    round(fisher.test(table(user_level$treatment, user_level$deviceType), 
                      simulate.p.value = TRUE)$p.value, 4),
    round(fisher.test(table(user_level$treatment, user_level$gender), 
                      simulate.p.value = TRUE)$p.value, 4)
  )
) %>%
  kable(caption = "Categorical Balance p-values (Fisher's exact)")
```

## Outcome Summary

```{r outcome-summary}
experiment_master %>%
  group_by(treatment) %>%
  summarize(
    `Rounds`          = n(),
    `Solve Rate`      = paste0(round(mean(isSolved, na.rm = TRUE) * 100, 1), "%"),
    `Avg Time (s)`    = round(mean(timeEngaged, na.rm = TRUE), 1),
    `Avg Attempts`    = round(mean(attemptCount, na.rm = TRUE), 2),
    `Skip Rate`       = paste0(round(mean(didSkip, na.rm = TRUE) * 100, 1), "%"),
    `Avg Rage Clicks` = round(mean(rageClicks, na.rm = TRUE), 2)
  ) %>%
  kable(caption = "Outcome Summary by Treatment")
```

## Data Cleaning

Session-level and round-level exclusions applied in order below.

```{r clean-data}
# ── Exclusion criteria ─────────────────────────────────────────────────────────
# sessionComplete is retained as a column — filter downstream (TRUE = completer).
# Session-level (user excluded entirely):
#   S2. age < 18: adults only
#   S3. sharedDevice == "returning": duplicate session from same device
#   S4. prolificRejected == TRUE: flagged by Prolific platform
#   S5. total tabSwitchTimePaused > 600s across session (verified away from browser)
#   S6. any round: didSkip + tabSwitchTimePaused == 0 + timeEngaged > per-puzzle p97
#       of skip distribution — walked away without switching tabs, came back and skipped
#   S7. 3+ zero-attempt skips with timeEngaged in 30–35s — skip button unlocks at 30s;
#       clicking immediately on 3+ puzzles signals waiting out the timer, not genuine engagement

# Per-puzzle p97 skip-time thresholds (computed from all skip rounds pre-cleaning)
skip_p97 <- experiment_master %>%
  filter(didSkip == TRUE) %>%
  group_by(puzzleId) %>%
  summarize(p97_skip = quantile(timeEngaged, 0.97), .groups = "drop")

# S6: users with at least one idle-walk-away skip round
idle_skip_users <- experiment_master %>%
  left_join(skip_p97, by = "puzzleId") %>%
  filter(didSkip == TRUE &
           (is.na(tabSwitchTimePaused) | tabSwitchTimePaused == 0) &
           timeEngaged > p97_skip) %>%
  pull(userId) %>%
  unique()

# S7: users with 5 zero-attempt skips at 30–35s (skip button unlocks at 30s; immediate click = timer-waiting)
cluster_skip_users <- experiment_master %>%
  filter(didSkip == TRUE & timeEngaged >= 30 & timeEngaged <= 35) %>%
  group_by(userId) %>%
  summarize(n_cluster_skips = n(), .groups = "drop") %>%
  filter(n_cluster_skips == 5) %>%
  pull(userId)

# Compute per-user total tab-switch pause time (S5)
user_tab_paused <- experiment_master %>%
  filter(isSolved | didSkip) %>%
  group_by(userId) %>%
  summarize(total_tab_paused = sum(tabSwitchTimePaused, na.rm = TRUE), .groups = "drop")

clean <- experiment_master %>%
  filter(age >= 18) %>%
  filter(is.na(sharedDevice) | sharedDevice == "new_player") %>%
  filter(is.na(prolificRejected) | prolificRejected != TRUE) %>%
  filter(!userId %in% idle_skip_users) %>%
  filter(!userId %in% cluster_skip_users) %>%
  filter(isSolved | didSkip) %>%
  left_join(user_tab_paused, by = "userId") %>%
  filter(total_tab_paused <= 600) %>%
  select(-total_tab_paused) %>%
  mutate(
    source     = ifelse(!is.na(prolificPid), "Prolific", "Organic"),
    logTE      = log(timeEngaged + 1),
    treatment  = relevel(factor(treatment), ref = "no_label"),
    puzzleId   = relevel(factor(puzzleId),  ref = "puzzle_01"),
    deviceType = factor(deviceType)
  )

cat("Idle-walk-away sessions excluded (S6):", length(idle_skip_users), "\n")
cat("Cluster-skip sessions excluded (S7):", length(cluster_skip_users), "\n")
cat("Clean rounds:", nrow(clean), "\n")
cat("Complete users:", n_distinct(clean$userId), "\n")

clean %>% count(treatment) %>% kable(col.names = c("Treatment", "Rounds"), caption = "Rounds per Treatment")

clean %>%
  group_by(puzzleId, treatment) %>%
  summarize(n = n(), skip_pct = paste0(round(mean(didSkip) * 100, 0), "%"), .groups = "drop") %>%
  pivot_wider(names_from = treatment, values_from = c(n, skip_pct)) %>%
  arrange(puzzleId) %>%
  kable(caption = "Skip Rate per Puzzle by Treatment")
```

## Exclusion Flags: AI Assistance Detection

Two-tier flag applied after deduplication on (userId, puzzleId,
roundNumber). Hard puzzles are puzzle_01 (number sequence), puzzle_02
(alphabet wheel), puzzle_05 (tile sequence). Thresholds are empirical
percentiles of solve-time distribution on first-attempt solves.

Rules: R1. Self-reported: honestyCheck == "assisted" R2a. Any single
hard puzzle solved in ≤ p5 time on first attempt (puzzle_01 ≤ 21s,
puzzle_02 ≤ 41s, puzzle_05 ≤ 23s) R2b. 2+ hard puzzles solved in ≤ p10
time on first attempt (puzzle_01 ≤ 30s, puzzle_02 ≤ 58s, puzzle_05 ≤
37s)

```{r exclusion-flags}
# Deduplicate: keep one row per user × puzzle × round
# Smarter dedup: prefer completed rows (isSolved | didSkip) over incomplete ones
dedup <- clean %>%
  arrange(userId, puzzleId, roundNumber, desc(isSolved | didSkip), didSkip) %>%
  distinct(userId, puzzleId, roundNumber, .keep_all = TRUE)

hard_puzzles <- c("puzzle_01", "puzzle_02", "puzzle_05")

# Solve-time percentile thresholds on hard puzzles (computed on clean data)
# p5  — flag any single hard puzzle solved below this time on first attempt (R2a)
# p10 — flag if 2+ hard puzzles below this time on first attempt (R2b)
p5_thresh  <- c(puzzle_01 = 21, puzzle_02 = 41, puzzle_05 = 23)
p10_thresh <- c(puzzle_01 = 30, puzzle_02 = 58, puzzle_05 = 37)

ai_flags <- dedup %>%
  mutate(
    below_p5  = puzzleId %in% hard_puzzles & isSolved & attemptCount == 1 &
                ((puzzleId == "puzzle_01" & timeEngaged <= p5_thresh["puzzle_01"]) |
                 (puzzleId == "puzzle_02" & timeEngaged <= p5_thresh["puzzle_02"]) |
                 (puzzleId == "puzzle_05" & timeEngaged <= p5_thresh["puzzle_05"])),
    below_p10 = puzzleId %in% hard_puzzles & isSolved & attemptCount == 1 &
                ((puzzleId == "puzzle_01" & timeEngaged <= p10_thresh["puzzle_01"]) |
                 (puzzleId == "puzzle_02" & timeEngaged <= p10_thresh["puzzle_02"]) |
                 (puzzleId == "puzzle_05" & timeEngaged <= p10_thresh["puzzle_05"]))
  ) %>%
  group_by(userId, prolificPid, treatment) %>%
  summarize(
    flag_R1  = any(honestyCheck == "assisted", na.rm = TRUE),  # self-reported
    flag_R2a = sum(below_p5,  na.rm = TRUE) >= 1,             # any p5 blitz
    flag_R2b = sum(below_p10, na.rm = TRUE) >= 2,             # 2+ p10 blitz
    ai_flagged = flag_R1 | flag_R2a | flag_R2b,
    .groups = "drop"
  )

# Manual overrides: participants reviewed and cleared despite automated flag
# 6977b9ea4ad70bb8a33c9b2d (local_lhde027og): p01 20s/a1 (1s below p5), p05 35s/a1 (2s below p10)
#   — margins too tight to be confident; struggled on p02 (118s/a3, tab switch); solo self-report
manual_clear <- c("local_lhde027og")

all_flags <- ai_flags %>%
  mutate(any_flag = ai_flagged & !userId %in% manual_clear)

cat("=== Exclusion summary ===\n")
cat("R1  Admitted assistance:         ", sum(ai_flags$flag_R1),  "\n")
cat("R2a Any hard puzzle below p5:    ", sum(ai_flags$flag_R2a), "\n")
cat("R2b 2+ hard puzzles below p10:   ", sum(ai_flags$flag_R2b), "\n")
cat("TOTAL excluded (R1|R2a|R2b):     ", sum(all_flags$any_flag), "\n")

# Master dataset: all users passing S2-S7 + AI exclusions
# One row per user × puzzle (earliest round kept). sessionComplete retained for downstream filtering.
master <- clean %>%
  left_join(all_flags %>% select(userId, any_flag), by = "userId") %>%
  filter(!any_flag) %>%
  mutate(
    zero_att_skip = as.integer(didSkip & attemptCount == 0),
    logAtt        = log(attemptCount + 1)
  ) %>%
  arrange(userId, puzzleId, roundNumber) %>%
  distinct(userId, puzzleId, .keep_all = TRUE)

cat("\nMaster dataset:", n_distinct(master$userId), "users |", nrow(master), "rounds\n")
cat("Completers (sessionComplete=TRUE):", n_distinct(master$userId[master$sessionComplete == TRUE]), "\n")
cat("Rounds per treatment:\n")
print(count(master %>% filter(sessionComplete == TRUE), treatment))

saveRDS(master, "analysis_data.rds")
cat("Saved analysis_data.rds\n")
```

## Attrition Analysis

```{r attrition, results = 'asis'}
user_level <- master %>% distinct(userId, .keep_all = TRUE)

# Attrition rates by source and treatment
user_level %>%
  group_by(source, treatment) %>%
  summarize(
    n_started     = n(),
    n_dropped     = sum(sessionComplete == FALSE),
    attrition_pct = paste0(round(mean(sessionComplete == FALSE) * 100, 1), "%"),
    .groups = "drop"
  ) %>%
  kable(caption = "Attrition by Treatment Arm and Recruitment Source")

# Logistic model: does treatment predict dropout?
user_level <- user_level %>%
  mutate(
    dropped   = as.integer(sessionComplete == FALSE),
    treatment = relevel(factor(treatment), ref = "no_label")
  )

m_att_full     <- glm(dropped ~ treatment, data = user_level, family = binomial)
m_att_organic  <- glm(dropped ~ treatment, data = filter(user_level, source == "Organic"),  family = binomial)
m_att_prolific <- glm(dropped ~ treatment, data = filter(user_level, source == "Prolific"), family = binomial)

stargazer(m_att_full, m_att_organic, m_att_prolific,
          type             = "text",
          column.labels    = c("Full", "Organic", "Prolific"),
          covariate.labels = c("Easy Label", "Hard Label"),
          dep.var.labels   = "Dropped Out",
          title            = "Differential Attrition: Effect of Treatment on Dropout",
          digits           = 3,
          star.cutoffs     = c(0.05, 0.01, 0.001),
          report           = "vcs*p")

# Completer vs dropout baseline comparison
user_level %>%
  group_by(Status = ifelse(sessionComplete == TRUE, "Completed", "Dropped")) %>%
  summarize(
    n                  = n(),
    Age                = round(mean(age, na.rm = TRUE), 1),
    `Puzzle Skill`     = median(puzzleSkill, na.rm = TRUE),
    `Puzzle Enjoyment` = median(puzzleEnjoyment, na.rm = TRUE),
    `Energy Level`     = median(energyLevel, na.rm = TRUE),
    `% Prolific`       = paste0(round(mean(source == "Prolific") * 100, 1), "%")
  ) %>%
  kable(caption = "Baseline Characteristics: Completers vs Dropouts")
```

------------------------------------------------------------------------

*Modeling and results are in `modeling_scratch.Rmd` and `report.Rmd`.
Load `analysis_data.rds` to access the final analysis dataset.*
