---
title: "Puzzle Engagement — Modeling Scratch"
date: "`r Sys.Date()`"
output: html_document
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-data}
library(dplyr)
library(tidyr)
library(purrr)
library(lme4)
library(clubSandwich)
library(knitr)
library(stargazer)

master      <- readRDS("analysis_data.rds")
df          <- master %>% filter(sessionComplete == TRUE)
df_prolific <- df %>% filter(source == "Prolific")
df_organic  <- df %>% filter(source == "Organic")

cat("Master:", n_distinct(master$userId), "users |", nrow(master), "rounds\n")
cat("Completers:", n_distinct(df$userId), "| Prolific:", n_distinct(df_prolific$userId),
    "| Organic:", n_distinct(df_organic$userId), "\n")
```

## Primary Model 1: log(timeEngaged)

Mixed effects model with random intercept per user (rounds nested within
users). Reference: treatment = no_label, puzzleId = puzzle_01.
Cluster-robust SEs (CR1, Satterthwaite df) account for within-user
correlation. Lean spec (treatment + puzzleId + puzzleSkill) is primary;
full covariate model is robustness check.

```{r primary-model}
run_primary <- function(data) {
  # Lean model
  m_lean    <- lmer(logTE ~ treatment + puzzleId + puzzleSkill + (1 | userId),
                    data = data, REML = FALSE)
  vcov_lean <- vcovCR(m_lean, cluster = data$userId, type = "CR1")
  rt_lean   <- coef_test(m_lean, vcov = vcov_lean, test = "Satterthwaite")

  # Full model
  m_full    <- lmer(logTE ~ treatment + puzzleId +
                      puzzleSkill + age + quantitativeExposure + energyLevel +
                      deviceType + tabSwitchTimePaused + (1 | userId),
                    data = data, REML = FALSE)
  vcov_full <- vcovCR(m_full, cluster = data$userId, type = "CR1")
  rt_full   <- coef_test(m_full, vcov = vcov_full, test = "Satterthwaite")

  trt_lean <- grepl("^treatment", rt_lean$Coef)
  trt_full <- grepl("^treatment", rt_full$Coef)

  bind_rows(
    data.frame(model = "Lean", term = rt_lean$Coef[trt_lean],
               est = round(rt_lean$beta[trt_lean], 3),
               se  = round(rt_lean$SE[trt_lean], 3),
               t   = round(rt_lean$tstat[trt_lean], 3),
               p   = round(rt_lean$p_Satt[trt_lean], 4)),
    data.frame(model = "Full", term = rt_full$Coef[trt_full],
               est = round(rt_full$beta[trt_full], 3),
               se  = round(rt_full$SE[trt_full], 3),
               t   = round(rt_full$tstat[trt_full], 3),
               p   = round(rt_full$p_Satt[trt_full], 4))
  )
}

bind_rows(
  run_primary(df)          %>% mutate(sample = "Full"),
  run_primary(df_prolific) %>% mutate(sample = "Prolific"),
  run_primary(df_organic)  %>% mutate(sample = "Organic")
) %>%
  select(sample, model, term, est, se, t, p) %>%
  arrange(term, model, sample) %>%
  mutate(p_adj = round(p.adjust(p, method = "BH"), 4)) %>%
  kable(caption = "Primary Model (logTE): Lean and Full by Sample (cluster-robust SEs, BH-adjusted)")
```

## Primary Model 2: Zero-Attempt Skip Rate

Zero-attempt skips (didSkip + attemptCount == 0) measure cold
disengagement — giving up without trying. Exploratory; raw p-values reported.

```{r zero-att-skip-model}
run_zas <- function(data) {
  m <- glmer(
    zero_att_skip ~ treatment + puzzleId + scale(puzzleSkill) + (1 | userId),
    data = data, family = binomial,
    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
  )
  cs   <- coef(summary(m))
  rows <- grep("^treatment", rownames(cs))
  data.frame(
    term = rownames(cs)[rows],
    OR   = round(exp(cs[rows, "Estimate"]), 3),
    z    = round(cs[rows, "z value"], 3),
    p    = round(cs[rows, "Pr(>|z|)"], 4)
  )
}

bind_rows(
  run_zas(df)          %>% mutate(sample = "Full"),
  run_zas(df_prolific) %>% mutate(sample = "Prolific"),
  run_zas(df_organic)  %>% mutate(sample = "Organic")
) %>%
  select(sample, term, OR, z, p) %>%
  arrange(term, sample) %>%
  kable(caption = "Zero-Attempt Skip Model: OR by Sample (exploratory; raw p-values)")
```

## Consistency Check: Attempt Count Among Skippers

Among rounds that ended in a skip, did the label change how hard people
tried before giving up? Outcome: log(attemptCount + 1). Models use cluster-robust SEs.

```{r att-skippers}
df_skip <- df %>% filter(didSkip == TRUE)
cat("Skipped rounds:", nrow(df_skip), "from", n_distinct(df_skip$userId), "users\n")

m_att_all  <- lmer(logAtt ~ treatment + puzzleId + puzzleSkill + (1 | userId),
                   data = df_skip, REML = FALSE)
rv_att_all <- vcovCR(m_att_all, cluster = df_skip$userId, type = "CR1")
rt_att_all <- coef_test(m_att_all, vcov = rv_att_all, test = "Satterthwaite")
trt_a      <- grepl("treatment", rt_att_all$Coef)

data.frame(
  Estimate  = round(rt_att_all$beta[trt_a], 4),
  SE        = round(rt_att_all$SE[trt_a], 4),
  t         = round(rt_att_all$tstat[trt_a], 3),
  p_Satt    = round(rt_att_all$p_Satt[trt_a], 4),
  row.names = rt_att_all$Coef[trt_a]
) %>% kable(caption = "log(attempts+1) among skippers — all puzzles (exploratory)")

# Hard number sequences only (puzzle_01 + puzzle_05)
df_hard_skip <- df %>% filter(puzzleId %in% c("puzzle_01", "puzzle_05"), didSkip == TRUE)
cat("\nHard puzzle (p01+p05) skipped rounds:", nrow(df_hard_skip),
    "from", n_distinct(df_hard_skip$userId), "users\n")

m_att_hard  <- lmer(logAtt ~ treatment + puzzleId + puzzleSkill + (1 | userId),
                    data = df_hard_skip, REML = FALSE)
rv_att_hard <- vcovCR(m_att_hard, cluster = df_hard_skip$userId, type = "CR1")
rt_att_hard <- coef_test(m_att_hard, vcov = rv_att_hard, test = "Satterthwaite")
trt_h       <- grepl("treatment", rt_att_hard$Coef)

data.frame(
  Estimate  = round(rt_att_hard$beta[trt_h], 4),
  SE        = round(rt_att_hard$SE[trt_h], 4),
  t         = round(rt_att_hard$tstat[trt_h], 3),
  p_Satt    = round(rt_att_hard$p_Satt[trt_h], 4),
  row.names = rt_att_hard$Coef[trt_h]
) %>% kable(caption = "log(attempts+1) among skippers — puzzle_01 + puzzle_05 only (exploratory)")

df_hard_skip %>%
  group_by(treatment, puzzleId) %>%
  summarize(n = n(), mean_att = round(mean(attemptCount), 2),
            pct_zero_att = round(mean(attemptCount == 0) * 100, 1), .groups = "drop") %>%
  arrange(puzzleId, treatment) %>%
  kable(caption = "Raw means — avg attempts before skip on hard puzzles")
```

## Robustness: Overall Skip Rate & Solve Rate

```{r robustness-models}
# Overall skip rate
m_skip <- suppressWarnings(glmer(
  didSkip ~ treatment + puzzleId + puzzleSkill + (1 | userId),
  data = df, family = binomial, control = glmerControl(optimizer = "bobyqa")
))
cs_skip <- coef(summary(m_skip))
data.frame(
  term = c("easy_label", "hard_label"),
  OR   = round(exp(cs_skip[c("treatmenteasy_label", "treatmenthard_label"), "Estimate"]), 3),
  p    = round(2 * pnorm(abs(cs_skip[c("treatmenteasy_label", "treatmenthard_label"), "z value"]),
                         lower.tail = FALSE), 4)
) %>% kable(caption = "Overall Skip Rate (didSkip) — robustness")

# Solve rate
m_solve <- suppressWarnings(glmer(
  isSolved ~ treatment + puzzleId + puzzleSkill + (1 | userId),
  data = df, family = binomial, control = glmerControl(optimizer = "bobyqa")
))
cs_solve <- coef(summary(m_solve))
data.frame(
  term = c("easy_label", "hard_label"),
  OR   = round(exp(cs_solve[c("treatmenteasy_label", "treatmenthard_label"), "Estimate"]), 3),
  p    = round(2 * pnorm(abs(cs_solve[c("treatmenteasy_label", "treatmenthard_label"), "z value"]),
                         lower.tail = FALSE), 4)
) %>% kable(caption = "Solve Rate (isSolved) — robustness")

# logTE among solvers only
clean_solved <- df %>% filter(isSolved == TRUE)
m_solved  <- lmer(logTE ~ treatment + puzzleId + puzzleSkill + (1 | userId),
                  data = clean_solved, REML = FALSE)
rv_sol    <- vcovCR(m_solved, cluster = clean_solved$userId, type = "CR1")
rt_sol    <- coef_test(m_solved, vcov = rv_sol, test = "Satterthwaite")
trt_s     <- grepl("treatment", rt_sol$Coef)

data.frame(
  Estimate  = round(rt_sol$beta[trt_s], 4),
  SE        = round(rt_sol$SE[trt_s], 4),
  t         = round(rt_sol$tstat[trt_s], 3),
  p_Satt    = round(rt_sol$p_Satt[trt_s], 4),
  row.names = rt_sol$Coef[trt_s]
) %>% kable(caption = "logTE among solved rounds only — robustness")
```



## Secondary Model: Attempt count
```{r}
run_primary <- function(data) {
  # Lean model
  m_lean    <- lmer(logAtt ~ treatment + puzzleId + puzzleSkill + (1 | userId),
                    data = data, REML = FALSE)
  vcov_lean <- vcovCR(m_lean, cluster = data$userId, type = "CR1")
  rt_lean   <- coef_test(m_lean, vcov = vcov_lean, test = "Satterthwaite")

  # Full model
  m_full    <- lmer(logAtt ~ treatment + puzzleId +
                      puzzleSkill + age + quantitativeExposure + energyLevel +
                      deviceType + tabSwitchTimePaused + (1 | userId),
                    data = data, REML = FALSE)
  vcov_full <- vcovCR(m_full, cluster = data$userId, type = "CR1")
  rt_full   <- coef_test(m_full, vcov = vcov_full, test = "Satterthwaite")

  trt_lean <- grepl("^treatment", rt_lean$Coef)
  trt_full <- grepl("^treatment", rt_full$Coef)

  bind_rows(
    data.frame(model = "Lean", term = rt_lean$Coef[trt_lean],
               est = round(rt_lean$beta[trt_lean], 3),
               se  = round(rt_lean$SE[trt_lean], 3),
               t   = round(rt_lean$tstat[trt_lean], 3),
               p   = round(rt_lean$p_Satt[trt_lean], 4)),
    data.frame(model = "Full", term = rt_full$Coef[trt_full],
               est = round(rt_full$beta[trt_full], 3),
               se  = round(rt_full$SE[trt_full], 3),
               t   = round(rt_full$tstat[trt_full], 3),
               p   = round(rt_full$p_Satt[trt_full], 4))
  )
}

bind_rows(
  run_primary(df)          %>% mutate(sample = "Full"),
  run_primary(df_prolific) %>% mutate(sample = "Prolific"),
  run_primary(df_organic)  %>% mutate(sample = "Organic")
) %>%
  select(sample, model, term, est, se, t, p) %>%
  arrange(term, model, sample) %>%
  mutate(p_adj = round(p.adjust(p, method = "BH"), 4)) %>%
  kable(caption = "Primary Model (logAtt): Lean and Full by Sample (cluster-robust SEs, BH-adjusted)")


```


