---
title: "The Effect of Difficulty Labels on Puzzle Engagement"
author: ""
date: "`r Sys.Date()`"
output: bookdown::pdf_document2
header-includes:
  - \usepackage{booktabs}
  - \usepackage[table]{xcolor}
  - \usepackage{threeparttable}
  - \usepackage{multirow}
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE, error = TRUE)
```

```{r load-data, include=FALSE}
library(dplyr)
library(tidyr)
library(purrr)
library(lme4)
library(clubSandwich)
library(knitr)
library(kableExtra)
library(modelsummary)

master      <- readRDS("analysis_data.rds")
df          <- master %>% filter(sessionComplete == TRUE)
df_prolific <- df %>% filter(source == "Prolific")
df_organic  <- df %>% filter(source == "Organic")
```

## Study Design

### Procedure

The experiment was delivered as a browser-based web application at
anotherpuzzle.io. The application was a single-page JavaScript game
engine that managed the complete participant session — from recruitment
intake through behavioral data collection — without any researcher
intervention at runtime. All data were logged directly to a cloud
database (Firestore) at both the session and round levels.

Upon arriving at the application, participants were presented with a
consent screen. Participants who agreed were immediately and invisibly
assigned to one of three treatment arms before proceeding. Following
consent, participants completed an eight-item survey collecting baseline
covariates: age, gender, education level, quantitative exposure, current
energy level, puzzle frequency, self-rated puzzle skill, and puzzle
enjoyment. After the survey, participants were shown a brief instruction
screen and then clicked "Start the Challenge," which triggered
randomization of puzzle order and treatment assignment and the creation
of the participant's session record.

The game engine logged behavioral data at the round level. For each
round, it recorded: total time engaged (a continuous timer that paused
automatically during tab switches and participant-initiated breaks),
number of answer attempts, whether the first attempt was correct, time
to first attempt, tab-switch count and total time paused, rage-click
events, and whether the round ended in a solve or a skip. Skip was
enabled after 30 seconds of engagement, preventing immediate abandonment
without any attempt. At the session level, the engine recorded device
type, puzzle presentation order, recruitment source, and session
completion status.

### Treatment

The experimental treatment was a difficulty label presented to
participants before each logic puzzle, designed to prime expectations
about puzzle difficulty prior to engagement. The anticipated mechanism
was that these primed expectations would influence persistence, measured
as time spent on a puzzle before solving or skipping.

Participants were randomly assigned at session entry to one of three
conditions: (1) no label, (2) easy label, or (3) hard label. The
assigned label was shown before every puzzle in the session, so a
participant assigned to the hard label condition saw that label at the
start of each of the five rounds. Within the application, all other
aspects of the experience were held constant across arms. The interface,
instructions, and the five logic puzzles were identical across
conditions, so differences in behavior across treatment groups can be
attributed to the difficulty label rather than any other feature of the
experience.

Each session consisted of five logic puzzles presented in randomized
order: a number sequence, an alphabet wheel, a shape sequence, a hybrid
shape-number puzzle, and a tile sequence.

```{r puzzle-skip-rate}
overall_rates <- df %>%
  mutate(
    Puzzle = case_when(
      puzzleId == "puzzle_01" ~ "Number Sequence",
      puzzleId == "puzzle_02" ~ "Alphabet Wheel",
      puzzleId == "puzzle_03" ~ "Shape Sequence",
      puzzleId == "puzzle_04" ~ "Hybrid Puzzle",
      puzzleId == "puzzle_05" ~ "Tile Sequence"
    )
  ) %>%
  group_by(Puzzle) %>%
  summarize(overall_raw = mean(didSkip, na.rm = TRUE), .groups = "drop")

df %>%
  mutate(
    Puzzle = case_when(
      puzzleId == "puzzle_01" ~ "Number Sequence",
      puzzleId == "puzzle_02" ~ "Alphabet Wheel",
      puzzleId == "puzzle_03" ~ "Shape Sequence",
      puzzleId == "puzzle_04" ~ "Hybrid Puzzle",
      puzzleId == "puzzle_05" ~ "Tile Sequence"
    )
  ) %>%
  group_by(Puzzle, treatment) %>%
  summarize(skip_rate = mean(didSkip, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = treatment, values_from = skip_rate) %>%
  left_join(overall_rates, by = "Puzzle") %>%
  arrange(desc(overall_raw)) %>%
  mutate(across(c(no_label, easy_label, hard_label, overall_raw),
                ~ paste0(round(. * 100, 1), "%"))) %>%
  select(Puzzle, no_label, easy_label, hard_label, overall_raw) %>%
  rename(
    `No Label`   = no_label,
    `Easy Label` = easy_label,
    `Hard Label` = hard_label,
    `Overall`    = overall_raw
  ) %>%
  kable(booktabs = TRUE,
        caption  = "Skip Rate by Puzzle and Treatment Arm") %>%
  kable_styling(latex_options = c("hold_position"))
```

All five puzzles were designed to be intentionally challenging, though
they varied in how difficult participants found them in practice. Skip
and solve rates differed meaningfully across puzzle types, as shown in
Table \@ref(tab:puzzle-skip-rate). Puzzle order was randomized across
participants to prevent any one puzzle from systematically falling at
the same point in the session, and the analysis accounts for
puzzle-level differences and within-participant correlation across
rounds.

Treatment delivery was controlled entirely through the application, and
the data contain no indication that any participant was shown a label
other than their assigned condition; we therefore identified no
treatment non-compliers. However, the application did not independently
log the randomized assignment alongside the label rendered on screen for
each round, so perfect delivery could not be independently verified.
Future experiments should log these separately to enable a formal
compliance check.

## Potential Outcomes

Each participant has three potential outcomes — the time engaged they
would produce under each of the three label conditions:
$Y_i(\text{no label})$, $Y_i(\text{easy label})$, and
$Y_i(\text{hard label})$. By the fundamental problem of causal
inference, only one of these is ever observed for any given participant.
Random assignment to one of the three arms makes the unobserved
potential outcomes independent of treatment assignment, allowing average
potential outcomes to be estimated from group means.

The primary comparisons of interest are the average treatment effects of
each label condition relative to the `no label` control:

$$\tau_{\text{easy}} = E[Y_i(\text{easy label})] - E[Y_i(\text{no label})]$$

$$\tau_{\text{hard}} = E[Y_i(\text{hard label})] - E[Y_i(\text{no label})]$$

The `no label` arm serves as the reference group, representing
engagement in the absence of any difficulty prime. The two primary
comparisons — `easy label` versus `no label` and `hard label` versus
`no label` — isolate the effect of priming participants with a
difficulty expectation, holding everything else constant. A positive
effect would indicate that priming increases persistence (time engaged);
a negative effect would indicate it discourages it. The direction is
expected to differ by label valence, with an easy prime potentially
encouraging engagement and a hard prime potentially suppressing it.

## Randomization and Attrition

### Randomization

Treatment was assigned entirely client-side at the moment a participant
clicked "I understand – let's go" on the consent screen. The game engine
drew a single uniform random variate and mapped it to one of three arms
with nearly equal probability:

```         
r = Math.random()
Treatment = r < 0.333 ? 'no_label'
          : r < 0.666 ? 'hard_label'
          : 'easy_label'
```

`Math.random()` draws a number between 0 and 1 uniformly, with the unit
interval cut into three equal slices. Each arm was assigned with
approximately one-third probability. The treatment draw was written to a
`localStorage` draft immediately after assignment, which occurs after
the consent screen. Participants who closed and returned mid-session
were shown a "Welcome back" screen that restored their original
treatment from the draft, ensuring the draw was not re-rolled on page
reload. A participant's IP address was additionally captured and stored
under a separate `localStorage` key to detect whether a device had
previously completed a full session. Returning devices were routed to a
shared-device screening question where participants self-reported
whether they were a new or returning player. Together, the IP-based flag
and the self-report allowed us to distinguish true first-time sessions
from replays and from new participants sharing a household device,
retaining only genuine first sessions.

### Recruitment Populations

This study drew from two distinct recruitment populations. Because the
recruitment source shapes both the attrition mechanism and potentially
the outcome measures themselves, we identify each participant's source
(organic or Prolific) and account for it throughout the analysis. The
organic sample was recruited through personal networks – friends,
family, and word of mouth – with no standardized referral message, no
controlled timing, and no single delivery channel. Organic participants
had no financial incentive to complete the study. The Prolific sample
was recruited through the Prolific platform, which delivers a
standardized study listing and compensates participants contingent on
completion and quality of the effort. Prolific participants are likely
to be experienced survey-takers operating under platform norms around
task completion.

### Attrition

Attrition is defined as beginning treatment, but failing to complete all
five rounds. Specifically, a participant is counted as an attritor if
they have a user record in the database – created at the moment they
click "Start the Challenge" and are shown a Round 1 label card, but
their session was never marked complete. Variable `sessionComplete` is
set to TRUE only when a participant reaches the results screen after
resolving all five rounds (each ending in either a correct answer or a
skip). Participants who passed through the survey and opted out before
clicking "Start the Challenge" are not counted as attritors, because no
user record is created for them and they were never exposed to the
treatment.

#### Differential Attrition in Organic and Prolific subsamples

The attrition analysis fits logistic regression models predicting
dropout on treatment arm, estimated separately for the full sample and
the organic subsample, with no_label as the reference arm. The Prolific
subsample had zero dropout across all arms and is reported descriptively
only.

```{r tab-attrition}
user_level <- master %>%
  distinct(userId, .keep_all = TRUE) %>%
  mutate(
    dropped   = as.integer(sessionComplete == FALSE),
    treatment = relevel(factor(treatment), ref = "no_label")
  )

m_att_full    <- glm(dropped ~ treatment, data = user_level, family = binomial)
m_att_organic <- glm(dropped ~ treatment,
                     data = filter(user_level, source == "Organic"),
                     family = binomial)

get_or_p <- function(model, term) {
  if (is.null(model)) return(c(NA, NA))
  s <- coef(summary(model))
  if (!term %in% rownames(s)) return(c(NA, NA))
  c(exp(s[term, "Estimate"]), s[term, "Pr(>|z|)"])
}

fmt_p <- function(p) {
  if (is.na(p)) return("—")
  ifelse(p < .001, "<.001", as.character(round(p, 3)))
}

or_lookup <- bind_rows(
  tibble(source = "Organic", treatment = "easy_label",
         OR = get_or_p(m_att_organic, "treatmenteasy_label")[1],
         p  = get_or_p(m_att_organic, "treatmenteasy_label")[2]),
  tibble(source = "Organic", treatment = "hard_label",
         OR = get_or_p(m_att_organic, "treatmenthard_label")[1],
         p  = get_or_p(m_att_organic, "treatmenthard_label")[2]),
  tibble(source = "Prolific", treatment = "easy_label", OR = NA, p = NA),
  tibble(source = "Prolific", treatment = "hard_label", OR = NA, p = NA),
  tibble(source = "Full", treatment = "easy_label",
         OR = get_or_p(m_att_full, "treatmenteasy_label")[1],
         p  = get_or_p(m_att_full, "treatmenteasy_label")[2]),
  tibble(source = "Full", treatment = "hard_label",
         OR = get_or_p(m_att_full, "treatmenthard_label")[1],
         p  = get_or_p(m_att_full, "treatmenthard_label")[2])
)

attrition_tbl <- user_level %>%
  group_by(source, treatment) %>%
  summarize(
    n_started     = n(),
    n_dropped     = sum(dropped),
    attrition_pct = paste0(round(mean(dropped) * 100, 1), "%"),
    .groups = "drop"
  )

get_pct <- function(src, trt) {
  attrition_tbl %>% filter(source == src, treatment == trt) %>% pull(attrition_pct)
}

org_no_pct   <- get_pct("Organic", "no_label")
org_easy_pct <- get_pct("Organic", "easy_label")
org_hard_pct <- get_pct("Organic", "hard_label")
org_easy_or  <- round(get_or_p(m_att_organic, "treatmenteasy_label")[1], 3)
org_easy_p   <- round(get_or_p(m_att_organic, "treatmenteasy_label")[2], 3)
org_hard_or  <- round(get_or_p(m_att_organic, "treatmenthard_label")[1], 3)
org_hard_p   <- round(get_or_p(m_att_organic, "treatmenthard_label")[2], 3)

attrition_tbl %>%
  mutate(treatment = factor(treatment,
                            levels = c("no_label", "easy_label", "hard_label"),
                            labels = c("No Label", "Easy Label", "Hard Label"))) %>%
  left_join(or_lookup %>%
              mutate(treatment = factor(treatment,
                                       levels = c("no_label", "easy_label", "hard_label"),
                                       labels = c("No Label", "Easy Label", "Hard Label"))),
            by = c("source", "treatment")) %>%
  mutate(
    OR = case_when(
      treatment == "No Label" ~ "ref.",
      is.na(OR)              ~ "—",
      TRUE                   ~ as.character(round(OR, 3))
    ),
    p = case_when(
      treatment == "No Label" ~ "—",
      is.na(p)               ~ "—",
      TRUE                   ~ sapply(p, fmt_p)
    )
  ) %>%
  arrange(source, treatment) %>%
  select(-source) %>%
  rename(
    Treatment      = treatment,
    `N Started`    = n_started,
    `N Dropped`    = n_dropped,
    `Dropout Rate` = attrition_pct,
    `Odds Ratio`   = OR,
    `p-value`      = p
  ) %>%
  kable(caption = "Attrition by Treatment Arm and Recruitment Source",
        booktabs = TRUE,
        align    = c("l", "c", "c", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = FALSE) %>%
  pack_rows("Organic",  1, 3) %>%
  pack_rows("Prolific", 4, 6) %>%
  footnote(general = "Odds ratios from logistic regression predicting dropout; reference arm: No Label. Prolific model omitted due to zero attrition. Full sample OR uses pooled model.",
           threeparttable = TRUE)
```

Among organic participants, dropout rates were `r org_no_pct` in the no
label arm, `r org_easy_pct` in the easy label arm, and `r org_hard_pct`
in the hard label arm. The easy label coefficient is negligible and
non-significant (OR = `r org_easy_or`, p = `r org_easy_p`), providing no
evidence of a difference from the control. The hard label arm shows a
suggestive but non-significant reduction in dropout odds relative to no
label (OR = `r org_hard_or`, p = `r org_hard_p`), contrary to the prior
expectation that a hard label would discourage persistence. The
difference in dropout rates across arms is not statistically
significant, so we cannot conclude that differential attrition
meaningfully biases the primary outcome estimates. If the pattern were
real, the direction of bias would work against finding a treatment
effect. The no label arm losing more participants than the hard label
arm would leave the no label completers as a comparatively more select
group, which would compress the estimated difference between arms. The
Prolific subsample had zero dropout across all three arms and is
reported descriptively only, consistent with payment-based compliance
overriding any label effect on attrition (see Table
\@ref(tab:tab-attrition)).

## Intent-to-Treat & Covariate Balance

Given the attrition observed above, the primary analysis proceeds on an
intent-to-treat basis restricted to participants who completed all five
rounds (`sessionComplete == TRUE`). This approach preserves the
integrity of the original randomization, where treatment effects are
estimated among those who received and complied with their assigned
condition and avoids the additional assumptions required to impute
outcomes for participants who dropped out.

```{r tab-balance}
bal <- df %>%
  distinct(userId, .keep_all = TRUE) %>%
  mutate(
    treatment   = relevel(factor(treatment), ref = "no_label"),
    is_prolific = as.integer(source == "Prolific"),
    is_desktop  = as.integer(deviceType == "Desktop"),
    is_female   = as.integer(gender == "Female")
  )

models <- list(
  "Age"            = lm(age                  ~ treatment, data = bal),
  "Quant Exp"      = lm(quantitativeExposure ~ treatment, data = bal),
  "Puzzle Skill"   = lm(puzzleSkill          ~ treatment, data = bal),
  "Puzzle Enjoy"   = lm(puzzleEnjoyment      ~ treatment, data = bal),
  "Energy"         = lm(energyLevel          ~ treatment, data = bal),
  "Recruit Source" = lm(is_prolific          ~ treatment, data = bal),
  "Desktop"        = lm(is_desktop           ~ treatment, data = bal),
  "Female"         = lm(is_female            ~ treatment, data = bal)
)

modelsummary(
  models,
  coef_map  = c("treatmenteasy_label" = "Easy Label",
                "treatmenthard_label" = "Hard Label"),
  stars     = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  gof_omit  = ".*",
  title     = "Covariate Balance Across Treatment Arms",
  notes     = "OLS coefficients vs. no-label reference arm. SEs in parentheses.",
  booktabs  = TRUE,
  output    = "kableExtra"
) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

Table @ref(tab:tab-balance) presents covariate balance across the three
treatment arms among completers. Each column reports an OLS regression
of one baseline characteristic on treatment indicators, with `no label`
as the reference arm. None of the `easy label` or `hard label`
coefficients reach conventional significance thresholds on any covariate
— age, quantitative exposure, puzzle skill, puzzle enjoyment, energy
level, recruitment source, device type, or gender. This indicates that
the randomization produced comparable groups on all measured
pre-treatment characteristics. The R² values are near zero throughout,
confirming that treatment assignment explains essentially none of the
variance in any baseline variable. This balance supports a clean
intent-to-treat interpretation: observed differences in outcomes across
arms are unlikely to reflect pre-existing differences in participant
composition.

## Discussion

[Interpretation of results, attrition limitations, and conclusions]
