---
title: "The Effect of Difficulty Labels on Puzzle Engagement"
author: ""
date: "`r Sys.Date()`"
output: bookdown::pdf_document2
bibliography: references.bib
header-includes:
  - \usepackage{booktabs}
  - \usepackage[table]{xcolor}
  - \usepackage{threeparttable}
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{graphicx}
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE, error = TRUE)
```

```{r load-data, include=FALSE}
library(MASS)
library(dplyr)
library(tidyr)
library(purrr)
library(lme4)
library(sandwich)
library(clubSandwich)
library(knitr)
library(kableExtra)
library(modelsummary)
library(jsonlite)
library(consort)
library(ggplot2)
library(patchwork)
library(stargazer)
library(lmtest)
library(margins)
master      <- readRDS("analysis_data.rds")
df          <- master %>% filter(sessionComplete == TRUE)
df_prolific <- df %>% filter(source == "Prolific")
df_organic  <- df %>% filter(source == "Organic")

theme_set(theme_bw(base_size = 11))
```

## Introduction

How difficult a task feels can shape how much effort a person puts into
it. This experiment asked whether showing participants a difficulty
label before a logic puzzle — telling them it is easy or hard — changes
how long they stay engaged with it. Participants were given the same set
of hard puzzles and randomly assigned to one of three conditions: no
label, a hard label (accurate), or an easy label (understated). We
compared mean time engaged across the three groups to test whether the
label changed behavior. Prior work on perceived task difficulty suggests
that the information people have about a task, along with their
expectations going in, influences how much effort they invest. By
varying the label shown before each puzzle, we manipulated those
expectations and observed the effect on engagement.

### Theory

Priming is when exposure to one stimulus influences how you respond to
the next. Research on semantic priming has shown that people recognize
related words faster when preceded by an associated cue
[@meyer1971facilitation], suggesting that prior context shapes how we
process what comes next.

@kahneman2011thinking extends this beyond words: he argues that System 1
— the fast, intuitive mode of thinking — is continuously shaped by
whatever context is available at the moment of judgment. A difficulty
label shown before a task is exactly this kind of cue. Before a
participant even reads the first clue, System 1 has already formed an
expectation about how hard the task will be and how much effort it
deserves. By the time deliberate reasoning engages, the frame is already
set.

When that frame is wrong — when someone primed with "easy" encounters a
genuinely hard puzzle — the mismatch creates cognitive friction.
@baumeister1998ego showed that sustained cognitive effort depletes a
limited self-regulatory resource, lowering persistence over time. A
violated expectation compounds this: participants are not just working
against the puzzle, they are also working against their own prior
expectation of how this should feel. That additional load, we argue,
lowers the threshold for giving up.

### Hypothesis

We hypothesize that the two label conditions affect time engaged in
opposite directions relative to the no label control. Participants shown
the hard label receive accurate information about task difficulty and
may mentally prepare for sustained effort, increasing persistence.
Participants shown the easy label encounter a mismatch between their
expectation and their experience: the puzzle is harder than primed,
which we expect to produce frustration that shortens engagement. When
participants expect a task to be easy and immediately find it is not,
the experience may feel like failure rather than challenge, lowering the
threshold for giving up and skipping. Formally:

$$H_1: \mu_{\text{hard}} > \mu_{\text{no label}}$$
$$H_2: \mu_{\text{easy}} < \mu_{\text{no label}}$$
$$H_3: \mu_{\text{hard}} > \mu_{\text{easy}}$$

where $\mu$ denotes mean time engaged per puzzle round.

## Power Calculation

To determine the necessary sample size, we conducted a simulation-based
power analysis rather than relying on a closed-form analytic formula.
This approach allows us to model the multi-round panel structure of the
experiment and account for participant-level clustering.

We set a control baseline average solve time of $\mu = 270$ seconds.
Variance was decomposed into a between-person component ($\sigma_u = 60$
seconds) and a within-person residual ($\sigma_e = 67$ seconds),
yielding a total standard deviation of
$\sigma = \sqrt{\sigma_u^2 + \sigma_e^2} \approx 90$ seconds, chosen
conservatively to capture natural individual variation in performance.
We specified treatment effects relative to the no label control in
opposite directions for the two label arms: a $+25$ second increase for
the hard label arm, reflecting anticipated persistence when participants
are accurately primed for difficulty, and a $-25$ second decrease for
the easy label arm, reflecting earlier disengagement when participants
encounter a task harder than the label suggested.

We simulated 500 iterations at escalating sample sizes. For each
iteration, a mixed-effects model was fit with puzzle fixed effects and a
participant-level random intercept:

$$Y_{it} = \beta_0 + \boldsymbol{\beta}_{\text{trt}} \text{Treatment}_i + \boldsymbol{\gamma}_t + u_i + \varepsilon_{it}$$

where $\boldsymbol{\gamma}_t$ are puzzle fixed effects and
$u_i \sim N(0, \sigma_u^2)$ is a participant random intercept. Because
data are generated exactly from the model, standard mixed-model
$t$-tests are used to assess significance; cluster-robust standard
errors are applied in the actual analysis but are not needed for
simulation where the variance structure is known.

```{r power-sim, message=FALSE, warning=FALSE}
if (file.exists("power_sim_results.rds")) {
  power_res <- readRDS("power_sim_results.rds")
} else {
  set.seed(241)
  mu_base   <- 270; beta_hard <- 25; beta_easy <- -25
  sigma_u   <-  60; sigma_e   <- 67; n_puzzles  <-  5; n_sim <- 500
  puzzle_fx <- c(-15, 10, 20, -8, -7)

  sim_power <- function(n_per_arm) {
    sig_hard <- sig_easy <- logical(n_sim)
    for (s in seq_len(n_sim)) {
      n_total  <- n_per_arm * 3
      trt_id   <- rep(c("no_label","easy_label","hard_label"), each = n_per_arm)
      u_i      <- rnorm(n_total, 0, sigma_u)
      userId   <- rep(seq_len(n_total), each = n_puzzles)
      puzzleId <- rep(seq_len(n_puzzles), times = n_total)
      trt_long <- rep(trt_id, each = n_puzzles)
      beta_trt <- ifelse(trt_long == "hard_label", beta_hard,
                  ifelse(trt_long == "easy_label", beta_easy, 0))
      Y <- mu_base + beta_trt + puzzle_fx[puzzleId] + u_i[userId] +
           rnorm(n_total * n_puzzles, 0, sigma_e)
      dat <- data.frame(
        Y         = Y,
        treatment = factor(trt_long, levels = c("no_label","easy_label","hard_label")),
        puzzleId  = factor(puzzleId),
        userId    = factor(userId)
      )
      m <- tryCatch(lmer(Y ~ treatment + puzzleId + (1|userId),
                         data = dat, REML = FALSE), error = function(e) NULL)
      if (is.null(m)) next
      ct <- summary(m)$coefficients
      th <- ct["treatmenthard_label", "t value"]
      te <- ct["treatmenteasy_label", "t value"]
      if (length(th)) sig_hard[s] <- abs(th) > 1.96
      if (length(te)) sig_easy[s] <- abs(te) > 1.96
    }
    data.frame(n_per_arm = n_per_arm,
               power_hard = mean(sig_hard),
               power_easy = mean(sig_easy))
  }

  power_res <- bind_rows(lapply(seq(40, 150, by = 10), sim_power))
  saveRDS(power_res, "power_sim_results.rds")
}

thresh_hard <- power_res$n_per_arm[which(power_res$power_hard >= 0.80)[1]]
thresh_easy <- power_res$n_per_arm[which(power_res$power_easy >= 0.80)[1]]
thresh_both <- max(thresh_hard, thresh_easy, na.rm = TRUE)
```

At $\alpha = 0.05$, power to detect the hard label effect crosses the
80% threshold at `r thresh_hard` participants per arm
(`r thresh_hard * 3` total), and the easy label effect crosses 80% at
`r thresh_easy` per arm (`r thresh_easy * 3` total). The
harder-to-detect effect requires `r thresh_both` per arm
(`r thresh_both * 3` total).

The final analytic sample of 256 completers falls below the
pre-specified target of `r thresh_both * 3`. Exclusions for behavioral
flags and attrition reduced the sample beyond what was anticipated at
the design stage, and arm sizes are unequal (No Label: 87, Easy Label:
80, Hard Label: 89), which modestly reduces power relative to equal
allocation. Beyond sample size, the analytic sample pools two
recruitment populations that differ systematically in engagement
conditions: Prolific participants completed the experiment in a
structured, incentivized setting with low distraction, while organic
participants, recruited through personal networks, varied more in
motivation, setting, and distraction level. This heterogeneity inflates
within-group variance beyond what a single $\sigma$ captures, further
eroding effective power. Both the sample shortfall and the variance
inflation from pooling two behaviorally distinct populations should be
kept in mind when interpreting null or marginal findings.

## Study Design

### Procedure

The experiment was delivered as a browser-based web application at
anotherpuzzle.io. The application was a single-page JavaScript game
engine that managed the complete participant session from recruitment
intake through behavioral data collection without any researcher
intervention at runtime. All data were logged directly to a cloud
database (Firestore) at both the session and round levels.

Upon arriving at the application, participants were presented with a
consent screen. Participants who agreed were immediately and invisibly
assigned to one of three treatment arms before proceeding. Following
consent, participants completed an eight-item survey collecting baseline
covariates: age, gender, education level, quantitative exposure, current
energy level, puzzle frequency, self-rated puzzle skill, and puzzle
enjoyment. After the survey, participants were shown a brief instruction
screen and then clicked "Start the Challenge," which triggered
randomization of puzzle order and the creation of the participant's
session record.

The game engine logged behavioral data at the round level. For each
round, it recorded: total time engaged (a continuous timer that paused
automatically during tab switches and participant-initiated breaks),
number of answer attempts, whether the first attempt was correct, time
to first attempt, tab-switch count and total time paused, rage-click
events, and whether the round ended in a solve or a skip. Skip was
enabled after 30 seconds of engagement, preventing immediate abandonment
without any attempt. At the session level, the engine recorded device
type, puzzle presentation order, recruitment source, and session
completion status.

### Treatment

The experimental treatment was a difficulty label presented to
participants before each logic puzzle. The hard label was expected to
prompt mental preparation for a demanding task, increasing persistence.
The easy label was expected to create an expectation mismatch:
participants anticipating an easy task would encounter greater
difficulty than expected, producing frustration that shortens engagement
before participants choose to skip.

Participants were randomly assigned at session entry to one of three
conditions: (1) no label, (2) easy label, or (3) hard label. The
assigned label was shown before every puzzle in the session, so a
participant assigned to the hard label condition saw that label at the
start of each of the five rounds. Within the application, all other
aspects of the experience were held constant across arms. The interface,
instructions, and the five logic puzzles were identical across
conditions, so differences in behavior across treatment groups can be
attributed to the difficulty label rather than any other feature of the
experience.

Each session consisted of five logic puzzles presented in randomized
order: a number sequence, an alphabet wheel, a shape sequence, a hybrid
shape-number puzzle, and a tile sequence. All five were designed to be
intentionally challenging, though they varied in how difficult
participants found them in practice.

```{r puzzle-skip-rate}
overall_rates <- df %>%
  mutate(
    Puzzle = case_when(
      puzzleId == "puzzle_01" ~ "Number Sequence",
      puzzleId == "puzzle_02" ~ "Alphabet Wheel",
      puzzleId == "puzzle_03" ~ "Shape Sequence",
      puzzleId == "puzzle_04" ~ "Hybrid Puzzle",
      puzzleId == "puzzle_05" ~ "Tile Sequence"
    )
  ) %>%
  group_by(Puzzle) %>%
  summarize(overall_raw = mean(didSkip, na.rm = TRUE), .groups = "drop")

df %>%
  mutate(
    Puzzle = case_when(
      puzzleId == "puzzle_01" ~ "Number Sequence",
      puzzleId == "puzzle_02" ~ "Alphabet Wheel",
      puzzleId == "puzzle_03" ~ "Shape Sequence",
      puzzleId == "puzzle_04" ~ "Hybrid Puzzle",
      puzzleId == "puzzle_05" ~ "Tile Sequence"
    )
  ) %>%
  group_by(Puzzle, treatment) %>%
  summarize(skip_rate = mean(didSkip, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = treatment, values_from = skip_rate) %>%
  left_join(overall_rates, by = "Puzzle") %>%
  arrange(desc(overall_raw)) %>%
  mutate(across(c(no_label, easy_label, hard_label, overall_raw),
                ~ paste0(round(. * 100, 1), "%"))) %>%
  dplyr::select(Puzzle, no_label, easy_label, hard_label, overall_raw) %>%
  rename(
    `No Label`   = no_label,
    `Easy Label` = easy_label,
    `Hard Label` = hard_label,
    `Overall`    = overall_raw
  ) %>%
  kable(booktabs = TRUE,
        caption  = "Skip Rate by Puzzle and Treatment Arm") %>%
  kable_styling(latex_options = c("hold_position"), font_size = 10)
```

Skip and solve rates differed meaningfully across puzzle types, as shown
in Table \@ref(tab:puzzle-skip-rate). Puzzle order was randomized across
participants to prevent any one puzzle from systematically falling at
the same point in the session, and the analysis accounts for
puzzle-level differences and within-participant correlation across
rounds.

Treatment delivery was controlled entirely through the application, and
the data contain no indication that any participant was shown a label
other than their assigned condition; we therefore identified no
treatment non-compliers. However, the application did not independently
log the randomized assignment alongside the label rendered on screen for
each round, so perfect delivery could not be independently verified.
Future experiments should log these separately to enable a formal
compliance check.

## Participants

The experiment database contained 398 unique participants. Eight were
excluded as returning or replay players whose devices had previously
completed a full session, leaving 390 first-time participants. An
additional 27 had no retained gamelog records and were excluded,
yielding 363 participants who started at least one puzzle round.

Twenty-one participants were excluded for eligibility: 8 were under 18
and 13 were rejected through Prolific's quality review. Of the remaining
342 eligible participants, 39 were removed for behavioral or
data-quality reasons:

*Behavioral exclusions*

1.  Total tab-switch pause time exceeded 600 seconds across the session,
    indicating the participant was away from the browser for an extended
    period.
2.  At least one puzzle showed an idle walk-away pattern: the
    participant skipped without switching tabs and with time engaged
    above the 97th percentile of the skip-time distribution for that
    puzzle, suggesting they left the page and returned.
3.  All five puzzles were skipped within 30 to 35 seconds with no answer
    attempts, indicating the participant waited out the skip-unlock
    timer on every round rather than engaging with the puzzles.

*AI assistance exclusions*

4.  Self-reported use of AI assistance in the post-session honesty
    check.
5.  Hard puzzles solved on the first attempt faster than population
    benchmarks for that puzzle (any single puzzle below the 5th
    percentile, or two or more below the 10th), subject to manual
    review. Flagged participants were retained if their fast solve times
    were consistent with their relative performance across all rounds,
    as genuine skill tends to produce proportionally fast times across
    puzzles rather than selectively fast times on isolated ones.

Attrition, defined as beginning treatment but failing to complete all
five rounds, accounted for 47 participants. The final analytic sample is
256 completers across three arms: No Label (n = 87), Easy Label (n =
80), and Hard Label (n = 89). Figure \@ref(fig:fig-consort) shows the
full participant flow.

```{r fig-consort, fig.width=9, fig.height=7, fig.cap="Participant flow through the experiment.", fig.pos="H", out.width="\\textwidth"}

raw_json  <- fromJSON("firestore_export.json", simplifyVector = FALSE)
users_all <- tibble(document_id = names(raw_json$users),
                    data        = raw_json$users) %>%
  unnest_wider(data) %>%
  distinct(userId, .keep_all = TRUE)

new_player_ids <- users_all %>%
  filter(!is.na(sharedDevice) & sharedDevice == "new_player") %>%
  pull(userId)

em_u          <- readRDS("experiment_master.rds") %>% distinct(userId, .keep_all = TRUE)
master_u      <- master %>% distinct(userId, .keep_all = TRUE)
completed_ids <- master_u %>% filter(sessionComplete == TRUE) %>% pull(userId)
dropped_ids   <- master_u %>% filter(sessionComplete == FALSE) %>% pull(userId)

arm_src <- master %>%
  filter(sessionComplete == TRUE) %>%
  distinct(userId, .keep_all = TRUE) %>%
  count(treatment, source)

gn <- function(trt, src) {
  v <- arm_src %>% filter(treatment == trt, source == src) %>% pull(n)
  if (length(v) == 0) 0L else v
}
n_no_org   <- gn("no_label",   "Organic")
n_no_pro   <- gn("no_label",   "Prolific")
n_easy_org <- gn("easy_label", "Organic")
n_easy_pro <- gn("easy_label", "Prolific")
n_hard_org <- gn("hard_label", "Organic")
n_hard_pro <- gn("hard_label", "Prolific")

flow <- users_all %>%
  mutate(
    exc_replay = case_when(
      !is.na(sharedDevice) & sharedDevice == "returning" ~ "Returning player",
      !is.na(isReplay) & isReplay == TRUE & !(userId %in% new_player_ids) ~ "Replay session",
      TRUE ~ NA_character_
    ),
    after_replay = if_else(is.na(exc_replay), userId, NA_character_),
    exc_no_game = case_when(
      !is.na(exc_replay) ~ NA_character_,
      !(userId %in% em_u$userId) ~ "No gamelog records",
      TRUE ~ NA_character_
    ),
    started = if_else(
      is.na(exc_replay) & (userId %in% em_u$userId), userId, NA_character_
    ),
    exc_elig = case_when(
      !is.na(exc_replay) | !(userId %in% em_u$userId) ~ NA_character_,
      !is.na(age) & age < 18 ~ "Age < 18",
      !is.na(prolificRejected) & prolificRejected == TRUE ~ "Prolific rejection",
      TRUE ~ NA_character_
    ),
    eligible = if_else(
      is.na(exc_replay) & (userId %in% em_u$userId) & is.na(exc_elig),
      userId, NA_character_
    ),
    exc_behav = case_when(
      !is.na(exc_replay) | !(userId %in% em_u$userId) | !is.na(exc_elig) ~ NA_character_,
      !(userId %in% master_u$userId) ~ "Behavioral or AI flag",
      TRUE ~ NA_character_
    ),
    quality = if_else(
      is.na(exc_replay) & (userId %in% em_u$userId) &
        is.na(exc_elig) & (userId %in% master_u$userId),
      userId, NA_character_
    ),
    exc_att = case_when(
      !is.na(exc_replay) | !(userId %in% em_u$userId) |
        !is.na(exc_elig) | !(userId %in% master_u$userId) ~ NA_character_,
      userId %in% dropped_ids ~ "Did not complete all 5 rounds",
      TRUE ~ NA_character_
    ),
    arm = case_when(
      !(userId %in% completed_ids) ~ NA_character_,
      treatment == "no_label" ~
        paste0("No Label\nOrganic: ", n_no_org, "\nProlific: ", n_no_pro, "\n"),
      treatment == "easy_label" ~
        paste0("Easy Label\nOrganic: ", n_easy_org, "\nProlific: ", n_easy_pro, "\n"),
      treatment == "hard_label" ~
        paste0("Hard Label\nOrganic: ", n_hard_org, "\nProlific: ", n_hard_pro, "\n")
    )
  )

p <- consort_plot(
  data   = flow,
  orders = c(
    userId       = 'Users with a session record',
    exc_replay   = 'Excluded',
    after_replay = 'Non-replay participants',
    exc_no_game  = 'Excluded',
    started      = 'Participants who started',
    exc_elig     = 'Excluded',
    eligible     = 'Eligible participants',
    exc_behav    = 'Excluded',
    quality      = 'Passed quality checks',
    exc_att      = 'Did not complete all 5 rounds',
    arm          = 'Analyzed'
  ),
  side_box   = c('exc_replay', 'exc_no_game', 'exc_elig', 'exc_behav', 'exc_att'),
  allocation = 'arm',
  cex        = 0.7
)

plot(p)
```

## Potential Outcomes

Each participant has three potential outcomes — how long they would
spend on the puzzles under each label condition: $Y_i(\text{no label})$,
$Y_i(\text{easy label})$, and $Y_i(\text{hard label})$. We only ever
observe one of these per person. Because participants were randomly
assigned, the treatment someone received is independent of what their
outcome would have been under a different condition, which means we can
estimate average outcomes for each arm using the observed group means.

The main comparisons we care about are the treatment effects of each
label relative to the no-label control, and the direct comparison
between the two label arms. Each $\tau$ below is the difference in group
means ($\mu$) between two arms:

$$\tau_{\text{easy}} = \mu_{\text{easy}} - \mu_{\text{no label}} = E[Y_i(\text{easy label})] - E[Y_i(\text{no label})]$$

$$\tau_{\text{hard}} = \mu_{\text{hard}} - \mu_{\text{no label}} = E[Y_i(\text{hard label})] - E[Y_i(\text{no label})]$$

$$\tau_{\text{hard vs. easy}} = \mu_{\text{hard}} - \mu_{\text{easy}} = E[Y_i(\text{hard label})] - E[Y_i(\text{easy label})]$$

The no-label arm is the reference group — participants who saw no
framing at all. The first two comparisons ask whether giving someone any
difficulty expectation changes their engagement relative to having none.
The third asks whether the direction of the label matters: we expect the
easy and hard labels to push behavior in opposite directions, so
$\tau_{\text{easy}} < 0$ (frustration leads to less engagement),
$\tau_{\text{hard}} > 0$ (preparation leads to more), and
$\tau_{\text{hard vs. easy}} > 0$.

## Randomization and Attrition

### Randomization

Treatment was assigned entirely client-side at the moment a participant
clicked "I understand -- let's go" on the consent screen. The game
engine drew a single uniform random variate and mapped it to one of
three arms with nearly equal probability:

```         
r = Math.random()
Treatment = r < 0.333 ? 'no_label'
          : r < 0.666 ? 'hard_label'
          : 'easy_label'
```

`Math.random()` draws a number between 0 and 1 uniformly, with the unit
interval cut into three equal slices. Each arm was assigned with
approximately one-third probability. The treatment draw was written to a
`localStorage` draft immediately after assignment. Participants who
closed and returned mid-session were shown a "Welcome back" screen that
restored their original treatment from the draft, ensuring the draw was
not re-rolled on page reload. A participant's IP address was
additionally captured and stored under a separate `localStorage` key to
detect whether a device had previously completed a full session.
Returning devices were routed to a shared-device screening question
where participants self-reported whether they were a new or returning
player. Together, the IP-based flag and the self-report allowed us to
distinguish true first-time sessions from replays and from new
participants sharing a household device, retaining only genuine first
sessions.

### Recruitment Populations

We recruited participants from two different sources, and we track this
throughout the analysis because recruitment context likely affects both
dropout and how people engage with the task. Organic participants were
recruited informally through personal networks — friends, family, and
word of mouth — with no financial incentive to complete the study.
Prolific participants were recruited through the Prolific platform,
which pays participants for completing studies. Prolific workers tend to
be experienced study participants who are used to following task
instructions carefully and completing what they start.

### Attrition

Attrition is defined as beginning treatment but failing to complete all
five rounds. A participant is counted as an attritor if they have a user
record in the database, created at the moment they click "Start the
Challenge" and are shown a Round 1 label card, but their session was
never marked complete. The variable `sessionComplete` is set to TRUE
only when a participant reaches the results screen after resolving all
five rounds, each ending in either a correct answer or a skip.
Participants who passed through the survey and opted out before clicking
"Start the Challenge" are not counted as attritors because no user
record is created for them and they were never exposed to the treatment.

#### Differential Attrition in Organic and Prolific Subsamples

The attrition analysis fits logistic regression models predicting
dropout on treatment arm, estimated separately for the full sample and
the organic subsample, with no label as the reference arm. The Prolific
subsample had zero dropout across all arms and is reported descriptively
only.

```{r tab-attrition}
user_level <- master %>%
  distinct(userId, .keep_all = TRUE) %>%
  mutate(
    dropped   = as.integer(sessionComplete == FALSE),
    treatment = relevel(factor(treatment), ref = "no_label")
  )

m_att_full    <- glm(dropped ~ treatment, data = user_level, family = binomial)
m_att_organic <- glm(dropped ~ treatment,
                     data = filter(user_level, source == "Organic"),
                     family = binomial)

get_or_p <- function(model, term) {
  if (is.null(model)) return(c(NA, NA))
  s <- coef(summary(model))
  if (!term %in% rownames(s)) return(c(NA, NA))
  c(exp(s[term, "Estimate"]), s[term, "Pr(>|z|)"])
}

fmt_p <- function(p) {
  if (is.na(p)) return("---")
  ifelse(p < .001, "<.001", as.character(round(p, 3)))
}

or_lookup <- bind_rows(
  tibble(source = "Organic", treatment = "easy_label",
         OR = get_or_p(m_att_organic, "treatmenteasy_label")[1],
         p  = get_or_p(m_att_organic, "treatmenteasy_label")[2]),
  tibble(source = "Organic", treatment = "hard_label",
         OR = get_or_p(m_att_organic, "treatmenthard_label")[1],
         p  = get_or_p(m_att_organic, "treatmenthard_label")[2]),
  tibble(source = "Prolific", treatment = "easy_label", OR = NA, p = NA),
  tibble(source = "Prolific", treatment = "hard_label", OR = NA, p = NA),
  tibble(source = "Full", treatment = "easy_label",
         OR = get_or_p(m_att_full, "treatmenteasy_label")[1],
         p  = get_or_p(m_att_full, "treatmenteasy_label")[2]),
  tibble(source = "Full", treatment = "hard_label",
         OR = get_or_p(m_att_full, "treatmenthard_label")[1],
         p  = get_or_p(m_att_full, "treatmenthard_label")[2])
)

attrition_tbl <- user_level %>%
  group_by(source, treatment) %>%
  summarize(
    n_started     = n(),
    n_dropped     = sum(dropped),
    attrition_pct = paste0(round(mean(dropped) * 100, 1), "%"),
    .groups = "drop"
  )

get_pct <- function(src, trt) {
  attrition_tbl %>% filter(source == src, treatment == trt) %>% pull(attrition_pct)
}

org_no_pct   <- get_pct("Organic", "no_label")
org_easy_pct <- get_pct("Organic", "easy_label")
org_hard_pct <- get_pct("Organic", "hard_label")
org_easy_or  <- round(get_or_p(m_att_organic, "treatmenteasy_label")[1], 3)
org_easy_p   <- round(get_or_p(m_att_organic, "treatmenteasy_label")[2], 3)
org_hard_or  <- round(get_or_p(m_att_organic, "treatmenthard_label")[1], 3)
org_hard_p   <- round(get_or_p(m_att_organic, "treatmenthard_label")[2], 3)

attrition_tbl %>%
  mutate(treatment = factor(treatment,
                            levels = c("no_label", "easy_label", "hard_label"),
                            labels = c("No Label", "Easy Label", "Hard Label"))) %>%
  left_join(or_lookup %>%
              mutate(treatment = factor(treatment,
                                       levels = c("no_label", "easy_label", "hard_label"),
                                       labels = c("No Label", "Easy Label", "Hard Label"))),
            by = c("source", "treatment")) %>%
  mutate(
    OR = case_when(
      treatment == "No Label" ~ "ref.",
      is.na(OR)              ~ "---",
      TRUE                   ~ as.character(round(OR, 3))
    ),
    p = case_when(
      treatment == "No Label" ~ "---",
      is.na(p)               ~ "---",
      TRUE                   ~ sapply(p, fmt_p)
    )
  ) %>%
  arrange(source, treatment) %>%
  dplyr::select(-source) %>%
  rename(
    Treatment      = treatment,
    `N Started`    = n_started,
    `N Dropped`    = n_dropped,
    `Dropout Rate` = attrition_pct,
    `Odds Ratio`   = OR,
    `p-value`      = p
  ) %>%
  kable(caption = "Attrition by Treatment Arm and Recruitment Source",
        booktabs = TRUE,
        align    = c("l", "c", "c", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = FALSE, font_size = 10) %>%
  pack_rows("Organic",  1, 3) %>%
  pack_rows("Prolific", 4, 6) %>%
  footnote(general = "Odds ratios from logistic regression predicting dropout; reference arm: No Label. Prolific model omitted due to zero attrition. Full sample OR uses pooled model.",
           threeparttable = TRUE)
```

Among organic participants, dropout rates were `r org_no_pct` in the no
label arm, `r org_easy_pct` in the easy label arm, and `r org_hard_pct`
in the hard label arm. The easy label showed no difference from the
control (OR = `r org_easy_or`, p = `r org_easy_p`). The hard label arm
had a suggestive but non-significant reduction in dropout relative to no
label (OR = `r org_hard_or`, p = `r org_hard_p`); notably, the primary
model found the hard label went in the opposite direction from what we
expected on engagement time. Dropout rates did not differ significantly
across arms, so we did not find evidence that differential attrition
meaningfully biased the primary outcome estimates. If the pattern were
real, the no label arm losing more participants would leave its
completers as a more selected group, which would compress the estimated
differences between arms. The Prolific subsample had zero dropout across
all three arms, which was expected given that Prolific participants are
paid to complete studies (see Table \@ref(tab:tab-attrition)).

## Completer Analysis & Covariate Balance

The primary analysis estimates average treatment effects among the 256
participants who completed all five rounds. Restricting to completers
introduces potential attrition bias if dropout is related to treatment.
As noted above, organic dropout was suggestively lower in the hard label
arm, though not statistically significant. If that pattern reflects a
real effect, the no label arm would lose a disproportionately higher
share of participants, leaving its completers as a more select group and
compressing estimated treatment effect differences toward zero.

```{r tab-balance}
bal <- df %>%
  distinct(userId, .keep_all = TRUE) %>%
  mutate(
    treatment   = relevel(factor(treatment), ref = "no_label"),
    is_prolific = as.integer(source == "Prolific"),
    is_desktop  = as.integer(deviceType == "Desktop"),
    is_female   = as.integer(gender == "Female")
  )

models <- list(
  "Age"            = lm(age                  ~ treatment, data = bal),
  "Quant Exp"      = lm(quantitativeExposure ~ treatment, data = bal),
  "Puzzle Skill"   = lm(puzzleSkill          ~ treatment, data = bal),
  "Puzzle Enjoy"   = lm(puzzleEnjoyment      ~ treatment, data = bal),
  "Energy"         = lm(energyLevel          ~ treatment, data = bal),
  "Recruit Source" = lm(is_prolific          ~ treatment, data = bal),
  "Desktop"        = lm(is_desktop           ~ treatment, data = bal),
  "Female"         = lm(is_female            ~ treatment, data = bal)
)

modelsummary(
  models,
  coef_map  = c("treatmenteasy_label" = "Easy Label",
                "treatmenthard_label" = "Hard Label"),
  stars     = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  gof_omit  = ".*",
  title     = "Covariate Balance Across Treatment Arms",
  notes     = "OLS coefficients vs. no-label reference arm. SEs in parentheses.",
  booktabs  = TRUE,
  output    = "kableExtra"
) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

Table \@ref(tab:tab-balance) presents covariate balance across the three
treatment arms among completers. Each column reports an OLS regression
of one baseline characteristic on treatment indicators, with no label as
the reference arm. None of the easy label or hard label coefficients
reach conventional significance thresholds on any covariate: age,
quantitative exposure, puzzle skill, puzzle enjoyment, energy level,
recruitment source, device type, or gender. The $R^2$ values are near
zero throughout, confirming that treatment assignment explains
essentially none of the variance in any baseline variable. This balance
supports the validity of the completer analysis: observed differences in
outcomes across arms are unlikely to reflect pre-existing differences in
participant composition.

## Results

### Behavioral Differences by Recruitment Source

To assess whether recruitment source introduced systematic differences
in data quality or engagement patterns, we examined four behavioral
indicators across Prolific and organic participants: across-person
variability in log time engaged, within-person variability in log time
engaged, tab switch frequency, and total time spent with the browser tab
inactive.

```{r tab-within-sd, results="asis", message=FALSE, warning=FALSE}
person_engage <- df %>%
  group_by(userId, source) %>%
  summarize(mean_logTE   = mean(logTE, na.rm = TRUE),
            sd_logTE     = sd(logTE, na.rm = TRUE),
            total_tabs   = sum(tabSwitchCount, na.rm = TRUE),
            total_paused = sum(tabSwitchTimePaused, na.rm = TRUE),
            .groups = "drop") %>%
  group_by(source) %>%
  mutate(across_sd = abs(mean_logTE - mean(mean_logTE))) %>%
  ungroup()

m_across    <- lm(across_sd    ~ source, data = person_engage)
m_within_sd <- lm(sd_logTE     ~ source, data = person_engage)
m_tabs      <- lm(total_tabs   ~ source, data = person_engage)
m_paused    <- lm(total_paused ~ source, data = person_engage)

stargazer(m_across, m_within_sd, m_tabs, m_paused,
          type             = "latex",
          title            = "Behavioral Differences by Recruitment Source",
          dep.var.labels   = c("Across-Person SD", "Within-Person SD", "Tab Switches", "Time Paused (s)"),
          covariate.labels = "Prolific (vs. Organic)",
          omit.stat        = c("f", "ser"),
          header           = FALSE,
          table.placement  = "H")
```

Across the four indicators, two differences emerge. Across-person
variability in log time engaged does not differ significantly between
groups, suggesting that the two subsamples are comparable in their
overall engagement levels. Within-person variability, however, is
significantly higher among organic participants: organic participants
were more erratic across their five puzzle rounds than Prolific
participants. One plausible explanation is that organic participants,
recruited informally through personal networks, varied more in their
environment, distraction level, and motivation from one puzzle to the
next. A participant might have been fully focused on one round and
interrupted on another.

Organic participants also spent significantly more total time with the
browser tab inactive, which points to more real-world interruptions
during the session. Tab switch counts did not differ significantly
between groups. Prolific participants likely kept more consistent
conditions throughout, given the platform's completion norms. Either
way, higher within-person variability among organic participants
increases the residual variance in the model, which means wider
confidence intervals and less power to detect a treatment effect. So the
null result doesn't necessarily mean there's no effect — it may just
mean we didn't have enough precision to detect a small one.

### Primary Model: Effect of Difficulty Labels on Log Time Engaged

#### Outcome Distribution

Raw time engaged is strongly right-skewed, with a long tail from
participants who spent disproportionately long on a single puzzle. Log
transformation is applied to reduce skew and stabilize variance before
modeling.

```{r fig-log-justification, fig.width=8, fig.height=3.5, fig.cap="Distribution of time engaged before and after log transformation.", fig.pos="H", out.width="\\textwidth"}
m1_raw <- ggplot(df, aes(x = timeEngaged)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  labs(x = "Time Engaged (seconds)", y = "Count", title = "Raw")

m1_log <- ggplot(df, aes(x = logTE)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  labs(x = "Log Time Engaged", y = "Count", title = "Log-Transformed")

m1_raw + m1_log
```

As shown in Figure \@ref(fig:fig-log-justification), the log
transformation substantially reduces skew and produces an approximately
symmetric distribution.

#### Model Specification

We estimated the effect of difficulty labels on log time engaged using a
linear mixed-effects model with puzzle fixed effects, a
participant-level random intercept, and pre-treatment covariates:

$$\log(\text{TimeEngaged}_{it}) = \beta_0 + \beta_1 \text{EasyLabel}_i + \beta_2 \text{HardLabel}_i + \mathbf{X}i'\boldsymbol{\delta} + u_i + \varepsilon_{it}$$

where $u_i \sim N(0, \sigma_u^2)$ is the participant-level random
intercept and $\mathbf{X}_i$ includes puzzle skill, recruitment source,
age, quantitative exposure, energy level, and device type. Standard
errors are clustered at the participant level using CR1 with
Satterthwaite degrees of freedom.

```{r primary-model, message=FALSE, warning=FALSE}
# Full sample (controls for source)
m      <- lmer(logTE ~ treatment + puzzleId + puzzleSkill + source +
                 age + quantitativeExposure + energyLevel + deviceType + (1 | userId),
               data = df, REML = FALSE)
vcov_m <- vcovCR(m, cluster = df$userId, type = "CR1")
rt     <- coef_test(m, vcov = vcov_m, test = "Satterthwaite")

# Organic subsample
m_org    <- lmer(logTE ~ treatment + puzzleId + puzzleSkill +
                   age + quantitativeExposure + energyLevel + deviceType + (1 | userId),
                 data = df_organic, REML = FALSE)
vcov_org <- vcovCR(m_org, cluster = df_organic$userId, type = "CR1")
rt_org   <- coef_test(m_org, vcov = vcov_org, test = "Satterthwaite")

# Prolific subsample
m_pro    <- lmer(logTE ~ treatment + puzzleId + puzzleSkill +
                   age + quantitativeExposure + energyLevel + deviceType + (1 | userId),
                 data = df_prolific, REML = FALSE)
vcov_pro <- vcovCR(m_pro, cluster = df_prolific$userId, type = "CR1")
rt_pro   <- coef_test(m_pro, vcov = vcov_pro, test = "Satterthwaite")

modelsummary(
  list("Full" = m, "Organic" = m_org, "Prolific" = m_pro),
  vcov      = list(vcov_m, vcov_org, vcov_pro),
  statistic = c("({std.error})", "p = {p.value}"),
  coef_map  = c("treatmenteasy_label" = "Easy Label",
                "treatmenthard_label" = "Hard Label"),
  gof_omit  = ".*",
  stars     = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  title     = "Effect of Difficulty Labels on Log Time Engaged",
  booktabs  = TRUE,
  output    = "kableExtra"
) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9) %>%
  footnote(general = "Mixed-effects model (lmer); cluster-robust SEs (CR1) by participant.",
           threeparttable = TRUE)
```

```{r m1-vals, include=FALSE}
m1_easy_b  <- round(rt$beta[rt$Coef == "treatmenteasy_label"], 3)
m1_easy_se <- round(rt$SE[rt$Coef == "treatmenteasy_label"], 3)
m1_easy_p  <- round(rt$p_Satt[rt$Coef == "treatmenteasy_label"], 3)
m1_hard_b  <- round(rt$beta[rt$Coef == "treatmenthard_label"], 3)
m1_hard_se <- round(rt$SE[rt$Coef == "treatmenthard_label"], 3)
m1_hard_p  <- round(rt$p_Satt[rt$Coef == "treatmenthard_label"], 3)
m1_hard_pct   <- round((1 - exp(rt$beta[rt$Coef == "treatmenthard_label"])) * 100, 1)
m1_hard_org_b <- round(rt_org$beta[rt_org$Coef == "treatmenthard_label"], 3)
m1_hard_pro_b <- round(rt_pro$beta[rt_pro$Coef == "treatmenthard_label"], 3)
```

Neither label had a statistically significant effect on log time engaged
relative to the no-label control in the full sample (Table
\@ref(tab:primary-model)). The easy label estimate is near zero overall
($\hat{\beta}_{\text{easy}}$ = `r m1_easy_b`, SE = `r m1_easy_se`, p =
`r m1_easy_p`), but this masks opposing effects across subgroups:
organic participants show a negative estimate consistent with H2, while
Prolific participants show a slight positive — the two cancel out in the
pooled sample. The hard label is consistently negative across both
subgroups (organic: `r m1_hard_org_b`, Prolific: `r m1_hard_pro_b`), and
the pooled estimate ($\hat{\beta}_{\text{hard}}$ = `r m1_hard_b`, SE =
`r m1_hard_se`, p = `r m1_hard_p`) is marginally significant at the 10%
level, corresponding to about `r m1_hard_pct`% less time engaged — the
opposite direction from H1, which predicted the hard label would
increase persistence.

This marginal result should be interpreted cautiously for two reasons.
First, across the primary and secondary models we ran multiple
comparisons — two treatment coefficients across three samples and two
outcomes — and we did not pre-register a correction procedure. Applying
a Bonferroni adjustment to the family of tests run here would push the
threshold well below the conventional 0.05 level, leaving no result
significant. Second, our power analysis targeted 120 participants per
arm to achieve 80% power, but we ended up with roughly 85 per arm among
completers. At that sample size the simulation suggests power closer to
65%, meaning we likely lack the precision to reliably detect effects of
the size we hypothesized, let alone smaller ones.

```{r fig-coef, fig.width=5, fig.height=2.5, fig.pos="H", fig.align="center", out.width="0.65\\textwidth", fig.cap="Point estimates and 95\\% confidence intervals for the effect of difficulty labels on log time engaged relative to the no label control. Error bars are based on cluster-robust standard errors."}
coef_df <- data.frame(
  term = c("Easy Label", "Hard Label"),
  est  = c(rt$beta[rt$Coef == "treatmenteasy_label"],
           rt$beta[rt$Coef == "treatmenthard_label"]),
  se   = c(rt$SE[rt$Coef == "treatmenteasy_label"],
           rt$SE[rt$Coef == "treatmenthard_label"])
) %>%
  mutate(lo = est - 1.96 * se, hi = est + 1.96 * se)

ggplot(coef_df, aes(x = est, y = term)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(aes(xmin = lo, xmax = hi), width = 0.15) +
  geom_point(size = 3) +
  labs(x = "Estimated Effect on Log Time Engaged", y = NULL,
       title    = "Effect of Difficulty Labels on Engagement",
       subtitle = "Relative to no-label control; error bars are 95% CIs") +
  theme_bw(base_size = 9)
```

As shown in Figure \@ref(fig:fig-coef), the confidence intervals for
both estimates are wide and cross zero, so we can't say much about the
true direction of either effect. H3 is also not supported — the hard
label point estimate is actually below the easy label estimate, the
opposite of what we predicted.

### Secondary Model: Effect of Difficulty Labels on Zero-Attempt Skips

```{r tab-skip-rate}
skip_by_puzzle <- df %>%
  mutate(puzzle = case_when(
    puzzleId == "puzzle_01" ~ "Number Sequence",
    puzzleId == "puzzle_02" ~ "Alphabet Wheel",
    puzzleId == "puzzle_03" ~ "Shape Sequence",
    puzzleId == "puzzle_04" ~ "Hybrid Puzzle",
    puzzleId == "puzzle_05" ~ "Tile Sequence"
  )) %>%
  group_by(puzzle, treatment) %>%
  summarize(rate = mean(zero_att_skip, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = treatment, values_from = rate) %>%
  rename("No Label" = no_label, "Easy Label" = easy_label, "Hard Label" = hard_label)

overall <- df %>%
  group_by(treatment) %>%
  summarize(rate = mean(zero_att_skip, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = treatment, values_from = rate) %>%
  rename("No Label" = no_label, "Easy Label" = easy_label, "Hard Label" = hard_label) %>%
  mutate(puzzle = "Overall")

bind_rows(skip_by_puzzle, overall) %>%
  mutate(across(where(is.numeric), ~ paste0(round(. * 100, 1), "%"))) %>%
  rename(Puzzle = puzzle) %>%
  kable(booktabs = TRUE, align = c("l", "c", "c", "c"),
        caption = "Zero-attempt skip rate by puzzle and treatment arm.") %>%
  kable_styling(latex_options = c("hold_position"), font_size = 10) %>%
  row_spec(nrow(skip_by_puzzle) + 1, bold = TRUE)
```

#### Model Specification

The secondary outcome is zero-attempt skips — rounds where a participant
skipped without entering any response, regardless of how much time or
cognitive effort preceded that decision. We model zero-attempt skips
using a Linear Probability Model (LPM), estimating the probability of a
zero-attempt skip as a linear function of treatment assignment and
covariates:

$$\text{ZeroAttSkip}_{it} = \beta_0 + \beta_1 \text{EasyLabel}_i + \beta_2 \text{HardLabel}_i + \mathbf{X}_i'\boldsymbol{\delta} + \varepsilon_{it}$$

where $\mathbf{X}_i$ includes puzzle skill, age, quantitative exposure,
energy level, and device type. The full-sample model additionally
controls for recruitment source. Standard errors are clustered by
participant (CR1, Satterthwaite degrees of freedom), and coefficients
are interpreted as percentage-point differences in zero-attempt skip
probability relative to the no-label control.

Zero-attempt skips are heavily concentrated in two puzzles. As shown in
Table \@ref(tab:tab-skip-rate), Number Sequence and Tile Sequence
account for the vast majority of cold skips, with overall rates of 23%
and 20% respectively, while Shape Sequence and Hybrid Puzzle see
virtually none. Alphabet Wheel falls in between at around 10%. The easy
label arm consistently shows lower skip rates across puzzles, while hard
label is slightly elevated relative to no label — a pattern that holds
across puzzle types rather than being driven by any single puzzle.

```{r zero-att-model, include=FALSE}
fit_zas <- function(data, control_source = FALSE) {
  fmla <- if (control_source)
    zero_att_skip ~ treatment + puzzleId + puzzleSkill + source +
      age + quantitativeExposure + energyLevel + deviceType
  else
    zero_att_skip ~ treatment + puzzleId + puzzleSkill +
      age + quantitativeExposure + energyLevel + deviceType
  m  <- lm(fmla, data = data)
  vc <- vcovCR(m, cluster = data$userId, type = "CR1")
  rt <- coef_test(m, vcov = vc, test = "Satterthwaite")
  list(m = m, se = sqrt(diag(vc)), p = rt$p_Satt, rt = rt)
}

za_full <- fit_zas(df, control_source = TRUE)
za_org  <- fit_zas(df_organic)
za_pro  <- fit_zas(df_prolific)
```

```{r tab-zas, results="asis", message=FALSE, warning=FALSE}
stargazer(za_full$m, za_org$m, za_pro$m,
          se               = list(za_full$se, za_org$se, za_pro$se),
          p                = list(za_full$p,  za_org$p,  za_pro$p),
          type             = "latex",
          title            = "Effect of Difficulty Labels on Zero-Attempt Skips",
          dep.var.caption  = "",
          dep.var.labels   = c("Full", "Organic", "Prolific"),
          multicolumn      = FALSE,
          covariate.labels = c("Easy Label", "Hard Label"),
          keep             = c("treatmenteasy_label", "treatmenthard_label"),
          report           = "vc*sp",
          omit.stat        = c("f", "ser"),
          label            = "tab:tab-zas",
          header           = FALSE,
          table.placement  = "H",
          font.size        = "small",
          notes            = "LPM; cluster-robust SEs (CR1) by participant.")
```

```{r zas-vals, include=FALSE}
easy_full_pp <- round(abs(za_full$rt$beta[za_full$rt$Coef == "treatmenteasy_label"]) * 100, 1)
easy_full_p  <- round(za_full$rt$p_Satt[za_full$rt$Coef == "treatmenteasy_label"], 3)
hard_full_b  <- round(za_full$rt$beta[za_full$rt$Coef == "treatmenthard_label"], 3)
hard_full_p  <- round(za_full$rt$p_Satt[za_full$rt$Coef == "treatmenthard_label"], 3)
easy_pro_pp  <- round(abs(za_pro$rt$beta[za_pro$rt$Coef == "treatmenteasy_label"]) * 100, 1)
easy_pro_p   <- round(za_pro$rt$p_Satt[za_pro$rt$Coef == "treatmenteasy_label"], 3)
easy_org_b   <- round(za_org$rt$beta[za_org$rt$Coef == "treatmenteasy_label"], 3)
easy_org_p   <- round(za_org$rt$p_Satt[za_org$rt$Coef == "treatmenteasy_label"], 3)
```

The linear model estimates the effect of difficulty labels on the
probability of skipping a puzzle without any answer attempt (Table
\@ref(tab:tab-zas)). The hard label shows no meaningful effect no matter
how we slice the sample — the coefficient is near zero and
non-significant in the full sample ($\hat{\beta}_{\text{hard}}$ =
`r hard_full_b`, p = `r hard_full_p`), among organic participants, and
among Prolific participants.

The easy label effect on zero-attempt skips is driven entirely by
Prolific participants. Among Prolific participants, the easy label
reduces the zero-attempt skip rate by `r easy_pro_pp` percentage points
(p = `r easy_pro_p`), while among organic participants the effect is
small and non-significant ($\hat{\beta}_{\text{easy}}$ = `r easy_org_b`,
p = `r easy_org_p`). When the two groups are pooled and recruitment
source is controlled for, the organic null attenuates the Prolific
signal, yielding a marginally significant estimate of `r easy_full_pp`
percentage points (p = `r easy_full_p`) in the full sample.

```{r fig-skip-coef, fig.width=5, fig.height=2.5, fig.pos="H", fig.align="center", fig.cap="Estimated effect of difficulty labels on zero-attempt skip probability (linear probability model). Coefficients represent percentage-point change relative to the no-label control. Error bars are 95\\% CIs based on cluster-robust SEs.", out.width="0.65\\textwidth"}
rt_zas <- za_full$rt
coef_df <- data.frame(
  term = c("Easy Label", "Hard Label"),
  est  = c(rt_zas$beta[rt_zas$Coef == "treatmenteasy_label"],
           rt_zas$beta[rt_zas$Coef == "treatmenthard_label"]),
  se   = c(rt_zas$SE[rt_zas$Coef == "treatmenteasy_label"],
           rt_zas$SE[rt_zas$Coef == "treatmenthard_label"])
) %>% mutate(lo = est - 1.96 * se, hi = est + 1.96 * se)

ggplot(coef_df, aes(x = est, y = term)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(aes(xmin = lo, xmax = hi), width = 0.15) +
  geom_point(size = 3) +
  labs(x = "Effect on Zero-Attempt Skip Rate (pp)", y = NULL,
       title    = "Effect of Difficulty Labels on Zero-Attempt Skips",
       subtitle = "Relative to no-label control; error bars are 95% CIs") +
  theme_bw(base_size = 9)
```

As shown in Figure \@ref(fig:fig-skip-coef), the easy label coefficient
is negative and pointing toward fewer cold skips relative to no label.
However, the confidence interval is wide and crosses zero in the full
sample. The hard label estimate is close to zero with a wide interval in
both directions, so there's no real signal there.

The difference in the easy label's effect between Prolific and organic
participants likely reflects differences in how each group engages with
task framing. Prolific participants are experienced study participants
who tend to read instructions carefully and take task cues seriously —
when told a puzzle is easy, they internalize that label and may feel a
stronger obligation to at least attempt an answer before skipping.
Organic participants, recruited informally through personal networks,
tend to be more distracted and less attentive to framing; we also found
they spent significantly more time with the browser tab inactive,
suggesting more real-world interruptions that could override any label
effect. In short, the easy label appears to work by making cold-skipping
feel less acceptable, but that mechanism requires the participant to be
paying attention to the framing in the first place.

### Secondary Model: Effect of Difficulty Labels on Time Before Skipping

#### Outcome Description

#### Model Specification

Similar to the primary model, we estimated the effect of difficulty
labels on log time engaged before skipping. This is a linear
mixed-effects model with puzzle fixed effects, a participant-level
random intercept, and pre-treatment covariates:

$$\log(\text{TimeEngagedBeforeSkip}_{it}) = \beta_0 + \beta_1 \text{EasyLabel}_i + \beta_2 \text{HardLabel}_i + \mathbf{X}_i'\boldsymbol{\delta} + \varepsilon_{it}$$

```{r prepare-skip-robustness, include=FALSE}

prepare_skip_sample <- function(data) {
  data %>%
    filter(didSkip == TRUE) %>%
    drop_na(
      timeEngaged,
      treatment,
      puzzleId,
      puzzleSkill,
      age,
      quantitativeExposure,
      energyLevel,
      deviceType,
      userId
    ) %>%
    mutate(
      treatment = relevel(
        factor(treatment),
        ref = "no_label"
      ),
      puzzleId = relevel(
        factor(puzzleId),
        ref = "puzzle_01"
      ),
      logTE = log(timeEngaged + 1)
    )
}

skip_full <- prepare_skip_sample(df) %>%
  drop_na(source)

skip_org <- prepare_skip_sample(df_organic)

skip_pro <- prepare_skip_sample(df_prolific)

```

```{r, include=FALSE}
tibble(
  Sample = c("Full", "Organic", "Prolific"),
  `Skipped Rounds` = c(
    nrow(skip_full),
    nrow(skip_org),
    nrow(skip_pro)
  ),
  Participants = c(
    n_distinct(skip_full$userId),
    n_distinct(skip_org$userId),
    n_distinct(skip_pro$userId)
  )
)
```

```{r fit-skip-robustness, include=FALSE}


fit_skip_source <- function(data, control_source = FALSE) {

  fmla <- if (control_source) {
    logTE ~ treatment + puzzleId + puzzleSkill + source +
      age + quantitativeExposure + energyLevel + deviceType +
      (1 | userId)
  } else {
    logTE ~ treatment + puzzleId + puzzleSkill +
      age + quantitativeExposure + energyLevel + deviceType +
      (1 | userId)
  }

  m <- lmer(
    fmla,
    data = data,
    REML = FALSE
  )

  vc <- vcovCR(
    m,
    cluster = data$userId,
    type = "CR1"
  )

  rt <- coef_test(
    m,
    vcov = vc,
    test = "Satterthwaite"
  )

  list(
    m = m,
    se = sqrt(diag(vc)),
    p = rt$p_Satt,
    rt = rt
  )
}

skip_fit_full <- fit_skip_source(
  skip_full,
  control_source = TRUE
)

skip_fit_org <- fit_skip_source(skip_org)

skip_fit_pro <- fit_skip_source(skip_pro)

```

```{r tab-skip-source-final, results="asis", message=FALSE, warning=FALSE}
stargazer(
  skip_fit_full$m,
  skip_fit_org$m,
  skip_fit_pro$m,
  se               = list(
    skip_fit_full$se,
    skip_fit_org$se,
    skip_fit_pro$se
  ),
  p                = list(
    skip_fit_full$p,
    skip_fit_org$p,
    skip_fit_pro$p
  ),
  type             = "latex",
  title            = "Effect of Difficulty Labels on Time Before Skipping",
  dep.var.caption  = "",
  dep.var.labels   = c("Full", "Organic", "Prolific"),
  multicolumn      = FALSE,
  covariate.labels = c("Easy Label", "Hard Label"),
  keep             = c(
    "treatmenteasy_label",
    "treatmenthard_label"
  ),
  report           = "vc*sp",
  omit.stat        = c("f", "ser", "ll", "aic", "bic"),
  label            = "tab:tab-skip-source",
  header           = FALSE,
  table.placement  = "H",
  font.size        = "small",
  notes            = "Mixed-effects models; CR1 cluster-robust SEs by participant."
)
```

### Exploratory Model: Effect of Difficulty Labels on Attempt Count

Given that the primary outcome showed no significant treatment effect,
we explored whether difficulty labels influenced participant behavior
through a different channel — how hard they tried before giving up. This
analysis is exploratory and not intended to support causal claims.

#### Outcome Distribution

```{r att-count-stats, include=FALSE}
mean_att          <- mean(df$attemptCount)
var_att           <- var(df$attemptCount)
num_zero_att      <- sum(df$attemptCount == 0)
num_responses_att <- length(df$attemptCount)
prop_of_zero      <- num_zero_att / num_responses_att
```

Attempt count measures how many answers a participant submitted before
either solving the puzzle or skipping it. Most participants make only a
few attempts, but a small number make many, producing a right-skewed
distribution with a mean of `r round(mean_att, 1)` attempts per round
and a variance of `r round(var_att, 1)` — well above the mean,
indicating overdispersion.

```{r fig-log-justification-ATT-COUNT, fig.width=5, fig.height=3.5, fig.cap="Distribution of attempt count per puzzle round. The distribution is right-skewed with most participants making few attempts and a sparse tail of high-attempt rounds.", fig.pos="H", out.width="0.65\\textwidth", fig.align="center"}
ggplot(df, aes(x = attemptCount)) +
  geom_histogram(bins = 20, fill = "steelblue", color = "white") +
  labs(x = "Attempt Count", y = "Count", title = "Attempt Count Distribution") +
  theme_bw(base_size = 9)
```

#### Model Specification

Attempt count is a non-negative integer with variance
(`r round(var_att, 1)`) much larger than the mean
(`r round(mean_att, 1)`), which rules out a Poisson model and points to
negative binomial regression. Zero inflation is not a concern —
`r round(prop_of_zero * 100, 1)`% of rounds have zero attempts, which is
consistent with natural count variation rather than a separate
zero-generating process.

Adding a participant-level random intercept is not feasible here. A
mixed-effects negative binomial model (glmer.nb) would be the natural
extension, but cluster-robust variance estimation is not supported for
mixed-effects models of this class in the software used. As an
alternative, we considered including participant as a fixed effect, but
since each participant is assigned to exactly one treatment arm,
participant dummies are perfectly collinear with treatment, making the
treatment effect unidentifiable. We therefore drop the random intercept
entirely and cluster standard errors at the participant level to account
for within-person correlation.

$$E[\text{attemptCount}_{it}] = \exp\!\Big(\beta_0 + \beta_1\,\text{EasyLabel}_i + \beta_2\,\text{HardLabel}_i + \boldsymbol{\gamma}_t + \mathbf{X}_i'\boldsymbol{\delta}\Big)$$

where $\boldsymbol{\gamma}_t$ are puzzle fixed effects and
$\mathbf{X}_i$ includes puzzle skill, age, quantitative exposure, energy
level, and device type. The full-sample model additionally controls for
recruitment source. Coefficients are interpreted as multiplicative
effects on expected attempt count.

```{r attcount-model, message=FALSE, warning=FALSE}

# Full sample (controls for source)
m_nb <- glm.nb(
  attemptCount ~ treatment + puzzleId + puzzleSkill + source +
            age + quantitativeExposure + energyLevel + deviceType,
  data = df
)

vcov_m <- vcovCL(m_nb, cluster = df$userId)
rt<-coeftest(m_nb, vcov = vcov_m)


# Organic subsample
m_org <- glm.nb(
  attemptCount ~ treatment + puzzleId + puzzleSkill  +
            age + quantitativeExposure + energyLevel + deviceType,
  data = df_organic
)

vcov_org <- vcovCL(m_org, cluster = df_organic$userId)
rt_org<-coeftest(m_org, vcov = vcov_org)


# Prolific subsample
m_pro <- glm.nb(
  attemptCount ~ treatment + puzzleId + puzzleSkill  +
            age + quantitativeExposure + energyLevel + deviceType,
  data = df_prolific
)

vcov_pro <- vcovCL(m_pro, cluster = df_prolific$userId)
rt_pro<-coeftest(m_pro, vcov = vcov_pro)

modelsummary(
  list("Full" = m_nb, "Organic" = m_org, "Prolific" = m_pro),
  vcov      = list(vcov_m, vcov_org, vcov_pro),
  statistic = c("({std.error})", "p = {p.value}"),
  coef_map  = c("treatmenteasy_label" = "Easy Label",
                "treatmenthard_label" = "Hard Label",
                "puzzleIdpuzzle_02" = "Puzzle 2",
                "puzzleIdpuzzle_03" = "Puzzle 3",
                "puzzleIdpuzzle_04" = "Puzzle 4",
                "puzzleIdpuzzle_05" = "Puzzle 5"),
  gof_omit  = ".*",
  stars     = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  title     = "Effect of Difficulty Labels on Attempt Count",
  booktabs  = TRUE,
  escape    = FALSE,
  output    = "kableExtra"
) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9) %>%
  footnote(general = "Negative binomial model; cluster-robust SEs (HC1) by participant.",
           threeparttable = TRUE, escape = FALSE)
```

```{r attcount-vals, include=FALSE}
exp(.08)
rt_pro
rt_pro[40]
m1_easy_b  <- round(rt[2], 3)
m1_easy_se <- round(rt[15], 3)
m1_easy_p  <- round(rt[41], 3)
m1_hard_b  <- round(rt[3], 3)
m1_hard_se <- round(rt[16], 3)
m1_hard_p  <- round(rt[42], 3)

m1_puz2_b_full  <- round(rt[4], 3)
m1_puz2_se_full <- round(rt[17], 3)
m1_puz2_p_full  <- round(rt[43], 3)

m1_puz2_b_org  <- round(rt_org[4], 3)
m1_puz2_se_org <- round(rt_org[16], 3)
m1_puz2_p_org  <- round(rt_org[40], 3)

m1_puz2_b_pro  <- round(rt_pro[4], 3)
m1_puz2_se_pro<- round(rt_pro[16], 3)
m1_puz2_p_pro  <- round(rt_pro[40], 3)

m1_puz3_b_full  <- round(rt[5], 3)
m1_puz3_se_full <- round(rt[18], 3)
m1_puz3_p_full  <- round(rt[44], 3)

m1_puz3_b_org  <- round(rt_org[5], 3)
m1_puz3_se_org <- round(rt_org[17], 3)
m1_puz3_p_org  <- round(rt_org[41], 3)

m1_puz3_b_pro  <- round(rt_pro[5], 3)
m1_puz3_se_pro <- round(rt_pro[17], 3)
m1_puz3_p_pro  <- round(rt_pro[41], 3)

m1_puz4_b_full  <- round(rt[6], 3)
m1_puz4_b_pro  <- round(rt_pro[6], 3)
m1_puz4_b_org  <- round(rt_org[6], 3)

m1_puz5_b_full  <- round(rt[7], 3)
m1_puz5_b_pro  <- round(rt_pro[7], 3)
m1_puz5_b_org  <- round(rt_org[7], 3)

m1_puz3_b_pro  
m1_puz3_se_pro
m1_puz3_p_pro
round(exp(m1_puz3_b_full),3)
round(exp(m1_puz3_b_pro),3)
round(exp(m1_puz3_b_org),3)
```

Neither label had a statistically significant effect on attempt count
across any of the three samples. Coefficients for easy and hard labels
are small with wide standard errors, indicating that difficulty labels
did not change how many times participants tried before solving or
skipping.

Puzzle identity, by contrast, consistently predicts attempt count.
Relative to the Number Sequence (puzzle 1), the Alphabet Wheel (puzzle
2) and Shape Sequence (puzzle 3) elicited substantially more attempts
across all three samples, significant at the 1% level in every group. In
the full sample, participants made
`r round((exp(m1_puz2_b_full) - 1) * 100, 1)`% more attempts on puzzle 2
and `r round((exp(m1_puz3_b_full) - 1) * 100, 1)`% more on puzzle 3
relative to puzzle 1. Among organic participants, the same puzzles
showed increases of `r round((exp(m1_puz2_b_org) - 1) * 100, 1)`% and
`r round((exp(m1_puz3_b_org) - 1) * 100, 1)`%, and among Prolific
participants, `r round((exp(m1_puz2_b_pro) - 1) * 100, 1)`% and
`r round((exp(m1_puz3_b_pro) - 1) * 100, 1)`%. Puzzle 4 and puzzle 5
also show positive associations, though significance varies by sample.
These puzzle-level differences reflect variation in objective puzzle
difficulty and should not be interpreted as causal — puzzle identity was
not randomly assigned.

## Conclusion

The primary model found no statistically significant effect of
difficulty labels on log time engaged in the full sample. The hard label
was consistently negative across both subgroups and marginally
significant at the 10% level, but in the opposite direction from our
hypothesis — participants shown the hard label spent slightly less time
engaged, not more.

The zero-attempt skip model showed a marginal effect of the easy label
in the full sample (p = `r easy_full_p`), driven entirely by Prolific
participants, who reduced their zero-attempt skip rate by
`r easy_pro_pp` percentage points under the easy label (p =
`r easy_pro_p`). This likely reflects how each group engages with task
framing: Prolific participants are experienced study participants who
tend to take task cues seriously and may feel more obligated to attempt
a puzzle they were told is easy. Organic participants, with no financial
incentive and more real-world distractions, showed no such response.

The exploratory attempt count model found no treatment effects across
any sample. Puzzle identity was the dominant predictor, with some
puzzles consistently eliciting far more attempts than others — but since
puzzle assignment was not randomized, these are not causal estimates.

Taken together, the results are inconclusive rather than null. The
marginal signals — a negative hard label effect on time engaged and a
reduction in cold skipping among Prolific participants shown the easy
label — suggest that framing may matter, but we did not have enough
precision to detect it cleanly. The easy label finding among Prolific
participants is the most interpretable: when participants are attentive
and motivated, telling them a puzzle is easy appears to make them less
willing to abandon it without trying. Organic participants, more
distracted and without a completion incentive, showed no such response.

The experiment was underpowered and mixing two recruitment populations
with very different engagement conditions inflated variance beyond what
was anticipated. A better-powered replication using Prolific only,
targeting `r thresh_both * 3` participants, would let us more cleanly
test whether the easy label truly reduces cold skipping, whether the
hard label's negative effect on engagement holds up or was noise, and
whether the framing effects on time engaged are detectable at adequate
power. It would also be worth testing whether the effect depends on the
mismatch itself, specifically whether an easy label on a genuinely easy
puzzle produces a different response than an easy label on a hard one.

## Appendix

### A. Residual Diagnostics

```{r fig-qq, fig.width=5, fig.height=5, fig.pos="H", fig.cap="Normal Q-Q plot of residuals from the primary model (log time engaged ~ difficulty label + puzzle fixed effects + covariates + participant random intercept). Residuals track the reference line closely across the central range, with minor deviation in the lower tail, consistent with a light left tail after log transformation."}
qqnorm(residuals(m), main = "Normal Q-Q Plot of Model Residuals")
qqline(residuals(m), col = "steelblue")
```
