---
title: "The Effect of Difficulty Labels on Puzzle Engagement"
author: ""
date: "`r Sys.Date()`"
output: bookdown::pdf_document2
bibliography: references.bib
header-includes:
  - \usepackage{booktabs}
  - \usepackage[table]{xcolor}
  - \usepackage{threeparttable}
  - \usepackage{multirow}
  - \usepackage{float}
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE, error = TRUE)
```

```{r load-data, include=FALSE}
library(dplyr)
library(tidyr)
library(purrr)
library(lme4)
library(clubSandwich)
library(knitr)
library(kableExtra)
library(modelsummary)
library(jsonlite)
library(consort)

master      <- readRDS("analysis_data.rds")
df          <- master %>% filter(sessionComplete == TRUE)
df_prolific <- df %>% filter(source == "Prolific")
df_organic  <- df %>% filter(source == "Organic")
```

## Introduction

### Theory

Priming refers to the phenomenon by which exposure to a stimulus
influences subsequent perception, cognition, or behavior. Early work on
semantic priming established that related concepts are recognized faster
when preceded by an associated word [@meyer1971facilitation], and
neuroimaging research has since confirmed consistent neural signatures
of repeated stimulus exposure [@lee2020neural]. Taken together, this
body of evidence suggests that priming can shape cognitive processing in
measurable ways.

Our study tests whether explicit difficulty labels, a concrete and
ecologically valid form of priming, alter persistence on cognitive
tasks. Difficulty labels are a natural mechanism through which priming
operates in everyday settings: a task framed as easy invites different
expectations and effort allocation than the same task framed as hard.
@kahneman2011thinking describes how the framing of information shapes
intuitive judgment before deliberate reasoning engages, which suggests
that a difficulty label encountered before a task begins could prime
effort expectations that carry into performance.

### Hypothesis

We hypothesize that the two label conditions affect time engaged in
opposite directions relative to the no label control. Participants shown
the hard label receive accurate information about task difficulty and
may mentally prepare for sustained effort, increasing persistence.
Participants shown the easy label encounter a mismatch between their
expectation and their experience: the puzzle is harder than primed,
which we expect to produce frustration that shortens engagement. When
participants expect a task to be easy and immediately find it is not,
the experience may feel like failure rather than challenge, lowering the
threshold for giving up and skipping.

Formally:

$$H_1: \mu_{\text{hard}} > \mu_{\text{no label}}$$
$$H_2: \mu_{\text{easy}} < \mu_{\text{no label}}$$
$$H_3: \mu_{\text{hard}} > \mu_{\text{easy}}$$

where $\mu$ denotes mean time engaged per puzzle round.

## Power Calculation

To determine the necessary sample size, we conducted a simulation-based
power analysis rather than relying on a closed-form analytic formula.
This approach allows us to model the multi-round panel structure of the
experiment and account for participant-level clustering.

We set a control baseline average solve time of $\mu = 270$ seconds.
Variance was decomposed into a between-person component ($\sigma_u = 60$
seconds) and a within-person residual ($\sigma_e = 67$ seconds),
yielding a total standard deviation of
$\sigma = \sqrt{\sigma_u^2 + \sigma_e^2} \approx 90$ seconds, chosen
conservatively to capture natural individual variation in performance.
We specified treatment effects relative to the no label control in
opposite directions for the two label arms: a $+25$ second increase for
the hard label arm, reflecting anticipated persistence when participants
are accurately primed for difficulty, and a $-25$ second decrease for
the easy label arm, reflecting earlier disengagement when participants
encounter a task harder than the label suggested.

We simulated 500 iterations at escalating sample sizes. For each
iteration, a mixed-effects model was fit with puzzle fixed effects and a
participant-level random intercept:

$$Y_{it} = \beta_0 + \boldsymbol{\beta}_{\text{trt}} \text{Treatment}_i + \boldsymbol{\gamma}_t + u_i + \varepsilon_{it}$$

where $\boldsymbol{\gamma}_t$ are puzzle fixed effects and
$u_i \sim N(0, \sigma_u^2)$ is a participant random intercept. Because
data are generated exactly from the model, standard mixed-model
$t$-tests are used to assess significance; cluster-robust standard
errors are applied in the actual analysis but are not needed for
simulation where the variance structure is known.

```{r power-sim, message=FALSE, warning=FALSE}
if (file.exists("power_sim_results.rds")) {
  power_res <- readRDS("power_sim_results.rds")
} else {
  set.seed(241)
  mu_base   <- 270; beta_hard <- 25; beta_easy <- -25
  sigma_u   <-  60; sigma_e   <- 67; n_puzzles  <-  5; n_sim <- 500
  puzzle_fx <- c(-15, 10, 20, -8, -7)

  sim_power <- function(n_per_arm) {
    sig_hard <- sig_easy <- logical(n_sim)
    for (s in seq_len(n_sim)) {
      n_total  <- n_per_arm * 3
      trt_id   <- rep(c("no_label","easy_label","hard_label"), each = n_per_arm)
      u_i      <- rnorm(n_total, 0, sigma_u)
      userId   <- rep(seq_len(n_total), each = n_puzzles)
      puzzleId <- rep(seq_len(n_puzzles), times = n_total)
      trt_long <- rep(trt_id, each = n_puzzles)
      beta_trt <- ifelse(trt_long == "hard_label", beta_hard,
                  ifelse(trt_long == "easy_label", beta_easy, 0))
      Y <- mu_base + beta_trt + puzzle_fx[puzzleId] + u_i[userId] +
           rnorm(n_total * n_puzzles, 0, sigma_e)
      dat <- data.frame(
        Y         = Y,
        treatment = factor(trt_long, levels = c("no_label","easy_label","hard_label")),
        puzzleId  = factor(puzzleId),
        userId    = factor(userId)
      )
      m <- tryCatch(lmer(Y ~ treatment + puzzleId + (1|userId),
                         data = dat, REML = FALSE), error = function(e) NULL)
      if (is.null(m)) next
      ct <- summary(m)$coefficients
      th <- ct["treatmenthard_label", "t value"]
      te <- ct["treatmenteasy_label", "t value"]
      if (length(th)) sig_hard[s] <- abs(th) > 1.96
      if (length(te)) sig_easy[s] <- abs(te) > 1.96
    }
    data.frame(n_per_arm = n_per_arm,
               power_hard = mean(sig_hard),
               power_easy = mean(sig_easy))
  }

  power_res <- bind_rows(lapply(seq(40, 150, by = 10), sim_power))
  saveRDS(power_res, "power_sim_results.rds")
}

thresh_hard <- power_res$n_per_arm[which(power_res$power_hard >= 0.80)[1]]
thresh_easy <- power_res$n_per_arm[which(power_res$power_easy >= 0.80)[1]]
thresh_both <- max(thresh_hard, thresh_easy, na.rm = TRUE)
```

At $\alpha = 0.05$, power to detect the hard label effect crosses the
80% threshold at `r thresh_hard` participants per arm
(`r thresh_hard * 3` total), and the easy label effect crosses 80% at
`r thresh_easy` per arm (`r thresh_easy * 3` total). The
harder-to-detect effect requires `r thresh_both` per arm
(`r thresh_both * 3` total).

The final analytic sample of 256 completers falls below the
pre-specified target of `r thresh_both * 3`. Exclusions for behavioral
flags and attrition reduced the sample beyond what was anticipated at
the design stage, and arm sizes are unequal (No Label: 87, Easy Label:
80, Hard Label: 89), which modestly reduces power relative to equal
allocation. Beyond sample size, the analytic sample pools two
recruitment populations that differ systematically in engagement
conditions: Prolific participants completed the experiment in a
structured, incentivized setting with low distraction, while organic
participants, recruited through personal networks, varied more in
motivation, setting, and distraction level. This heterogeneity inflates
within-group variance beyond what a single $\sigma$ captures, further
eroding effective power. Both the sample shortfall and the variance
inflation from pooling two behaviorally distinct populations should be
kept in mind when interpreting null or marginal findings.

## Study Design

### Procedure

The experiment was delivered as a browser-based web application at
anotherpuzzle.io. The application was a single-page JavaScript game
engine that managed the complete participant session from recruitment
intake through behavioral data collection without any researcher
intervention at runtime. All data were logged directly to a cloud
database (Firestore) at both the session and round levels.

Upon arriving at the application, participants were presented with a
consent screen. Participants who agreed were immediately and invisibly
assigned to one of three treatment arms before proceeding. Following
consent, participants completed an eight-item survey collecting baseline
covariates: age, gender, education level, quantitative exposure, current
energy level, puzzle frequency, self-rated puzzle skill, and puzzle
enjoyment. After the survey, participants were shown a brief instruction
screen and then clicked "Start the Challenge," which triggered
randomization of puzzle order and the creation of the participant's
session record.

The game engine logged behavioral data at the round level. For each
round, it recorded: total time engaged (a continuous timer that paused
automatically during tab switches and participant-initiated breaks),
number of answer attempts, whether the first attempt was correct, time
to first attempt, tab-switch count and total time paused, rage-click
events, and whether the round ended in a solve or a skip. Skip was
enabled after 30 seconds of engagement, preventing immediate abandonment
without any attempt. At the session level, the engine recorded device
type, puzzle presentation order, recruitment source, and session
completion status.

### Treatment

The experimental treatment was a difficulty label presented to
participants before each logic puzzle. The hard label was expected to
prompt mental preparation for a demanding task, increasing persistence.
The easy label was expected to create an expectation mismatch:
participants anticipating an easy task would encounter greater
difficulty than expected, producing frustration that shortens engagement
before participants choose to skip.

Participants were randomly assigned at session entry to one of three
conditions: (1) no label, (2) easy label, or (3) hard label. The
assigned label was shown before every puzzle in the session, so a
participant assigned to the hard label condition saw that label at the
start of each of the five rounds. Within the application, all other
aspects of the experience were held constant across arms. The interface,
instructions, and the five logic puzzles were identical across
conditions, so differences in behavior across treatment groups can be
attributed to the difficulty label rather than any other feature of the
experience.

Each session consisted of five logic puzzles presented in randomized
order: a number sequence, an alphabet wheel, a shape sequence, a hybrid
shape-number puzzle, and a tile sequence. All five were designed to be
intentionally challenging, though they varied in how difficult
participants found them in practice.

```{r puzzle-skip-rate}
overall_rates <- df %>%
  mutate(
    Puzzle = case_when(
      puzzleId == "puzzle_01" ~ "Number Sequence",
      puzzleId == "puzzle_02" ~ "Alphabet Wheel",
      puzzleId == "puzzle_03" ~ "Shape Sequence",
      puzzleId == "puzzle_04" ~ "Hybrid Puzzle",
      puzzleId == "puzzle_05" ~ "Tile Sequence"
    )
  ) %>%
  group_by(Puzzle) %>%
  summarize(overall_raw = mean(didSkip, na.rm = TRUE), .groups = "drop")

df %>%
  mutate(
    Puzzle = case_when(
      puzzleId == "puzzle_01" ~ "Number Sequence",
      puzzleId == "puzzle_02" ~ "Alphabet Wheel",
      puzzleId == "puzzle_03" ~ "Shape Sequence",
      puzzleId == "puzzle_04" ~ "Hybrid Puzzle",
      puzzleId == "puzzle_05" ~ "Tile Sequence"
    )
  ) %>%
  group_by(Puzzle, treatment) %>%
  summarize(skip_rate = mean(didSkip, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = treatment, values_from = skip_rate) %>%
  left_join(overall_rates, by = "Puzzle") %>%
  arrange(desc(overall_raw)) %>%
  mutate(across(c(no_label, easy_label, hard_label, overall_raw),
                ~ paste0(round(. * 100, 1), "%"))) %>%
  select(Puzzle, no_label, easy_label, hard_label, overall_raw) %>%
  rename(
    `No Label`   = no_label,
    `Easy Label` = easy_label,
    `Hard Label` = hard_label,
    `Overall`    = overall_raw
  ) %>%
  kable(booktabs = TRUE,
        caption  = "Skip Rate by Puzzle and Treatment Arm") %>%
  kable_styling(latex_options = c("hold_position"))
```

Skip and solve rates differed meaningfully across puzzle types, as shown
in Table \@ref(tab:puzzle-skip-rate). Puzzle order was randomized across
participants to prevent any one puzzle from systematically falling at
the same point in the session, and the analysis accounts for
puzzle-level differences and within-participant correlation across
rounds.

Treatment delivery was controlled entirely through the application, and
the data contain no indication that any participant was shown a label
other than their assigned condition; we therefore identified no
treatment non-compliers. However, the application did not independently
log the randomized assignment alongside the label rendered on screen for
each round, so perfect delivery could not be independently verified.
Future experiments should log these separately to enable a formal
compliance check.

## Participants

The experiment database contained 399 session records. One duplicate
user ID was removed, leaving 398 unique participants. Eight were
excluded as returning or replay players whose devices had previously
completed a full session, leaving 390 first-time participants. An
additional 27 had no retained gamelog records and were excluded,
yielding 363 participants who started at least one puzzle round.

Twenty-one participants were excluded for eligibility: 8 were under 18
and 13 were rejected through Prolific's quality review. Of the remaining
342 eligible participants, 39 were removed for behavioral or
data-quality reasons:

*Behavioral exclusions*

1.  Total tab-switch pause time exceeded 600 seconds across the session,
    indicating the participant was away from the browser for an extended
    period.
2.  At least one puzzle showed an idle walk-away pattern: the
    participant skipped without switching tabs and with time engaged
    above the 97th percentile of the skip-time distribution for that
    puzzle, suggesting they left the page and returned.
3.  All five puzzles were skipped within 30 to 35 seconds with no answer
    attempts, indicating the participant waited out the skip-unlock
    timer on every round rather than engaging with the puzzles.

*AI assistance exclusions*

4.  Self-reported use of AI assistance in the post-session honesty
    check.
5.  Hard puzzles solved on the first attempt faster than population
    benchmarks for that puzzle (any single puzzle below the 5th
    percentile, or two or more below the 10th), subject to manual
    review. Flagged participants were retained if their fast solve times
    were consistent with their relative performance across all rounds,
    as genuine skill tends to produce proportionally fast times across
    puzzles rather than selectively fast times on isolated ones.

Attrition, defined as beginning treatment but failing to complete all
five rounds, accounted for 47 participants. The final analytic sample is
256 completers across three arms: No Label (n = 87), Easy Label (n =
80), and Hard Label (n = 89). Figure \@ref(fig:fig-consort) shows the
full participant flow.

```{r fig-consort, fig.width=9, fig.height=7, fig.cap="Participant flow through the experiment.", fig.pos="H", out.width="\\textwidth"}

raw_json  <- fromJSON("firestore_export.json", simplifyVector = FALSE)
users_all <- tibble(document_id = names(raw_json$users),
                    data        = raw_json$users) %>%
  unnest_wider(data) %>%
  distinct(userId, .keep_all = TRUE)

new_player_ids <- users_all %>%
  filter(!is.na(sharedDevice) & sharedDevice == "new_player") %>%
  pull(userId)

em_u          <- readRDS("experiment_master.rds") %>% distinct(userId, .keep_all = TRUE)
master_u      <- master %>% distinct(userId, .keep_all = TRUE)
completed_ids <- master_u %>% filter(sessionComplete == TRUE) %>% pull(userId)
dropped_ids   <- master_u %>% filter(sessionComplete == FALSE) %>% pull(userId)

arm_src <- master %>%
  filter(sessionComplete == TRUE) %>%
  distinct(userId, .keep_all = TRUE) %>%
  count(treatment, source)

gn <- function(trt, src) {
  v <- arm_src %>% filter(treatment == trt, source == src) %>% pull(n)
  if (length(v) == 0) 0L else v
}
n_no_org   <- gn("no_label",   "Organic")
n_no_pro   <- gn("no_label",   "Prolific")
n_easy_org <- gn("easy_label", "Organic")
n_easy_pro <- gn("easy_label", "Prolific")
n_hard_org <- gn("hard_label", "Organic")
n_hard_pro <- gn("hard_label", "Prolific")

flow <- users_all %>%
  mutate(
    exc_replay = case_when(
      !is.na(sharedDevice) & sharedDevice == "returning" ~ "Returning player",
      !is.na(isReplay) & isReplay == TRUE & !(userId %in% new_player_ids) ~ "Replay session",
      TRUE ~ NA_character_
    ),
    after_replay = if_else(is.na(exc_replay), userId, NA_character_),
    exc_no_game = case_when(
      !is.na(exc_replay) ~ NA_character_,
      !(userId %in% em_u$userId) ~ "No gamelog records",
      TRUE ~ NA_character_
    ),
    started = if_else(
      is.na(exc_replay) & (userId %in% em_u$userId), userId, NA_character_
    ),
    exc_elig = case_when(
      !is.na(exc_replay) | !(userId %in% em_u$userId) ~ NA_character_,
      !is.na(age) & age < 18 ~ "Age < 18",
      !is.na(prolificRejected) & prolificRejected == TRUE ~ "Prolific rejection",
      TRUE ~ NA_character_
    ),
    eligible = if_else(
      is.na(exc_replay) & (userId %in% em_u$userId) & is.na(exc_elig),
      userId, NA_character_
    ),
    exc_behav = case_when(
      !is.na(exc_replay) | !(userId %in% em_u$userId) | !is.na(exc_elig) ~ NA_character_,
      !(userId %in% master_u$userId) ~ "Behavioral or AI flag",
      TRUE ~ NA_character_
    ),
    quality = if_else(
      is.na(exc_replay) & (userId %in% em_u$userId) &
        is.na(exc_elig) & (userId %in% master_u$userId),
      userId, NA_character_
    ),
    exc_att = case_when(
      !is.na(exc_replay) | !(userId %in% em_u$userId) |
        !is.na(exc_elig) | !(userId %in% master_u$userId) ~ NA_character_,
      userId %in% dropped_ids ~ "Did not complete all 5 rounds",
      TRUE ~ NA_character_
    ),
    arm = case_when(
      !(userId %in% completed_ids) ~ NA_character_,
      treatment == "no_label" ~
        paste0("No Label\nOrganic: ", n_no_org, "\nProlific: ", n_no_pro, "\n"),
      treatment == "easy_label" ~
        paste0("Easy Label\nOrganic: ", n_easy_org, "\nProlific: ", n_easy_pro, "\n"),
      treatment == "hard_label" ~
        paste0("Hard Label\nOrganic: ", n_hard_org, "\nProlific: ", n_hard_pro, "\n")
    )
  )

p <- consort_plot(
  data   = flow,
  orders = c(
    userId       = 'Users with a session record',
    exc_replay   = 'Excluded',
    after_replay = 'Non-replay participants',
    exc_no_game  = 'Excluded',
    started      = 'Participants who started',
    exc_elig     = 'Excluded',
    eligible     = 'Eligible participants',
    exc_behav    = 'Excluded',
    quality      = 'Passed quality checks',
    exc_att      = 'Did not complete all 5 rounds',
    arm          = 'Analyzed'
  ),
  side_box   = c('exc_replay', 'exc_no_game', 'exc_elig', 'exc_behav', 'exc_att'),
  allocation = 'arm',
  cex        = 0.7
)

plot(p)
```

## Potential Outcomes

Each participant has three potential outcomes, the time engaged they
would produce under each of the three label conditions:
$Y_i(\text{no label})$, $Y_i(\text{easy label})$, and
$Y_i(\text{hard label})$. By the fundamental problem of causal
inference, only one of these is ever observed for any given participant.
Random assignment to one of the three arms makes the unobserved
potential outcomes independent of treatment assignment, allowing average
potential outcomes to be estimated from group means.

The primary comparisons of interest are the average treatment effects of
each label condition relative to the no label control, and the direct
contrast between the two label conditions:

$$\tau_{\text{easy}} = E[Y_i(\text{easy label})] - E[Y_i(\text{no label})]$$

$$\tau_{\text{hard}} = E[Y_i(\text{hard label})] - E[Y_i(\text{no label})]$$

$$\tau_{\text{hard vs. easy}} = E[Y_i(\text{hard label})] - E[Y_i(\text{easy label})]$$

The no label arm serves as the reference group, representing engagement
in the absence of any difficulty prime. The first two comparisons
isolate the effect of priming participants with a difficulty expectation
relative to no expectation at all, holding everything else constant. The
third captures whether the direction of the prime matters: even if both
labels shift behavior away from the no-label baseline, the frustration
mechanism and the persistence mechanism predict that their effects move
in opposite directions. Consistent with the hypotheses, we expect
$\tau_{\text{easy}} < 0$, reflecting frustration-driven disengagement,
$\tau_{\text{hard}} > 0$, reflecting preparation-driven persistence, and
$\tau_{\text{hard vs. easy}} > 0$, reflecting the combined spread
between the two priming directions.

## Randomization and Attrition

### Randomization

Treatment was assigned entirely client-side at the moment a participant
clicked "I understand -- let's go" on the consent screen. The game
engine drew a single uniform random variate and mapped it to one of
three arms with nearly equal probability:

```         
r = Math.random()
Treatment = r < 0.333 ? 'no_label'
          : r < 0.666 ? 'hard_label'
          : 'easy_label'
```

`Math.random()` draws a number between 0 and 1 uniformly, with the unit
interval cut into three equal slices. Each arm was assigned with
approximately one-third probability. The treatment draw was written to a
`localStorage` draft immediately after assignment. Participants who
closed and returned mid-session were shown a "Welcome back" screen that
restored their original treatment from the draft, ensuring the draw was
not re-rolled on page reload. A participant's IP address was
additionally captured and stored under a separate `localStorage` key to
detect whether a device had previously completed a full session.
Returning devices were routed to a shared-device screening question
where participants self-reported whether they were a new or returning
player. Together, the IP-based flag and the self-report allowed us to
distinguish true first-time sessions from replays and from new
participants sharing a household device, retaining only genuine first
sessions.

### Recruitment Populations

This study drew from two distinct recruitment populations. Because the
recruitment source shapes both the attrition mechanism and potentially
the outcome measures themselves, we identify each participant's source
(organic or Prolific) and account for it throughout the analysis. The
organic sample was recruited through personal networks, friends, family,
and word of mouth, with no standardized referral message, no controlled
timing, and no single delivery channel. Organic participants had no
financial incentive to complete the study. The Prolific sample was
recruited through the Prolific platform, which delivers a standardized
study listing and compensates participants contingent on completion and
quality of effort. Prolific participants are likely to be experienced
survey-takers operating under platform norms around task completion.

### Attrition

Attrition is defined as beginning treatment but failing to complete all
five rounds. A participant is counted as an attritor if they have a user
record in the database, created at the moment they click "Start the
Challenge" and are shown a Round 1 label card, but their session was
never marked complete. The variable `sessionComplete` is set to TRUE
only when a participant reaches the results screen after resolving all
five rounds, each ending in either a correct answer or a skip.
Participants who passed through the survey and opted out before clicking
"Start the Challenge" are not counted as attritors because no user
record is created for them and they were never exposed to the treatment.

#### Differential Attrition in Organic and Prolific Subsamples

The attrition analysis fits logistic regression models predicting
dropout on treatment arm, estimated separately for the full sample and
the organic subsample, with no label as the reference arm. The Prolific
subsample had zero dropout across all arms and is reported descriptively
only.

```{r tab-attrition}
user_level <- master %>%
  distinct(userId, .keep_all = TRUE) %>%
  mutate(
    dropped   = as.integer(sessionComplete == FALSE),
    treatment = relevel(factor(treatment), ref = "no_label")
  )

m_att_full    <- glm(dropped ~ treatment, data = user_level, family = binomial)
m_att_organic <- glm(dropped ~ treatment,
                     data = filter(user_level, source == "Organic"),
                     family = binomial)

get_or_p <- function(model, term) {
  if (is.null(model)) return(c(NA, NA))
  s <- coef(summary(model))
  if (!term %in% rownames(s)) return(c(NA, NA))
  c(exp(s[term, "Estimate"]), s[term, "Pr(>|z|)"])
}

fmt_p <- function(p) {
  if (is.na(p)) return("---")
  ifelse(p < .001, "<.001", as.character(round(p, 3)))
}

or_lookup <- bind_rows(
  tibble(source = "Organic", treatment = "easy_label",
         OR = get_or_p(m_att_organic, "treatmenteasy_label")[1],
         p  = get_or_p(m_att_organic, "treatmenteasy_label")[2]),
  tibble(source = "Organic", treatment = "hard_label",
         OR = get_or_p(m_att_organic, "treatmenthard_label")[1],
         p  = get_or_p(m_att_organic, "treatmenthard_label")[2]),
  tibble(source = "Prolific", treatment = "easy_label", OR = NA, p = NA),
  tibble(source = "Prolific", treatment = "hard_label", OR = NA, p = NA),
  tibble(source = "Full", treatment = "easy_label",
         OR = get_or_p(m_att_full, "treatmenteasy_label")[1],
         p  = get_or_p(m_att_full, "treatmenteasy_label")[2]),
  tibble(source = "Full", treatment = "hard_label",
         OR = get_or_p(m_att_full, "treatmenthard_label")[1],
         p  = get_or_p(m_att_full, "treatmenthard_label")[2])
)

attrition_tbl <- user_level %>%
  group_by(source, treatment) %>%
  summarize(
    n_started     = n(),
    n_dropped     = sum(dropped),
    attrition_pct = paste0(round(mean(dropped) * 100, 1), "%"),
    .groups = "drop"
  )

get_pct <- function(src, trt) {
  attrition_tbl %>% filter(source == src, treatment == trt) %>% pull(attrition_pct)
}

org_no_pct   <- get_pct("Organic", "no_label")
org_easy_pct <- get_pct("Organic", "easy_label")
org_hard_pct <- get_pct("Organic", "hard_label")
org_easy_or  <- round(get_or_p(m_att_organic, "treatmenteasy_label")[1], 3)
org_easy_p   <- round(get_or_p(m_att_organic, "treatmenteasy_label")[2], 3)
org_hard_or  <- round(get_or_p(m_att_organic, "treatmenthard_label")[1], 3)
org_hard_p   <- round(get_or_p(m_att_organic, "treatmenthard_label")[2], 3)

attrition_tbl %>%
  mutate(treatment = factor(treatment,
                            levels = c("no_label", "easy_label", "hard_label"),
                            labels = c("No Label", "Easy Label", "Hard Label"))) %>%
  left_join(or_lookup %>%
              mutate(treatment = factor(treatment,
                                       levels = c("no_label", "easy_label", "hard_label"),
                                       labels = c("No Label", "Easy Label", "Hard Label"))),
            by = c("source", "treatment")) %>%
  mutate(
    OR = case_when(
      treatment == "No Label" ~ "ref.",
      is.na(OR)              ~ "---",
      TRUE                   ~ as.character(round(OR, 3))
    ),
    p = case_when(
      treatment == "No Label" ~ "---",
      is.na(p)               ~ "---",
      TRUE                   ~ sapply(p, fmt_p)
    )
  ) %>%
  arrange(source, treatment) %>%
  select(-source) %>%
  rename(
    Treatment      = treatment,
    `N Started`    = n_started,
    `N Dropped`    = n_dropped,
    `Dropout Rate` = attrition_pct,
    `Odds Ratio`   = OR,
    `p-value`      = p
  ) %>%
  kable(caption = "Attrition by Treatment Arm and Recruitment Source",
        booktabs = TRUE,
        align    = c("l", "c", "c", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = FALSE) %>%
  pack_rows("Organic",  1, 3) %>%
  pack_rows("Prolific", 4, 6) %>%
  footnote(general = "Odds ratios from logistic regression predicting dropout; reference arm: No Label. Prolific model omitted due to zero attrition. Full sample OR uses pooled model.",
           threeparttable = TRUE)
```

Among organic participants, dropout rates were `r org_no_pct` in the no
label arm, `r org_easy_pct` in the easy label arm, and `r org_hard_pct`
in the hard label arm. The easy label coefficient is negligible and
non-significant (OR = `r org_easy_or`, p = `r org_easy_p`), providing no
evidence of a difference from the control. The hard label arm shows a
suggestive but non-significant reduction in dropout odds relative to no
label (OR = `r org_hard_or`, p = `r org_hard_p`), consistent with H1's
prediction that a hard label primes persistence rather than
discouragement. The difference in dropout rates across arms is not
statistically significant, so we cannot conclude that differential
attrition meaningfully biases the primary outcome estimates. If the
pattern were real, the no label arm losing more participants than the
hard label arm would leave the no label completers as a comparatively
more select group, compressing the estimated difference between arms.
The Prolific subsample had zero dropout across all three arms,
consistent with payment-based compliance overriding any label effect on
attrition (see Table \@ref(tab:tab-attrition)).

## Completer Analysis & Covariate Balance

The primary analysis estimates average treatment effects among the 256
participants who completed all five rounds. Restricting to completers
introduces potential attrition bias if dropout is related to treatment.
Prolific participants had zero dropout across all three arms. Among
organic participants, differential attrition was not statistically
significant, but the hard label arm showed a suggestive reduction in
dropout relative to no label (OR = `r org_hard_or`, p = `r org_hard_p`).
If this pattern reflects a real effect, the no label arm would lose a
disproportionately higher share of participants, leaving its completers
as a more select group and compressing estimated treatment effect
differences toward zero.

```{r tab-balance}
bal <- df %>%
  distinct(userId, .keep_all = TRUE) %>%
  mutate(
    treatment   = relevel(factor(treatment), ref = "no_label"),
    is_prolific = as.integer(source == "Prolific"),
    is_desktop  = as.integer(deviceType == "Desktop"),
    is_female   = as.integer(gender == "Female")
  )

models <- list(
  "Age"            = lm(age                  ~ treatment, data = bal),
  "Quant Exp"      = lm(quantitativeExposure ~ treatment, data = bal),
  "Puzzle Skill"   = lm(puzzleSkill          ~ treatment, data = bal),
  "Puzzle Enjoy"   = lm(puzzleEnjoyment      ~ treatment, data = bal),
  "Energy"         = lm(energyLevel          ~ treatment, data = bal),
  "Recruit Source" = lm(is_prolific          ~ treatment, data = bal),
  "Desktop"        = lm(is_desktop           ~ treatment, data = bal),
  "Female"         = lm(is_female            ~ treatment, data = bal)
)

modelsummary(
  models,
  coef_map  = c("treatmenteasy_label" = "Easy Label",
                "treatmenthard_label" = "Hard Label"),
  stars     = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  gof_omit  = ".*",
  title     = "Covariate Balance Across Treatment Arms",
  notes     = "OLS coefficients vs. no-label reference arm. SEs in parentheses.",
  booktabs  = TRUE,
  output    = "kableExtra"
) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

Table \@ref(tab:tab-balance) presents covariate balance across the three
treatment arms among completers. Each column reports an OLS regression
of one baseline characteristic on treatment indicators, with no label as
the reference arm. None of the easy label or hard label coefficients
reach conventional significance thresholds on any covariate: age,
quantitative exposure, puzzle skill, puzzle enjoyment, energy level,
recruitment source, device type, or gender. The $R^2$ values are near
zero throughout, confirming that treatment assignment explains
essentially none of the variance in any baseline variable. This balance
supports the validity of the completer analysis: observed differences in
outcomes across arms are unlikely to reflect pre-existing differences in
participant composition.

## Discussion

[Interpretation of results, attrition limitations, and conclusions]
