---
title: "The Effect of Difficulty Labels on Puzzle Engagement"
author: ""
date: "`r Sys.Date()`"
output: bookdown::pdf_document2
bibliography: references.bib
header-includes:
  - \usepackage{booktabs}
  - \usepackage[table]{xcolor}
  - \usepackage{threeparttable}
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{graphicx}
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE, error = TRUE)
```

```{r load-data, include=FALSE}
library(dplyr)
library(tidyr)
library(purrr)
library(lme4)
library(clubSandwich)
library(knitr)
library(kableExtra)
library(modelsummary)
library(jsonlite)
library(consort)
library(ggplot2)
library(patchwork)
library(stargazer)

master      <- readRDS("analysis_data.rds")
df          <- master %>% filter(sessionComplete == TRUE)
df_prolific <- df %>% filter(source == "Prolific")
df_organic  <- df %>% filter(source == "Organic")

theme_set(theme_bw(base_size = 11))
```

## Introduction

How difficult a task feels can shape how much effort a person puts into
it. This experiment asked whether showing participants a difficulty
label before a logic puzzle — telling them it is easy or hard — changes
how long they stay engaged with it. Participants were given the same set
of hard puzzles and randomly assigned to one of three conditions: no
label, a hard label (accurate), or an easy label (understated). We
compare mean time engaged across the three groups to test whether the
label changed behavior. Prior work on perceived task difficulty suggests
that the information people have about a task, along with their
expectations going in, influences how much effort they invest. By
varying the label shown before each puzzle, we can manipulate those
expectations and observe the effect on engagement.

### Theory

Priming is when exposure to one stimulus influences how you respond to
the next. Research on semantic priming has shown that people recognize
related words faster when preceded by an associated cue
[@meyer1971facilitation], and neuroimaging work has found consistent
brain signatures of repeated stimulus exposure [@lee2020neural]. These
findings suggest that priming shapes cognitive processing in real,
measurable ways.

Our study looks at whether a difficulty label shown before a puzzle acts
as a prime that changes how long people stick with it. When someone sees
a label saying a task is easy, they form an expectation — and if the
task turns out harder than expected, that mismatch can affect their
behavior. @kahneman2011thinking argues that framing shapes intuitive
judgment before deliberate reasoning even kicks in, which is why a
difficulty label seen before the task might already shape how much
effort someone is willing to put in.

### Hypothesis

We hypothesize that the two label conditions affect time engaged in
opposite directions relative to the no label control. Participants shown
the hard label receive accurate information about task difficulty and
may mentally prepare for sustained effort, increasing persistence.
Participants shown the easy label encounter a mismatch between their
expectation and their experience: the puzzle is harder than primed,
which we expect to produce frustration that shortens engagement. When
participants expect a task to be easy and immediately find it is not,
the experience may feel like failure rather than challenge, lowering the
threshold for giving up and skipping. Formally:

$$H_1: \mu_{\text{hard}} > \mu_{\text{no label}}$$
$$H_2: \mu_{\text{easy}} < \mu_{\text{no label}}$$
$$H_3: \mu_{\text{hard}} > \mu_{\text{easy}}$$

where $\mu$ denotes mean time engaged per puzzle round.

## Power Calculation

To determine the necessary sample size, we conducted a simulation-based
power analysis rather than relying on a closed-form analytic formula.
This approach allows us to model the multi-round panel structure of the
experiment and account for participant-level clustering.

We set a control baseline average solve time of $\mu = 270$ seconds.
Variance was decomposed into a between-person component ($\sigma_u = 60$
seconds) and a within-person residual ($\sigma_e = 67$ seconds),
yielding a total standard deviation of
$\sigma = \sqrt{\sigma_u^2 + \sigma_e^2} \approx 90$ seconds, chosen
conservatively to capture natural individual variation in performance.
We specified treatment effects relative to the no label control in
opposite directions for the two label arms: a $+25$ second increase for
the hard label arm, reflecting anticipated persistence when participants
are accurately primed for difficulty, and a $-25$ second decrease for
the easy label arm, reflecting earlier disengagement when participants
encounter a task harder than the label suggested.

We simulated 500 iterations at escalating sample sizes. For each
iteration, a mixed-effects model was fit with puzzle fixed effects and a
participant-level random intercept:

$$Y_{it} = \beta_0 + \boldsymbol{\beta}_{\text{trt}} \text{Treatment}_i + \boldsymbol{\gamma}_t + u_i + \varepsilon_{it}$$

where $\boldsymbol{\gamma}_t$ are puzzle fixed effects and
$u_i \sim N(0, \sigma_u^2)$ is a participant random intercept. Because
data are generated exactly from the model, standard mixed-model
$t$-tests are used to assess significance; cluster-robust standard
errors are applied in the actual analysis but are not needed for
simulation where the variance structure is known.

```{r power-sim, message=FALSE, warning=FALSE}
if (file.exists("power_sim_results.rds")) {
  power_res <- readRDS("power_sim_results.rds")
} else {
  set.seed(241)
  mu_base   <- 270; beta_hard <- 25; beta_easy <- -25
  sigma_u   <-  60; sigma_e   <- 67; n_puzzles  <-  5; n_sim <- 500
  puzzle_fx <- c(-15, 10, 20, -8, -7)

  sim_power <- function(n_per_arm) {
    sig_hard <- sig_easy <- logical(n_sim)
    for (s in seq_len(n_sim)) {
      n_total  <- n_per_arm * 3
      trt_id   <- rep(c("no_label","easy_label","hard_label"), each = n_per_arm)
      u_i      <- rnorm(n_total, 0, sigma_u)
      userId   <- rep(seq_len(n_total), each = n_puzzles)
      puzzleId <- rep(seq_len(n_puzzles), times = n_total)
      trt_long <- rep(trt_id, each = n_puzzles)
      beta_trt <- ifelse(trt_long == "hard_label", beta_hard,
                  ifelse(trt_long == "easy_label", beta_easy, 0))
      Y <- mu_base + beta_trt + puzzle_fx[puzzleId] + u_i[userId] +
           rnorm(n_total * n_puzzles, 0, sigma_e)
      dat <- data.frame(
        Y         = Y,
        treatment = factor(trt_long, levels = c("no_label","easy_label","hard_label")),
        puzzleId  = factor(puzzleId),
        userId    = factor(userId)
      )
      m <- tryCatch(lmer(Y ~ treatment + puzzleId + (1|userId),
                         data = dat, REML = FALSE), error = function(e) NULL)
      if (is.null(m)) next
      ct <- summary(m)$coefficients
      th <- ct["treatmenthard_label", "t value"]
      te <- ct["treatmenteasy_label", "t value"]
      if (length(th)) sig_hard[s] <- abs(th) > 1.96
      if (length(te)) sig_easy[s] <- abs(te) > 1.96
    }
    data.frame(n_per_arm = n_per_arm,
               power_hard = mean(sig_hard),
               power_easy = mean(sig_easy))
  }

  power_res <- bind_rows(lapply(seq(40, 150, by = 10), sim_power))
  saveRDS(power_res, "power_sim_results.rds")
}

thresh_hard <- power_res$n_per_arm[which(power_res$power_hard >= 0.80)[1]]
thresh_easy <- power_res$n_per_arm[which(power_res$power_easy >= 0.80)[1]]
thresh_both <- max(thresh_hard, thresh_easy, na.rm = TRUE)
```

At $\alpha = 0.05$, power to detect the hard label effect crosses the
80% threshold at `r thresh_hard` participants per arm
(`r thresh_hard * 3` total), and the easy label effect crosses 80% at
`r thresh_easy` per arm (`r thresh_easy * 3` total). The
harder-to-detect effect requires `r thresh_both` per arm
(`r thresh_both * 3` total).

The final analytic sample of 256 completers falls below the
pre-specified target of `r thresh_both * 3`. Exclusions for behavioral
flags and attrition reduced the sample beyond what was anticipated at
the design stage, and arm sizes are unequal (No Label: 87, Easy Label:
80, Hard Label: 89), which modestly reduces power relative to equal
allocation. Beyond sample size, the analytic sample pools two
recruitment populations that differ systematically in engagement
conditions: Prolific participants completed the experiment in a
structured, incentivized setting with low distraction, while organic
participants, recruited through personal networks, varied more in
motivation, setting, and distraction level. This heterogeneity inflates
within-group variance beyond what a single $\sigma$ captures, further
eroding effective power. Both the sample shortfall and the variance
inflation from pooling two behaviorally distinct populations should be
kept in mind when interpreting null or marginal findings.

## Study Design

### Procedure

The experiment was delivered as a browser-based web application at
anotherpuzzle.io. The application was a single-page JavaScript game
engine that managed the complete participant session from recruitment
intake through behavioral data collection without any researcher
intervention at runtime. All data were logged directly to a cloud
database (Firestore) at both the session and round levels.

Upon arriving at the application, participants were presented with a
consent screen. Participants who agreed were immediately and invisibly
assigned to one of three treatment arms before proceeding. Following
consent, participants completed an eight-item survey collecting baseline
covariates: age, gender, education level, quantitative exposure, current
energy level, puzzle frequency, self-rated puzzle skill, and puzzle
enjoyment. After the survey, participants were shown a brief instruction
screen and then clicked "Start the Challenge," which triggered
randomization of puzzle order and the creation of the participant's
session record.

The game engine logged behavioral data at the round level. For each
round, it recorded: total time engaged (a continuous timer that paused
automatically during tab switches and participant-initiated breaks),
number of answer attempts, whether the first attempt was correct, time
to first attempt, tab-switch count and total time paused, rage-click
events, and whether the round ended in a solve or a skip. Skip was
enabled after 30 seconds of engagement, preventing immediate abandonment
without any attempt. At the session level, the engine recorded device
type, puzzle presentation order, recruitment source, and session
completion status.

### Treatment

The experimental treatment was a difficulty label presented to
participants before each logic puzzle. The hard label was expected to
prompt mental preparation for a demanding task, increasing persistence.
The easy label was expected to create an expectation mismatch:
participants anticipating an easy task would encounter greater
difficulty than expected, producing frustration that shortens engagement
before participants choose to skip.

Participants were randomly assigned at session entry to one of three
conditions: (1) no label, (2) easy label, or (3) hard label. The
assigned label was shown before every puzzle in the session, so a
participant assigned to the hard label condition saw that label at the
start of each of the five rounds. Within the application, all other
aspects of the experience were held constant across arms. The interface,
instructions, and the five logic puzzles were identical across
conditions, so differences in behavior across treatment groups can be
attributed to the difficulty label rather than any other feature of the
experience.

Each session consisted of five logic puzzles presented in randomized
order: a number sequence, an alphabet wheel, a shape sequence, a hybrid
shape-number puzzle, and a tile sequence. All five were designed to be
intentionally challenging, though they varied in how difficult
participants found them in practice.

```{r puzzle-skip-rate}
overall_rates <- df %>%
  mutate(
    Puzzle = case_when(
      puzzleId == "puzzle_01" ~ "Number Sequence",
      puzzleId == "puzzle_02" ~ "Alphabet Wheel",
      puzzleId == "puzzle_03" ~ "Shape Sequence",
      puzzleId == "puzzle_04" ~ "Hybrid Puzzle",
      puzzleId == "puzzle_05" ~ "Tile Sequence"
    )
  ) %>%
  group_by(Puzzle) %>%
  summarize(overall_raw = mean(didSkip, na.rm = TRUE), .groups = "drop")

df %>%
  mutate(
    Puzzle = case_when(
      puzzleId == "puzzle_01" ~ "Number Sequence",
      puzzleId == "puzzle_02" ~ "Alphabet Wheel",
      puzzleId == "puzzle_03" ~ "Shape Sequence",
      puzzleId == "puzzle_04" ~ "Hybrid Puzzle",
      puzzleId == "puzzle_05" ~ "Tile Sequence"
    )
  ) %>%
  group_by(Puzzle, treatment) %>%
  summarize(skip_rate = mean(didSkip, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = treatment, values_from = skip_rate) %>%
  left_join(overall_rates, by = "Puzzle") %>%
  arrange(desc(overall_raw)) %>%
  mutate(across(c(no_label, easy_label, hard_label, overall_raw),
                ~ paste0(round(. * 100, 1), "%"))) %>%
  select(Puzzle, no_label, easy_label, hard_label, overall_raw) %>%
  rename(
    `No Label`   = no_label,
    `Easy Label` = easy_label,
    `Hard Label` = hard_label,
    `Overall`    = overall_raw
  ) %>%
  kable(booktabs = TRUE,
        caption  = "Skip Rate by Puzzle and Treatment Arm") %>%
  kable_styling(latex_options = c("hold_position"), font_size = 10)
```

Skip and solve rates differed meaningfully across puzzle types, as shown
in Table \@ref(tab:puzzle-skip-rate). Puzzle order was randomized across
participants to prevent any one puzzle from systematically falling at
the same point in the session, and the analysis accounts for
puzzle-level differences and within-participant correlation across
rounds.

Treatment delivery was controlled entirely through the application, and
the data contain no indication that any participant was shown a label
other than their assigned condition; we therefore identified no
treatment non-compliers. However, the application did not independently
log the randomized assignment alongside the label rendered on screen for
each round, so perfect delivery could not be independently verified.
Future experiments should log these separately to enable a formal
compliance check.

## Participants

The experiment database contained 398 unique participants. Eight were
excluded as returning or replay players whose devices had previously
completed a full session, leaving 390 first-time participants. An
additional 27 had no retained gamelog records and were excluded,
yielding 363 participants who started at least one puzzle round.

Twenty-one participants were excluded for eligibility: 8 were under 18
and 13 were rejected through Prolific's quality review. Of the remaining
342 eligible participants, 39 were removed for behavioral or
data-quality reasons:

*Behavioral exclusions*

1.  Total tab-switch pause time exceeded 600 seconds across the session,
    indicating the participant was away from the browser for an extended
    period.
2.  At least one puzzle showed an idle walk-away pattern: the
    participant skipped without switching tabs and with time engaged
    above the 97th percentile of the skip-time distribution for that
    puzzle, suggesting they left the page and returned.
3.  All five puzzles were skipped within 30 to 35 seconds with no answer
    attempts, indicating the participant waited out the skip-unlock
    timer on every round rather than engaging with the puzzles.

*AI assistance exclusions*

4.  Self-reported use of AI assistance in the post-session honesty
    check.
5.  Hard puzzles solved on the first attempt faster than population
    benchmarks for that puzzle (any single puzzle below the 5th
    percentile, or two or more below the 10th), subject to manual
    review. Flagged participants were retained if their fast solve times
    were consistent with their relative performance across all rounds,
    as genuine skill tends to produce proportionally fast times across
    puzzles rather than selectively fast times on isolated ones.

Attrition, defined as beginning treatment but failing to complete all
five rounds, accounted for 47 participants. The final analytic sample is
256 completers across three arms: No Label (n = 87), Easy Label (n =
80), and Hard Label (n = 89). Figure \@ref(fig:fig-consort) shows the
full participant flow.

```{r fig-consort, fig.width=9, fig.height=7, fig.cap="Participant flow through the experiment.", fig.pos="H", out.width="\\textwidth"}

raw_json  <- fromJSON("firestore_export.json", simplifyVector = FALSE)
users_all <- tibble(document_id = names(raw_json$users),
                    data        = raw_json$users) %>%
  unnest_wider(data) %>%
  distinct(userId, .keep_all = TRUE)

new_player_ids <- users_all %>%
  filter(!is.na(sharedDevice) & sharedDevice == "new_player") %>%
  pull(userId)

em_u          <- readRDS("experiment_master.rds") %>% distinct(userId, .keep_all = TRUE)
master_u      <- master %>% distinct(userId, .keep_all = TRUE)
completed_ids <- master_u %>% filter(sessionComplete == TRUE) %>% pull(userId)
dropped_ids   <- master_u %>% filter(sessionComplete == FALSE) %>% pull(userId)

arm_src <- master %>%
  filter(sessionComplete == TRUE) %>%
  distinct(userId, .keep_all = TRUE) %>%
  count(treatment, source)

gn <- function(trt, src) {
  v <- arm_src %>% filter(treatment == trt, source == src) %>% pull(n)
  if (length(v) == 0) 0L else v
}
n_no_org   <- gn("no_label",   "Organic")
n_no_pro   <- gn("no_label",   "Prolific")
n_easy_org <- gn("easy_label", "Organic")
n_easy_pro <- gn("easy_label", "Prolific")
n_hard_org <- gn("hard_label", "Organic")
n_hard_pro <- gn("hard_label", "Prolific")

flow <- users_all %>%
  mutate(
    exc_replay = case_when(
      !is.na(sharedDevice) & sharedDevice == "returning" ~ "Returning player",
      !is.na(isReplay) & isReplay == TRUE & !(userId %in% new_player_ids) ~ "Replay session",
      TRUE ~ NA_character_
    ),
    after_replay = if_else(is.na(exc_replay), userId, NA_character_),
    exc_no_game = case_when(
      !is.na(exc_replay) ~ NA_character_,
      !(userId %in% em_u$userId) ~ "No gamelog records",
      TRUE ~ NA_character_
    ),
    started = if_else(
      is.na(exc_replay) & (userId %in% em_u$userId), userId, NA_character_
    ),
    exc_elig = case_when(
      !is.na(exc_replay) | !(userId %in% em_u$userId) ~ NA_character_,
      !is.na(age) & age < 18 ~ "Age < 18",
      !is.na(prolificRejected) & prolificRejected == TRUE ~ "Prolific rejection",
      TRUE ~ NA_character_
    ),
    eligible = if_else(
      is.na(exc_replay) & (userId %in% em_u$userId) & is.na(exc_elig),
      userId, NA_character_
    ),
    exc_behav = case_when(
      !is.na(exc_replay) | !(userId %in% em_u$userId) | !is.na(exc_elig) ~ NA_character_,
      !(userId %in% master_u$userId) ~ "Behavioral or AI flag",
      TRUE ~ NA_character_
    ),
    quality = if_else(
      is.na(exc_replay) & (userId %in% em_u$userId) &
        is.na(exc_elig) & (userId %in% master_u$userId),
      userId, NA_character_
    ),
    exc_att = case_when(
      !is.na(exc_replay) | !(userId %in% em_u$userId) |
        !is.na(exc_elig) | !(userId %in% master_u$userId) ~ NA_character_,
      userId %in% dropped_ids ~ "Did not complete all 5 rounds",
      TRUE ~ NA_character_
    ),
    arm = case_when(
      !(userId %in% completed_ids) ~ NA_character_,
      treatment == "no_label" ~
        paste0("No Label\nOrganic: ", n_no_org, "\nProlific: ", n_no_pro, "\n"),
      treatment == "easy_label" ~
        paste0("Easy Label\nOrganic: ", n_easy_org, "\nProlific: ", n_easy_pro, "\n"),
      treatment == "hard_label" ~
        paste0("Hard Label\nOrganic: ", n_hard_org, "\nProlific: ", n_hard_pro, "\n")
    )
  )

p <- consort_plot(
  data   = flow,
  orders = c(
    userId       = 'Users with a session record',
    exc_replay   = 'Excluded',
    after_replay = 'Non-replay participants',
    exc_no_game  = 'Excluded',
    started      = 'Participants who started',
    exc_elig     = 'Excluded',
    eligible     = 'Eligible participants',
    exc_behav    = 'Excluded',
    quality      = 'Passed quality checks',
    exc_att      = 'Did not complete all 5 rounds',
    arm          = 'Analyzed'
  ),
  side_box   = c('exc_replay', 'exc_no_game', 'exc_elig', 'exc_behav', 'exc_att'),
  allocation = 'arm',
  cex        = 0.7
)

plot(p)
```

## Potential Outcomes

Each participant has three potential outcomes — how long they would
spend on the puzzles under each label condition: $Y_i(\text{no label})$,
$Y_i(\text{easy label})$, and $Y_i(\text{hard label})$. We only ever
observe one of these per person. Because participants were randomly
assigned, the treatment someone received is independent of what their
outcome would have been under a different condition, which means we can
estimate average outcomes for each arm using the observed group means.

The main comparisons we care about are the treatment effects of each
label relative to the no-label control, and the direct comparison
between the two label arms. Each $\tau$ below is the difference in group
means ($\mu$) between two arms:

$$\tau_{\text{easy}} = \mu_{\text{easy}} - \mu_{\text{no label}} = E[Y_i(\text{easy label})] - E[Y_i(\text{no label})]$$

$$\tau_{\text{hard}} = \mu_{\text{hard}} - \mu_{\text{no label}} = E[Y_i(\text{hard label})] - E[Y_i(\text{no label})]$$

$$\tau_{\text{hard vs. easy}} = \mu_{\text{hard}} - \mu_{\text{easy}} = E[Y_i(\text{hard label})] - E[Y_i(\text{easy label})]$$

The no-label arm is the reference group — participants who saw no
framing at all. The first two comparisons ask whether giving someone any
difficulty expectation changes their engagement relative to having none.
The third asks whether the direction of the label matters: we expect the
easy and hard labels to push behavior in opposite directions, so
$\tau_{\text{easy}} < 0$ (frustration leads to less engagement),
$\tau_{\text{hard}} > 0$ (preparation leads to more), and
$\tau_{\text{hard vs. easy}} > 0$.

## Randomization and Attrition

### Randomization

Treatment was assigned entirely client-side at the moment a participant
clicked "I understand -- let's go" on the consent screen. The game
engine drew a single uniform random variate and mapped it to one of
three arms with nearly equal probability:

```         
r = Math.random()
Treatment = r < 0.333 ? 'no_label'
          : r < 0.666 ? 'hard_label'
          : 'easy_label'
```

`Math.random()` draws a number between 0 and 1 uniformly, with the unit
interval cut into three equal slices. Each arm was assigned with
approximately one-third probability. The treatment draw was written to a
`localStorage` draft immediately after assignment. Participants who
closed and returned mid-session were shown a "Welcome back" screen that
restored their original treatment from the draft, ensuring the draw was
not re-rolled on page reload. A participant's IP address was
additionally captured and stored under a separate `localStorage` key to
detect whether a device had previously completed a full session.
Returning devices were routed to a shared-device screening question
where participants self-reported whether they were a new or returning
player. Together, the IP-based flag and the self-report allowed us to
distinguish true first-time sessions from replays and from new
participants sharing a household device, retaining only genuine first
sessions.

### Recruitment Populations

We recruited participants from two different sources, and we track this
throughout the analysis because recruitment context likely affects both
dropout and how people engage with the task. Organic participants were
recruited informally through personal networks — friends, family, and
word of mouth — with no financial incentive to complete the study.
Prolific participants were recruited through the Prolific platform,
which pays participants for completing studies. Prolific workers tend to
be experienced study participants who are used to following task
instructions carefully and completing what they start.

### Attrition

Attrition is defined as beginning treatment but failing to complete all
five rounds. A participant is counted as an attritor if they have a user
record in the database, created at the moment they click "Start the
Challenge" and are shown a Round 1 label card, but their session was
never marked complete. The variable `sessionComplete` is set to TRUE
only when a participant reaches the results screen after resolving all
five rounds, each ending in either a correct answer or a skip.
Participants who passed through the survey and opted out before clicking
"Start the Challenge" are not counted as attritors because no user
record is created for them and they were never exposed to the treatment.

#### Differential Attrition in Organic and Prolific Subsamples

The attrition analysis fits logistic regression models predicting
dropout on treatment arm, estimated separately for the full sample and
the organic subsample, with no label as the reference arm. The Prolific
subsample had zero dropout across all arms and is reported descriptively
only.

```{r tab-attrition}
user_level <- master %>%
  distinct(userId, .keep_all = TRUE) %>%
  mutate(
    dropped   = as.integer(sessionComplete == FALSE),
    treatment = relevel(factor(treatment), ref = "no_label")
  )

m_att_full    <- glm(dropped ~ treatment, data = user_level, family = binomial)
m_att_organic <- glm(dropped ~ treatment,
                     data = filter(user_level, source == "Organic"),
                     family = binomial)

get_or_p <- function(model, term) {
  if (is.null(model)) return(c(NA, NA))
  s <- coef(summary(model))
  if (!term %in% rownames(s)) return(c(NA, NA))
  c(exp(s[term, "Estimate"]), s[term, "Pr(>|z|)"])
}

fmt_p <- function(p) {
  if (is.na(p)) return("---")
  ifelse(p < .001, "<.001", as.character(round(p, 3)))
}

or_lookup <- bind_rows(
  tibble(source = "Organic", treatment = "easy_label",
         OR = get_or_p(m_att_organic, "treatmenteasy_label")[1],
         p  = get_or_p(m_att_organic, "treatmenteasy_label")[2]),
  tibble(source = "Organic", treatment = "hard_label",
         OR = get_or_p(m_att_organic, "treatmenthard_label")[1],
         p  = get_or_p(m_att_organic, "treatmenthard_label")[2]),
  tibble(source = "Prolific", treatment = "easy_label", OR = NA, p = NA),
  tibble(source = "Prolific", treatment = "hard_label", OR = NA, p = NA),
  tibble(source = "Full", treatment = "easy_label",
         OR = get_or_p(m_att_full, "treatmenteasy_label")[1],
         p  = get_or_p(m_att_full, "treatmenteasy_label")[2]),
  tibble(source = "Full", treatment = "hard_label",
         OR = get_or_p(m_att_full, "treatmenthard_label")[1],
         p  = get_or_p(m_att_full, "treatmenthard_label")[2])
)

attrition_tbl <- user_level %>%
  group_by(source, treatment) %>%
  summarize(
    n_started     = n(),
    n_dropped     = sum(dropped),
    attrition_pct = paste0(round(mean(dropped) * 100, 1), "%"),
    .groups = "drop"
  )

get_pct <- function(src, trt) {
  attrition_tbl %>% filter(source == src, treatment == trt) %>% pull(attrition_pct)
}

org_no_pct   <- get_pct("Organic", "no_label")
org_easy_pct <- get_pct("Organic", "easy_label")
org_hard_pct <- get_pct("Organic", "hard_label")
org_easy_or  <- round(get_or_p(m_att_organic, "treatmenteasy_label")[1], 3)
org_easy_p   <- round(get_or_p(m_att_organic, "treatmenteasy_label")[2], 3)
org_hard_or  <- round(get_or_p(m_att_organic, "treatmenthard_label")[1], 3)
org_hard_p   <- round(get_or_p(m_att_organic, "treatmenthard_label")[2], 3)

attrition_tbl %>%
  mutate(treatment = factor(treatment,
                            levels = c("no_label", "easy_label", "hard_label"),
                            labels = c("No Label", "Easy Label", "Hard Label"))) %>%
  left_join(or_lookup %>%
              mutate(treatment = factor(treatment,
                                       levels = c("no_label", "easy_label", "hard_label"),
                                       labels = c("No Label", "Easy Label", "Hard Label"))),
            by = c("source", "treatment")) %>%
  mutate(
    OR = case_when(
      treatment == "No Label" ~ "ref.",
      is.na(OR)              ~ "---",
      TRUE                   ~ as.character(round(OR, 3))
    ),
    p = case_when(
      treatment == "No Label" ~ "---",
      is.na(p)               ~ "---",
      TRUE                   ~ sapply(p, fmt_p)
    )
  ) %>%
  arrange(source, treatment) %>%
  select(-source) %>%
  rename(
    Treatment      = treatment,
    `N Started`    = n_started,
    `N Dropped`    = n_dropped,
    `Dropout Rate` = attrition_pct,
    `Odds Ratio`   = OR,
    `p-value`      = p
  ) %>%
  kable(caption = "Attrition by Treatment Arm and Recruitment Source",
        booktabs = TRUE,
        align    = c("l", "c", "c", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = FALSE, font_size = 10) %>%
  pack_rows("Organic",  1, 3) %>%
  pack_rows("Prolific", 4, 6) %>%
  footnote(general = "Odds ratios from logistic regression predicting dropout; reference arm: No Label. Prolific model omitted due to zero attrition. Full sample OR uses pooled model.",
           threeparttable = TRUE)
```

Among organic participants, dropout rates were `r org_no_pct` in the no
label arm, `r org_easy_pct` in the easy label arm, and `r org_hard_pct`
in the hard label arm. The easy label coefficient is negligible and
non-significant (OR = `r org_easy_or`, p = `r org_easy_p`), providing no
evidence of a difference from the control. The hard label arm shows a
suggestive but non-significant reduction in dropout odds relative to no
label (OR = `r org_hard_or`, p = `r org_hard_p`), consistent with H1's
prediction that a hard label primes persistence rather than
discouragement. The difference in dropout rates across arms is not
statistically significant, so we cannot conclude that differential
attrition meaningfully biases the primary outcome estimates. If the
pattern were real, the no label arm losing more participants than the
hard label arm would leave the no label completers as a comparatively
more select group, compressing the estimated difference between arms.
The Prolific subsample had zero dropout across all three arms,
consistent with payment-based compliance overriding any label effect on
attrition (see Table \@ref(tab:tab-attrition)).

## Completer Analysis & Covariate Balance

The primary analysis estimates average treatment effects among the 256
participants who completed all five rounds. Restricting to completers
introduces potential attrition bias if dropout is related to treatment.
Prolific participants had zero dropout across all three arms. Among
organic participants, differential attrition was not statistically
significant, but the hard label arm showed a suggestive reduction in
dropout relative to no label (OR = `r org_hard_or`, p = `r org_hard_p`).
If this pattern reflects a real effect, the no label arm would lose a
disproportionately higher share of participants, leaving its completers
as a more select group and compressing estimated treatment effect
differences toward zero.

```{r tab-balance}
bal <- df %>%
  distinct(userId, .keep_all = TRUE) %>%
  mutate(
    treatment   = relevel(factor(treatment), ref = "no_label"),
    is_prolific = as.integer(source == "Prolific"),
    is_desktop  = as.integer(deviceType == "Desktop"),
    is_female   = as.integer(gender == "Female")
  )

models <- list(
  "Age"            = lm(age                  ~ treatment, data = bal),
  "Quant Exp"      = lm(quantitativeExposure ~ treatment, data = bal),
  "Puzzle Skill"   = lm(puzzleSkill          ~ treatment, data = bal),
  "Puzzle Enjoy"   = lm(puzzleEnjoyment      ~ treatment, data = bal),
  "Energy"         = lm(energyLevel          ~ treatment, data = bal),
  "Recruit Source" = lm(is_prolific          ~ treatment, data = bal),
  "Desktop"        = lm(is_desktop           ~ treatment, data = bal),
  "Female"         = lm(is_female            ~ treatment, data = bal)
)

modelsummary(
  models,
  coef_map  = c("treatmenteasy_label" = "Easy Label",
                "treatmenthard_label" = "Hard Label"),
  stars     = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  gof_omit  = ".*",
  title     = "Covariate Balance Across Treatment Arms",
  notes     = "OLS coefficients vs. no-label reference arm. SEs in parentheses.",
  booktabs  = TRUE,
  output    = "kableExtra"
) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

Table \@ref(tab:tab-balance) presents covariate balance across the three
treatment arms among completers. Each column reports an OLS regression
of one baseline characteristic on treatment indicators, with no label as
the reference arm. None of the easy label or hard label coefficients
reach conventional significance thresholds on any covariate: age,
quantitative exposure, puzzle skill, puzzle enjoyment, energy level,
recruitment source, device type, or gender. The $R^2$ values are near
zero throughout, confirming that treatment assignment explains
essentially none of the variance in any baseline variable. This balance
supports the validity of the completer analysis: observed differences in
outcomes across arms are unlikely to reflect pre-existing differences in
participant composition.

## Limitations, Models & Analysis

### Limitations of the Pooled Subsamples

To assess whether recruitment source introduced systematic differences
in data quality or engagement patterns, we examined four behavioral
indicators across Prolific and organic participants: across-person
variability in log time engaged, within-person variability in log time
engaged, tab switch frequency, and total time spent with the browser tab
inactive.

```{r tab-within-sd, results="asis", message=FALSE, warning=FALSE}
person_engage <- df %>%
  group_by(userId, source) %>%
  summarize(mean_logTE   = mean(logTE, na.rm = TRUE),
            sd_logTE     = sd(logTE, na.rm = TRUE),
            total_tabs   = sum(tabSwitchCount, na.rm = TRUE),
            total_paused = sum(tabSwitchTimePaused, na.rm = TRUE),
            .groups = "drop") %>%
  group_by(source) %>%
  mutate(across_sd = abs(mean_logTE - mean(mean_logTE))) %>%
  ungroup()

m_across    <- lm(across_sd    ~ source, data = person_engage)
m_within_sd <- lm(sd_logTE     ~ source, data = person_engage)
m_tabs      <- lm(total_tabs   ~ source, data = person_engage)
m_paused    <- lm(total_paused ~ source, data = person_engage)

stargazer(m_across, m_within_sd, m_tabs, m_paused,
          type             = "latex",
          title            = "Behavioral Differences by Recruitment Source",
          dep.var.labels   = c("Across-Person SD", "Within-Person SD", "Tab Switches", "Time Paused (s)"),
          covariate.labels = "Prolific (vs. Organic)",
          omit.stat        = c("f", "ser"),
          header           = FALSE,
          table.placement  = "H")
```

Across the four indicators, two differences emerge. Across-person
variability in log time engaged does not differ significantly between
groups, suggesting that the two subsamples are comparable in their
overall engagement levels. Within-person variability, however, is
significantly higher among organic participants: organic participants
were more erratic across their five puzzle rounds than Prolific
participants. One plausible explanation is that organic participants,
recruited informally through personal networks, varied more in their
environment, distraction level, and motivation from one puzzle to the
next. A participant might have been fully focused on one round and
interrupted on another.

Organic participants also spent significantly more total time with the
browser tab inactive, which points to more real-world interruptions
during the session. Tab switch counts did not differ significantly
between groups. Prolific participants likely kept more consistent
conditions throughout, given the platform's completion norms. Either
way, higher within-person variability among organic participants
increases the residual variance in the model, which means wider
confidence intervals and less power to detect a treatment effect. So the
null result doesn't necessarily mean there's no effect — it may just
mean we didn't have enough precision to detect a small one.

### Primary Model: Effect of Difficulty Labels on Log Time Engaged

#### Outcome Distribution

Raw time engaged is strongly right-skewed, with a long tail from
participants who spent disproportionately long on a single puzzle. Log
transformation is applied to reduce skew and stabilize variance before
modeling.

```{r fig-log-justification, fig.width=8, fig.height=3.5, fig.cap="Distribution of time engaged before and after log transformation.", fig.pos="H", out.width="\\textwidth"}
m1_raw <- ggplot(df, aes(x = timeEngaged)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  labs(x = "Time Engaged (seconds)", y = "Count", title = "Raw")

m1_log <- ggplot(df, aes(x = logTE)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  labs(x = "Log Time Engaged", y = "Count", title = "Log-Transformed")

m1_raw + m1_log
```

As shown in Figure \@ref(fig:fig-log-justification), the log
transformation substantially reduces skew and produces an approximately
symmetric distribution.

#### Model Specification

We estimated the effect of difficulty labels on log time engaged using a
linear mixed-effects model with puzzle fixed effects, a
participant-level random intercept, and pre-treatment covariates:

$$\log(\text{TimeEngaged}_{it}) = \beta_0 + \beta_1 \text{EasyLabel}_i + \beta_2 \text{HardLabel}_i + \mathbf{X}_i'\boldsymbol{\delta} + \varepsilon_{it}$$

where $\mathbf{X}_i$ includes puzzle skill, recruitment source, age,
quantitative exposure, energy level, and device type. Standard errors
are clustered at the participant level using CR1 with Satterthwaite
degrees of freedom.

```{r primary-model}
m <- lmer(logTE ~ treatment + puzzleId + puzzleSkill + source +
            age + quantitativeExposure + energyLevel + deviceType + (1 | userId),
          data = df, REML = FALSE)

vcov_m <- vcovCR(m, cluster = df$userId, type = "CR1")
rt     <- coef_test(m, vcov = vcov_m, test = "Satterthwaite")

modelsummary(
  list("Log Time Engaged" = m),
  vcov      = vcov_m,
  statistic = c("SE: {std.error}", "p: {p.value}"),
  coef_map  = c("treatmenteasy_label" = "Easy Label",
                "treatmenthard_label" = "Hard Label"),
  gof_omit  = ".*",
  stars     = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  title     = "Effect of Difficulty Labels on Log Time Engaged",
  booktabs  = TRUE,
  output    = "kableExtra"
) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

```{r m1-vals, include=FALSE}
m1_easy_b  <- round(rt$beta[rt$Coef == "treatmenteasy_label"], 3)
m1_easy_se <- round(rt$SE[rt$Coef == "treatmenteasy_label"], 3)
m1_easy_p  <- round(rt$p_Satt[rt$Coef == "treatmenteasy_label"], 3)
m1_hard_b  <- round(rt$beta[rt$Coef == "treatmenthard_label"], 3)
m1_hard_se <- round(rt$SE[rt$Coef == "treatmenthard_label"], 3)
m1_hard_p  <- round(rt$p_Satt[rt$Coef == "treatmenthard_label"], 3)
m1_hard_pct <- round((1 - exp(rt$beta[rt$Coef == "treatmenthard_label"])) * 100, 1)
```

Neither label had a statistically significant effect on log time engaged
relative to the no-label control (Table \@ref(tab:primary-model)). The
easy label coefficient is essentially zero ($\hat{\beta}_{\text{easy}}$
= `r m1_easy_b`, SE = `r m1_easy_se`, p = `r m1_easy_p`), so we see no
evidence of the frustration and disengagement we predicted in H2. The
hard label coefficient is actually negative and marginally
non-significant ($\hat{\beta}_{\text{hard}}$ = `r m1_hard_b`, SE =
`r m1_hard_se`, p = `r m1_hard_p`), corresponding to about
`r m1_hard_pct`% less time engaged — the opposite direction from what H1
predicted.

```{r fig-coef, fig.width=5, fig.height=2.5, fig.pos="H", fig.align="center", out.width="0.65\\textwidth", fig.cap="Point estimates and 95\\% confidence intervals for the effect of difficulty labels on log time engaged relative to the no label control. Error bars are based on cluster-robust standard errors."}
coef_df <- data.frame(
  term = c("Easy Label", "Hard Label"),
  est  = c(rt$beta[rt$Coef == "treatmenteasy_label"],
           rt$beta[rt$Coef == "treatmenthard_label"]),
  se   = c(rt$SE[rt$Coef == "treatmenteasy_label"],
           rt$SE[rt$Coef == "treatmenthard_label"])
) %>%
  mutate(lo = est - 1.96 * se, hi = est + 1.96 * se)

ggplot(coef_df, aes(x = est, y = term)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(aes(xmin = lo, xmax = hi), width = 0.15) +
  geom_point(size = 3) +
  labs(x = "Estimated Effect on Log Time Engaged", y = NULL,
       title    = "Effect of Difficulty Labels on Engagement",
       subtitle = "Relative to no-label control; error bars are 95% CIs") +
  theme_bw(base_size = 9)
```

As shown in Figure \@ref(fig:fig-coef), the confidence intervals for
both estimates are wide and cross zero, so we can't say much about the
true direction of either effect. H3 is also not supported — the hard
label point estimate is actually below the easy label estimate, the
opposite of what we predicted.

### Secondary Model: Effect of Difficulty Labels on Zero-Attempt Skips

#### Outcome Description

In our secondary model, we are looking at zero-attempt skips. A
zero-attempt skip is when a participant skips the puzzle round without
entering any response. This could reflect several things: a participant
who looked at the puzzle and eventually gave up, or one who spent time
thinking through it, but ultimately could not arrive at an answer
confident enough to submit. What this model captures is the absence of a
committed attempt, regardless of how much cognitive effort proceeded
that decision.

```{r tab-skip-rate}
skip_by_puzzle <- df %>%
  mutate(puzzle = case_when(
    puzzleId == "puzzle_01" ~ "Number Sequence",
    puzzleId == "puzzle_02" ~ "Alphabet Wheel",
    puzzleId == "puzzle_03" ~ "Shape Sequence",
    puzzleId == "puzzle_04" ~ "Hybrid Puzzle",
    puzzleId == "puzzle_05" ~ "Tile Sequence"
  )) %>%
  group_by(puzzle, treatment) %>%
  summarize(rate = mean(zero_att_skip, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = treatment, values_from = rate) %>%
  rename("No Label" = no_label, "Easy Label" = easy_label, "Hard Label" = hard_label)

overall <- df %>%
  group_by(treatment) %>%
  summarize(rate = mean(zero_att_skip, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = treatment, values_from = rate) %>%
  rename("No Label" = no_label, "Easy Label" = easy_label, "Hard Label" = hard_label) %>%
  mutate(puzzle = "Overall")

bind_rows(skip_by_puzzle, overall) %>%
  mutate(across(where(is.numeric), ~ paste0(round(. * 100, 1), "%"))) %>%
  rename(Puzzle = puzzle) %>%
  kable(booktabs = TRUE, align = c("l", "c", "c", "c"),
        caption = "Zero-attempt skip rate by puzzle and treatment arm.") %>%
  kable_styling(latex_options = c("hold_position"), font_size = 10) %>%
  row_spec(nrow(skip_by_puzzle) + 1, bold = TRUE)
```

#### Model Specification

We model zero-attempt skips using a Linear Probability Model (LPM),
estimating the probability of a zero-attempt skip as a linear function
of treatment assignment and covariates:

$$\text{ZeroAttSkip}_{it} = \beta_0 + \beta_1 \text{EasyLabel}_i + \beta_2 \text{HardLabel}_i + \mathbf{X}_i'\boldsymbol{\delta} + \varepsilon_{it}$$

where $\mathbf{X}_i$ includes puzzle skill, age, quantitative exposure,
energy level, and device type. The full-sample model additionally
controls for recruitment source. Standard errors are clustered by
participant (CR1, Satterthwaite degrees of freedom), and coefficients
are interpreted as percentage-point differences in zero-attempt skip
probability relative to the no-label control.

Zero-attempt skips are heavily concentrated in two puzzles. As shown in
Table \@ref(tab:tab-skip-rate), Number Sequence and Tile Sequence
account for the vast majority of cold skips, with overall rates of 23%
and 20% respectively, while Shape Sequence and Hybrid Puzzle see
virtually none. Alphabet Wheel falls in between at around 10%. The easy
label arm consistently shows lower skip rates across puzzles, while hard
label is slightly elevated relative to no label — a pattern that holds
across puzzle types rather than being driven by any single puzzle.

```{r zero-att-model, include=FALSE}
fit_zas <- function(data, control_source = FALSE) {
  fmla <- if (control_source)
    zero_att_skip ~ treatment + puzzleId + puzzleSkill + source +
      age + quantitativeExposure + energyLevel + deviceType
  else
    zero_att_skip ~ treatment + puzzleId + puzzleSkill +
      age + quantitativeExposure + energyLevel + deviceType
  m  <- lm(fmla, data = data)
  vc <- vcovCR(m, cluster = data$userId, type = "CR1")
  rt <- coef_test(m, vcov = vc, test = "Satterthwaite")
  list(m = m, se = sqrt(diag(vc)), p = rt$p_Satt, rt = rt)
}

za_full <- fit_zas(df, control_source = TRUE)
za_org  <- fit_zas(df_organic)
za_pro  <- fit_zas(df_prolific)
```

```{r tab-zas, results="asis", message=FALSE, warning=FALSE}
stargazer(za_full$m, za_org$m, za_pro$m,
          se               = list(za_full$se, za_org$se, za_pro$se),
          p                = list(za_full$p,  za_org$p,  za_pro$p),
          type             = "latex",
          title            = "Effect of Difficulty Labels on Zero-Attempt Skips",
          dep.var.caption  = "",
          dep.var.labels   = c("Full", "Organic", "Prolific"),
          multicolumn      = FALSE,
          covariate.labels = c("Easy Label", "Hard Label"),
          keep             = c("treatmenteasy_label", "treatmenthard_label"),
          report           = "vc*sp",
          omit.stat        = c("f", "ser"),
          label            = "tab:tab-zas",
          header           = FALSE,
          table.placement  = "H",
          font.size        = "small",
          notes            = "LPM; cluster-robust SEs (CR1) by participant.")
```

```{r zas-vals, include=FALSE}
easy_full_pp <- round(abs(za_full$rt$beta[za_full$rt$Coef == "treatmenteasy_label"]) * 100, 1)
easy_full_p  <- round(za_full$rt$p_Satt[za_full$rt$Coef == "treatmenteasy_label"], 3)
hard_full_b  <- round(za_full$rt$beta[za_full$rt$Coef == "treatmenthard_label"], 3)
hard_full_p  <- round(za_full$rt$p_Satt[za_full$rt$Coef == "treatmenthard_label"], 3)
easy_pro_pp  <- round(abs(za_pro$rt$beta[za_pro$rt$Coef == "treatmenteasy_label"]) * 100, 1)
easy_pro_p   <- round(za_pro$rt$p_Satt[za_pro$rt$Coef == "treatmenteasy_label"], 3)
easy_org_b   <- round(za_org$rt$beta[za_org$rt$Coef == "treatmenteasy_label"], 3)
easy_org_p   <- round(za_org$rt$p_Satt[za_org$rt$Coef == "treatmenteasy_label"], 3)
```

The linear model estimates the effect of difficulty labels on the
probability of skipping a puzzle without any answer attempt (Table
\@ref(tab:tab-zas)). The hard label shows no meaningful effect no matter
how we slice the sample — the coefficient is near zero and
non-significant in the full sample ($\hat{\beta}_{\text{hard}}$ =
`r hard_full_b`, p = `r hard_full_p`), among organic participants, and
among Prolific participants.

The easy label effect on zero-attempt skips is driven entirely by
Prolific participants. Among Prolific participants, the easy label
reduces the zero-attempt skip rate by `r easy_pro_pp` percentage points
(p = `r easy_pro_p`), while among organic participants the effect is
small and non-significant ($\hat{\beta}_{\text{easy}}$ = `r easy_org_b`,
p = `r easy_org_p`). When the two groups are pooled and recruitment
source is controlled for, the organic null attenuates the Prolific
signal, yielding a marginally significant estimate of `r easy_full_pp`
percentage points (p = `r easy_full_p`) in the full sample.

```{r fig-skip-coef, fig.width=5, fig.height=2.5, fig.pos="H", fig.align="center", fig.cap="Estimated effect of difficulty labels on zero-attempt skip probability (linear probability model). Coefficients represent percentage-point change relative to the no-label control. Error bars are 95\\% CIs based on cluster-robust SEs.", out.width="0.65\\textwidth"}
rt <- za_full$rt
coef_df <- data.frame(
  term = c("Easy Label", "Hard Label"),
  est  = c(rt$beta[rt$Coef == "treatmenteasy_label"],
           rt$beta[rt$Coef == "treatmenthard_label"]),
  se   = c(rt$SE[rt$Coef == "treatmenteasy_label"],
           rt$SE[rt$Coef == "treatmenthard_label"])
) %>% mutate(lo = est - 1.96 * se, hi = est + 1.96 * se)

ggplot(coef_df, aes(x = est, y = term)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(aes(xmin = lo, xmax = hi), width = 0.15) +
  geom_point(size = 3) +
  labs(x = "Effect on Zero-Attempt Skip Rate (pp)", y = NULL,
       title    = "Effect of Difficulty Labels on Zero-Attempt Skips",
       subtitle = "Relative to no-label control; error bars are 95% CIs") +
  theme_bw(base_size = 9)
```

As shown in Figure \@ref(fig:fig-skip-coef), the easy label coefficient
is negative and pointing toward fewer cold skips relative to no label.
However, thee confidence interval is wide and crosses zero in the full
sample. The hard label estimate is close to zero with a wide interval in
both directions, so there's no real signal there.

The difference in the easy label's effect between Prolific and organic
participants likely reflects differences in how each group engages with
task framing. Prolific participants are experienced study participants
who tend to read instructions carefully and take task cues seriously —
when told a puzzle is easy, they internalize that label and may feel a
stronger obligation to at least attempt an answer before skipping.
Organic participants, recruited informally through personal networks,
tend to be more distracted and less attentive to framing; we also found
they spent significantly more time with the browser tab inactive,
suggesting more real-world interruptions that could override any label
effect. In short, the easy label appears to work by making cold-skipping
feel less acceptable, but that mechanism requires the participant to be
paying attention to the framing in the first place.

## Appendix

### A. Residual Diagnostics

```{r fig-qq, fig.width=5, fig.height=5, fig.pos="H", fig.cap="Normal Q-Q plot of residuals from the primary model (log time engaged ~ difficulty label + puzzle fixed effects + covariates + participant random intercept). Residuals track the reference line closely across the central range, with minor deviation in the lower tail, consistent with a light left tail after log transformation."}
qqnorm(residuals(m), main = "Normal Q-Q Plot of Model Residuals")
qqline(residuals(m), col = "steelblue")
```
